---
title: "analysis"
author: "Krishna Anujan"
date: "20/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r adding env columns}
library(raster)

msg <- read.csv("data/raw_data/ObsLevel_Sept22.csv")
# dem_stack
# light_vals

msg_coords <- msg
coordinates(msg_coords) <- ~ long_new2 + lat_new2

dem_stack <- stack("C://Users/krish/Documents/SeasonWatch/MSG/other_data/env_vars/dem_stack")
lights_KL <- raster("C://Users/krish/Documents/SeasonWatch/MSG/other_data/env_vars/lights_KL")
dem_vals <- extract(dem_stack, msg_coords)
light_vals <- extract(lights_KL, msg_coords)

# writing the dem stack separately
# elev<-writeRaster(dem_stack$elev, "C://Users/krish/Documents/SeasonWatch/MSG/other_data/env_vars/elev.grd")

# slope<-writeRaster(dem_stack$slope, "C://Users/krish/Documents/SeasonWatch/MSG/other_data/env_vars/slope.grd")

# aspect<-writeRaster(dem_stack$aspect, "C://Users/krish/Documents/SeasonWatch/MSG/other_data/env_vars/aspect.grd")

msg <- cbind(msg, dem_vals, light_vals)
# write.csv(msg, "fromMSG/All3_ObservationLevel_spat.csv", row.names=F)

sw_2014 <- read.csv("data/modified_data/sw_msg_from2014.csv")
sw_spat <- msg

sw_spat$Date <- sw_2014$Date2[match(sw_spat$Observation_ID, sw_2014$Observation_ID)]
sw_spat$day <- sw_2014$date[match(sw_spat$Observation_ID, sw_2014$Observation_ID)]
sw_spat$month <- sw_2014$month[match(sw_spat$Observation_ID, sw_2014$Observation_ID)]
sw_spat$year <- sw_2014$year[match(sw_spat$Observation_ID, sw_2014$Observation_ID)]
sw_spat$User_Tree_id <- sw_2014$User_Tree_id[match(sw_spat$Observation_ID, sw_2014$Observation_ID)]
sw_spat$User_id <- sw_2014$User_id[match(sw_spat$Observation_ID, sw_2014$Observation_ID)]
sw_spat$is.KL2 <- sw_2014$is.KL2[match(sw_spat$Observation_ID, sw_2014$Observation_ID)]
# write.csv(sw_spat, "fromMSG/All3_ObservationLevel_spat.csv", row.names=F)
```

```{r correlations between variables}
library(sp)
library(lubridate)
library(plyr)
library(ggplot2)
library(tidyr)


# sw_spat<-read.csv("fromMSG/All3_ObservationLevel_spat.csv") #has DEM and night lights and Tree ID etc

sw_spat <- sw_spat[sw_spat$is.KL2 == 1, ]

sw_spat_forcor <- ddply(sw_spat, .(Species_name, User_Tree_id), summarise,
  lat = mean(lat_new2), long = mean(long_new2),
  elev = mean(elev), slope = mean(slope),
  aspect = mean(aspect), light = mean(light_vals),
  pcp = mean(FNSum_PCP), maxtemp = mean(FNMean_MaxTemp),
  mintemp = mean(FNMean_MinTemp), meantemp = mean(FNMean_MeanTemp),
  SM = mean(FNMean_MeanSM), DryDays = mean(FNSum_ConsDryDays),
  mn_pcp = mean(MNSum_PCP), mn_maxtemp = mean(MNMean_MaxTemp),
  mn_mintemp = mean(MNMean_MinTemp), mn_meantemp = mean(MNMean_MeanTemp),
  mn_SM = mean(MNMean_MeanSM), mn_DryDays = mean(MNSum_ConsDryDays),
  qr_pcp = mean(QrSum_PCP), qr_maxtemp = mean(QrMean_MaxTemp),
  qr_mintemp = mean(QrMean_MinTemp), qr_meantemp = mean(QrMean_MeanTemp),
  qr_SM = mean(QrMean_MeanSM), qr_DryDays = mean(QrSum_ConsDryDays)
)

# png("plots/distr_jack.png", width=8, height=8, res=300, units="in")
ggplot(gather(sw_spat_forcor[sw_spat_forcor$Species_name == "Artocarpus heterophyllus", 3:8]), aes(value)) +
  geom_histogram(bins = 10) +
  ggtitle("jackfruit") +
  facet_wrap(~key, scales = "free_x")
# dev.off()

# png("plots/distr_mango.png", width=8, height=8, res=300, units="in")
ggplot(gather(sw_spat_forcor[sw_spat_forcor$Species_name == "Mangifera indica", 3:8]), aes(value)) +
  geom_histogram(bins = 10) +
  ggtitle("mango") +
  facet_wrap(~key, scales = "free_x")
# dev.off()

# png("plots/distr_tam.png", width=8, height=8, res=300, units="in")
ggplot(gather(sw_spat_forcor[sw_spat_forcor$Species_name == "Tamarindus indica", 3:8]), aes(value)) +
  geom_histogram(bins = 10) +
  ggtitle("tamarind") +
  facet_wrap(~key, scales = "free_x")
# dev.off()

spatcor <- cor(sw_spat_forcor[, 3:8], use = "complete.obs")

# png("plots/spat_cor.png", width=4, height=4, res=300, units="in")
ggcorrplot::ggcorrplot(spatcor, lab = T)
# dev.off()

spattempcor <- cor(sw_spat_forcor[, 3:19], use = "complete.obs")

# png("plots/spattemp_cor.png", width=8, height=8, res=300, units="in")
ggcorrplot::ggcorrplot(spattempcor, lab = T, type = "lower")
# dev.off()


sw_fn_cor <- ddply(sw_spat, .(Grid_ID, month), summarise,
  fn_pcp = mean(FNSum_PCP), fn_maxtemp = mean(FNMean_MaxTemp),
  fn_mintemp = mean(FNMean_MinTemp), fn_meantemp = mean(FNMean_MeanTemp),
  fn_SM = mean(FNMean_MeanSM), fn_DryDays = mean(FNSum_ConsDryDays),
  mn_pcp = mean(MNSum_PCP), mn_maxtemp = mean(MNMean_MaxTemp),
  mn_mintemp = mean(MNMean_MinTemp), mn_meantemp = mean(MNMean_MeanTemp),
  mn_SM = mean(MNMean_MeanSM), mn_DryDays = mean(MNSum_ConsDryDays),
  qr_pcp = mean(QrSum_PCP), qr_maxtemp = mean(QrMean_MaxTemp),
  qr_mintemp = mean(QrMean_MinTemp), qr_meantemp = mean(QrMean_MeanTemp),
  qr_SM = mean(QrMean_MeanSM), qr_DryDays = mean(QrSum_ConsDryDays)
)

fn_cor <- cor(sw_fn_cor[, 3:ncol(sw_fn_cor)])

png("doc/display/SFig_corr_2.png", width = 8, height = 8, res = 300, units = "in")
ggcorrplot::ggcorrplot(fn_cor, lab = T, type = "lower")
dev.off()

# scale everything
# https://stackoverflow.com/questions/41766181/correct-way-to-scale-for-multilevel-regression-using-lmer-r

sw_spat_rescale <- transform(sw_spat,
  FNSum_PCP_cs = scale(FNSum_PCP),
  FNMean_MaxTemp_cs = scale(FNMean_MaxTemp),
  FNMean_MeanSM_cs = scale(FNMean_MeanSM),
  FNSum_DryDays_cs = scale(FNSum_ConsDryDays),
  MNSum_PCP_cs = scale(MNSum_PCP),
  MNMean_MaxTemp_cs = scale(MNMean_MaxTemp),
  MNMean_MeanSM_cs = scale(MNMean_MeanSM),
  MNSum_DryDays_cs = scale(MNSum_ConsDryDays),
  QrSum_PCP_cs = scale(QrSum_PCP),
  QrMean_MaxTemp_cs = scale(QrMean_MaxTemp),
  QrMean_MeanSM_cs = scale(QrMean_MeanSM),
  QrSum_DryDays_cs = scale(QrSum_ConsDryDays),
  elev_cs = scale(elev), slope_cs = scale(slope),
  light_cs = scale(light_vals), aspect_cs = scale(aspect),
  lat_sc = scale(lat_new2), long_sc = scale(long_new2)
)


spat_cs_cor <- cor(sw_spat_rescale[, 54:59], use = "complete.obs")
ggcorrplot::ggcorrplot(spat_cs_cor, lab = T, type = "lower")
```

```{r explore models}
# library(bbmle)
library(lme4)



# flbin_glmm<-glmer(fl_binary~FNSum_PCP+FNMean_MaxTemp+FNMean_MeanSM+FNSum_DryDays+
#                    MNSum_PCP+MNMean_MaxTemp+MNMean_MeanSM+
#                    QrSum_PCP+QrMean_MeanTemp+QrMean_MeanSM+
#                    elev+slope+light_vals+(1|User_Tree_id)+
#                    (1|Species_name), data=sw_spat, family="binomial") #Species as a fixed effect

# all species together

flbin_glmm_cs <- glmer(
  fl_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs + FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs + slope_cs + light_cs + Species_name + # Maybe remove SM
    # lat_sc+long_sc+
    (1 | User_Tree_id) + (1 | year), #+(1|User_id)
  data = sw_spat_rescale,
  family = "binomial"
)

frbin_glmm_cs <- glmer(
  fr_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs + FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs + slope_cs + light_cs + Species_name +
    # lat_sc+long_sc+
    (1 | User_Tree_id) + (1 | year),
  data = sw_spat_rescale,
  family = "binomial"
)


frbin_glmm_cs <- glmer(
  fr_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs + FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs + slope_cs + light_cs + Species_name +
    # lat_sc+long_sc+
    (1 | User_Tree_id) + (1 | year),
  data = sw_spat_rescale,
  family = "binomial"
)

library(sjPlot)

tab_model(flbin_glmm_cs, frbin_glmm_cs, file = "plots/flbin_frbin_all_cs.html")

# flbin_glmm_jack<-glmer(fl_binary~FNSum_PCP+FNMean_MaxTemp+FNMean_MeanSM+FNSum_DryDays+
# MNSum_PCP+MNMean_MaxTemp+MNMean_MeanSM+
# QrSum_PCP+QrMean_MeanTemp+QrMean_MeanSM+
#                    elev+slope+light_vals+(1|User_Tree_id),
#                    data=sw_spat[sw_spat$Species_name=="Artocarpus heterophyllus",],
#                    family="binomial")

flbin_glmm_jack_cs <- glmer(
  fl_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs + FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs + slope_cs + light_cs +
    # lat_sc+long_sc+
    (1 | User_Tree_id) + (1 | year),
  data = sw_spat_rescale[sw_spat_rescale$Species_name == "Artocarpus heterophyllus", ],
  family = "binomial"
)
flbin_glmm_mango_cs <- glmer(
  fl_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs + FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs + slope_cs + light_cs +
    # lat_new2+long_new2+
    (1 | User_Tree_id) + (1 | year),
  data = sw_spat_rescale[sw_spat_rescale$Species_name == "Mangifera indica", ],
  family = "binomial"
)
flbin_glmm_tam_cs <- glmer(
  fl_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs + FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs + slope_cs + light_cs +
    # lat_new2+long_new2+
    (1 | User_Tree_id) + (1 | year),
  data = sw_spat_rescale[sw_spat_rescale$Species_name == "Tamarindus indica", ],
  family = "binomial"
)

# Fruiting separate models
frbin_glmm_jack_cs <- glmer(
  fr_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs + FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs + slope_cs + light_cs +
    # lat_sc+long_sc+
    (1 | User_Tree_id) + (1 | year),
  data = sw_spat_rescale[sw_spat_rescale$Species_name == "Artocarpus heterophyllus", ],
  family = "binomial"
)
frbin_glmm_mango_cs <- glmer(
  fr_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs + FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs + slope_cs + light_cs +
    # lat_new2+long_new2+
    (1 | User_Tree_id) + (1 | year),
  data = sw_spat_rescale[sw_spat_rescale$Species_name == "Mangifera indica", ],
  family = "binomial"
)
frbin_glmm_tam_cs <- glmer(
  fr_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs + FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs + slope_cs + light_cs +
    # lat_new2+long_new2+
    (1 | User_Tree_id) + (1 | year),
  data = sw_spat_rescale[sw_spat_rescale$Species_name == "Tamarindus indica", ],
  family = "binomial"
)


# summary(flbin_glmm_jack_cs)

library(sjPlot)


tab_model(flbin_glmm_jack_cs, flbin_glmm_mango_cs, flbin_glmm_tam_cs, file = "plots/flbin_3sp_cs.html")
tab_model(frbin_glmm_jack_cs, frbin_glmm_mango_cs, frbin_glmm_tam_cs, file = "plots/frbin_3sp_cs.html")
```

```{r ordinal models}
# borrowing code from here - https://drizopoulos.github.io/GLMMadaptive/articles/Ordinal_Mixed_Models.html

# for proportion odds, code from here-
# https://www.r-bloggers.com/2019/06/how-to-perform-ordinal-logistic-regression-in-r/
# and https://www.rdocumentation.org/packages/ordinal/versions/2019.12-10/topics/clmm

library(GLMMadaptive)

# make the response variable a factor
sw_spat_rescale$fl_intensity_fac <- factor(sw_spat_rescale$fl_intensity, levels = c("0", "1", "2"))
sw_spat_rescale$fr_intensity_fac <- factor(sw_spat_rescale$fr_intensity, levels = c("0", "1", "2"))

# setting up the data for ordinal analysis
cr_vals_fl <- cr_setup(sw_spat_rescale$fl_intensity_fac)
cr_data_fl <- sw_spat_rescale[cr_vals_fl$subs, ]
cr_data_fl$y_new <- cr_vals_fl$y
cr_data_fl$cohort <- cr_vals_fl$cohort

fl_int_mod <- mixed_model(
  y_new ~ FNSum_PCP_cs + FNMean_MaxTemp_cs + Species_name +
    FNMean_MeanSM_cs + FNSum_DryDays_cs + elev_cs + slope_cs + light_cs,
  random = ~ 1 | User_Tree_id,
  data = cr_data_fl, family = binomial()
)

fl_int_mod
# library(sjPlot)

# setting up the data for ordinal analysis for fruiting
cr_vals_fr <- cr_setup(sw_spat_rescale$fr_intensity_fac)
cr_data_fr <- sw_spat_rescale[cr_vals_fr$subs, ]
cr_data_fr$y_new <- cr_vals_fr$y
cr_data_fr$cohort <- cr_vals_fr$cohort

fr_int_mod <- mixed_model(
  y_new ~ FNSum_PCP_cs + FNMean_MaxTemp_cs + Species_name +
    FNMean_MeanSM_cs + FNSum_DryDays_cs + elev_cs + slope_cs + light_cs,
  random = ~ 1 | User_Tree_id,
  data = cr_data_fr, family = binomial()
)

fr_int_mod
# library(sjPlot)
tab_model(fl_int_mod, fr_int_mod, file = "plots/fl_fr_int_mod.html")
```

```{r ordinal models for individual species}
# ordinal models for each species separately--------------------

library(MASS)

library(ordinal)

# cr_data_fl$Species_namefac<-factor(cr_data_fl$Species_name, levels=c("Artocarpus heterophyllus", "Mangifera indica", "Tamarindus indica"))

# make the response variable a factor
sw_spat_rescale$fl_intensity_fac <- factor(sw_spat_rescale$fl_intensity, levels = c("0", "1", "2"), ordered = T)
sw_spat_rescale$fr_intensity_fac <- factor(sw_spat_rescale$fr_intensity, levels = c("0", "1", "2"), ordered = T)


# Dividing data into training and test set
# Random sampling

# jack-----------------------------------
jack <- sw_spat_rescale[sw_spat_rescale$Species_name == "Artocarpus heterophyllus" & sw_spat_rescale$is.KL2 == 1, ]

samplesize <- 0.70 * nrow(jack)
set.seed(100)
index <- sample(seq_len(nrow(jack)), size = samplesize)
# Creating training and test set
jacktrain <- jack[index, ]
jacktest <- jack[-index, ]


# fl_int_jack_po<- clmm2(fl_intensity_fac ~ FNSum_PCP_cs+FNMean_MaxTemp_cs+
#                          FNMean_MeanSM_cs+FNSum_DryDays_cs+elev_cs+aspect_cs+light_cs,
#                          (1 | User_Tree_id),
#                  data = jacktrain, Hess=T)

# summary(fl_int_jack_po)
# fl_int_jack_po

ctable_jack <- coef(summary(fl_int_jack_po))
write.csv(ctable_jack, "plots/coefs_jack_fl.csv")

# p<-pnorm(abs(ctable_jack[,"t value"]), lower.tail = F)*2
# ctable_jack<-cbind(ctable_jack, "p value"=p)
ci <- confint(fl_int_jack_po)

jacktrain$User_Tree_id <- as.factor(jacktrain$User_Tree_id)

fl_inter_jack_po <- clmm2(
  fl_intensity_fac ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs,
  random = User_Tree_id,
  data = jacktrain, Hess = T
)

summary(fl_inter_jack_po)
fl_inter_jack_po

pred_jack <- predict(fl_inter_jack_po, newdata = jacktest)
plot(jacktest$fl_intensity, pred_jack)

saveRDS(fl_inter_jack_po, file = "models/fl_inter_jack_po.rds")
```
```{r plotting probabilities}
library(effects)

png("plots/jack_SM_aspect_effect.png", width = 8, height = 8, res = 300, units = "in")
plot(Effect(focal.predictors = c("FNMean_MeanSM_cs", "aspect_cs"), fl_inter_jack_po), rug = T)
dev.off()

png("plots/jack_DryDays_light_effect.png", width = 8, height = 8, res = 300, units = "in")
plot(Effect(focal.predictors = c("FNSum_DryDays_cs", "light_cs"), fl_inter_jack_po), rug = T)
dev.off()

png("plots/jack_SumPCP_elev_effect.png", width = 8, height = 8, res = 300, units = "in")
plot(Effect(focal.predictors = c("FNSum_PCP_cs", "elev_cs"), fl_inter_jack_po), rug = T)
dev.off()
```

{r predictions}

#Solution 1----------------------
#prediction using test data - https://gist.github.com/Martin-Jung/33ca6591193cb59f0a88

# Arguments:
#  - model = a clmm model
#  - modelAvg = a clmm model average (object of class averaging)
#  - newdata = a dataframe of new data to apply the model to
# Returns a dataframe of predicted probabilities for each row and response level
#fake.predict.clmm <- function(modelAvg, newdata) {
fake.predict.clmm <- function(model, newdata) {
  # Actual prediction function
  pred <- function(eta, theta, cat = 1:(length(theta) + 1), inv.link = plogis) {
    Theta <- c(-1000, theta, 1000)
    sapply(cat, function(j) inv.link(Theta[j + 1] - eta) - inv.link(Theta[j] - 
eta))
  }

  # Multiply each row by the coefficients
  coefs <- c(model$beta, unlist(model$ST))##turn off if a model average is used
  
  #beta <- modelAvg$coefficients[2,3:12]
  #coefs <- c(beta, unlist(modelAvg$ST))

  xbetas <- sweep(newdata, MARGIN=2, coefs, `*`)

  # Make predictions
  #Theta<-modelAvg$coefficients[2,1:2]
  pred.mat <- data.frame(pred(eta=rowSums(xbetas), theta=model$Theta))
  
  #trying
  #pred.mat <- data.frame(pred(eta=xbetas, theta=model$Theta))
  
  #pred.mat <- data.frame(pred(eta=rowSums(xbetas), theta=Theta))
  colnames(pred.mat) <- levels(model$model[,1])
  
  #a<-attr(modelAvg, "modelList")
  #colnames(pred.mat) <- levels(a[[1]]$model[,1])

  pred.mat
}

jack.pred<-fake.predict.clmm(fl_int_jack_po, newdata=jacktest[,match(names(coefs), colnames(jacktest))])

#r-squared of prediction and test data
plot(jacktest$fl_intensity, jack.pred)


#Solution 2-------------------------------------------
#https://petrilankoski.com/2013/09/20/model-comparison-with-data/

library("ordinal")
library("ggplot2")
pred <- function(eta, theta, cat = 1:(length(theta) + 1), inv.link = plogis) {
  Theta <- c(-1000, theta, 1000)
  sapply(cat, function(j) inv.link(Theta[j + 1] - eta) - inv.link(Theta[j] - eta))}

plot.probabilities3<-function(grid, model, comp.data=NULL, title="", ylim=NULL) {
  co <- model$coefficients[1:length(model$y.levels)-1]
  pre.mat <- pred(eta=rowSums(grid), theta=co)
  df<-data.frame(levels=as.numeric(model$y.levels))
  df["avg"] <- pre.mat[1,]
  df["low"] <- pre.mat[2,]
  df["high"] <- pre.mat[3,]
  if(!is.null(comp.data)) {
     df["freq"] <- summary(comp.data)/sum(summary(comp.data))
  }
  plot1 <- ggplot(data=df, aes(x=levels, y=avg)) + geom_line() + geom_point() +
  ggtitle(title) + ylab("probability") + xlab("") +
  geom_line(aes(x=levels, y=low), colour="#CCCCCC") +
  geom_line(aes(x=levels, y=high), colour="#CCCCCC")
  if(!is.null(comp.data)) {
     plot1 <- plot1 +
     geom_line(aes(x=levels, y=freq), lty="dotted")
  }
  if(!is.null(ylim)) {
     plot1 <- plot1 + ylim(0, ylim)
  }
  return(plot1)
}


library("ggplot2")
pred <- function(eta, theta, cat = 1:(length(theta) + 1), inv.link = plogis) {
  Theta <- c(-1000, theta, 1000)
  sapply(cat, function(j) inv.link(Theta[j + 1] - eta) - inv.link(Theta[j] - eta))}

plot.probabilities3<-function(grid, model, comp.data=NULL, title="", ylim=NULL) {
  co <- model$coefficients[1:length(model$y.levels)-1]
  pre.mat <- pred(eta=rowSums(grid), theta=co)
  df<-data.frame(levels=as.numeric(model$y.levels))
  df["avg"] <- pre.mat[1,]
  df["low"] <- pre.mat[2,]
  df["high"] <- pre.mat[3,]
  if(!is.null(comp.data)) {
     df["freq"] <- summary(comp.data)/sum(summary(comp.data))
  }
  plot1 <- ggplot(data=df, aes(x=levels, y=avg)) + geom_line() + geom_point() +
  ggtitle(title) + ylab("probability") + xlab("") +
  geom_line(aes(x=levels, y=low), colour="#CCCCCC") +
  geom_line(aes(x=levels, y=high), colour="#CCCCCC")
  if(!is.null(comp.data)) {
     plot1 <- plot1 +
     geom_line(aes(x=levels, y=freq), lty="dotted")
  }
  if(!is.null(ylim)) {
     plot1 <- plot1 + ylim(0, ylim)
  }
  return(plot1)
}


#data(wine)
#fm2 <- clmm(rating ~ temp + contact + (1|judge), data = wine,Hess = TRUE, nAGQ = 10)
mat_no_cold <- expand.grid(judge = qnorm(0.95) * c(0, -1, 1) * fm2$stDev, contact = c(0), temp=c(0))
mat_yes_cold <- expand.grid(judge = qnorm(0.95) * c(0, -1, 1) * fm2$stDev, contact = c(fm2$beta[2]), temp=c(0))
mat_no_warm <- expand.grid(judge = qnorm(0.95) * c(0, -1, 1) * fm2$stDev, contact = c(0), temp=c(fm2$beta[1]))
mat_yes_warm <- expand.grid(judge = qnorm(0.95) * c(0, -1, 1) * fm2$stDev, contact = c(fm2$beta[2]), temp=c(fm2$beta[1]))


jacktest2<-jacktest[complete.cases(jacktest),]
jack.pred.plot<-plot.probabilities3(newdata, fl_int_jack_po, jacktest2$fl_intensity_fac)
jack.pred.plot

dat_no_cold.tmp<-wine[wine["contact"]=="no",]
dat_no_cold<-dat_no_cold.tmp[dat_no_cold.tmp["temp"]=="cold",]
plot1<-plot.probabilities3(mat_no_cold, fm2, dat_no_cold$rating, title="Contact=no,Temperature=cold")

dat_yes_cold.tmp<-wine[wine["contact"]=="yes",]
dat_yes_cold<-dat_yes_cold.tmp[dat_yes_cold.tmp["temp"]=="cold",]
plot2<-plot.probabilities3(mat_yes_cold, fm2, dat_yes_cold$rating, title="Contact=yes,Temperature=cold")

dat_no_warm.tmp<-wine[wine["contact"]=="no",]
dat_no_warm<-dat_no_warm.tmp[dat_no_warm.tmp["temp"]=="warm",]
plot3<-plot.probabilities3(mat_no_warm, fm2, dat_no_warm$rating, title="Contact=no,Temperature=warm")

dat_yes_warm.tmp<-wine[wine["contact"]=="yes",]
dat_yes_warm<-dat_yes_warm.tmp[dat_yes_warm.tmp["temp"]=="warm",]
plot4<-plot.probabilities3(mat_yes_warm, fm2, dat_yes_warm$rating, title="Contact=yes,Temperature=warm")

grid.arrange(plot1, plot2, plot3, plot4, ncol=2)

#Solution 3------------------------------------

#https://marissabarlaz.github.io/portfolio/ols/

library(ggeffects)
ggpred_jack <- data.frame(ggpredict(fl_int_jack_po, terms = c("FNSum_PCP_cs", "FNSum_DryDays_cs"), type = "fe"))
colnames(ggpred_jack)[6]<-"fl_intensity"
ggplot(ggpred_jack, aes(x=fl_intensity, y=predicted))+geom_point()

```{r others}
# mango------------------------------------
mango <- sw_spat_rescale[sw_spat_rescale$Species_name == "Mangifera indica" & sw_spat_rescale$is.KL2 == 1, ]

samplesize <- 0.70 * nrow(mango)
set.seed(100)
index <- sample(seq_len(nrow(mango)), size = samplesize)
# Creating training and test set
mangotrain <- mango[index, ]
mangotest <- mango[-index, ]


mangotrain$User_Tree_id <- as.factor(mangotrain$User_Tree_id)

fl_inter_mango_po <- clmm2(
  fl_intensity_fac ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs,
  random = User_Tree_id,
  data = mangotrain, Hess = T
)

summary(fl_inter_mango_po)
# saveRDS(fl_inter_mango_po, file="models/fl_inter_mango_po.rds")

pred_mango <- predict(fl_inter_jack_po, newdata = mangotest)
plot(mangotest$fl_intensity_fac, pred_mango)


ctable_mango <- coef(summary(fl_int_mango_po))
# write.csv(ctable_mango, "plots/coefs_mango_fl.csv")
ci <- confint(fl_int_mango_po)

# plots
# png("plots/mango_SM_aspect_effect.png", width=8, height=8, res=300, units="in")
plot(Effect(focal.predictors = c("FNMean_MeanSM_cs", "aspect_cs"), fl_inter_mango_po), rug = T)
# dev.off()

# png("plots/mango_DryDays_light_effect.png", width=8, height=8, res=300, units="in")
plot(Effect(focal.predictors = c("FNSum_DryDays_cs", "light_cs"), fl_inter_mango_po), rug = T)
# dev.off()

# png("plots/mango_SumPCP_elev_effect.png", width=8, height=8, res=300, units="in")
plot(Effect(focal.predictors = c("FNSum_PCP_cs", "elev_cs"), fl_inter_mango_po), rug = T)
# dev.off()


# tamarind----------------------
tam <- sw_spat_rescale[sw_spat_rescale$Species_name == "Tamarindus indica" & sw_spat_rescale$is.KL2 == 1, ]

samplesize <- 0.70 * nrow(tam)
set.seed(100)
index <- sample(seq_len(nrow(tam)), size = samplesize)
# Creating training and test set
tamtrain <- tam[index, ]
tamtest <- tam[-index, ]

tamtrain$User_Tree_id <- as.factor(tamtrain$User_Tree_id)

fl_inter_tam_po <- clmm2(
  fl_intensity_fac ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs,
  random = User_Tree_id,
  data = tamtrain, Hess = T
)

summary(fl_inter_tam_po)
saveRDS(fl_inter_tam_po, file = "models/fl_inter_tam_po.rds")

png("plots/tam_SM_aspect_effect.png", width = 8, height = 8, res = 300, units = "in")
plot(Effect(focal.predictors = c("FNMean_MeanSM_cs", "aspect_cs"), fl_inter_tam_po), rug = T)
dev.off()

png("plots/tam_DryDays_light_effect.png", width = 8, height = 8, res = 300, units = "in")
plot(Effect(focal.predictors = c("FNSum_DryDays_cs", "light_cs"), fl_inter_tam_po), rug = T)
dev.off()

png("plots/tam_SumPCP_elev_effect.png", width = 8, height = 8, res = 300, units = "in")
plot(Effect(focal.predictors = c("FNSum_PCP_cs", "elev_cs"), fl_inter_tam_po), rug = T)
dev.off()



# plots----------------
# https://stackoverflow.com/questions/38977094/how-to-use-dotplot-after-clmm
# https://www.rdocumentation.org/packages/ordinal/versions/2011.05-10/topics/profile.clmm

all3_data <- list(jack, mango, tam)
saveRDS(all3_data, file = "models/all3_data.rds")


library(sjPlot)
tab_model(fl_inter_jack_po, fl_inter_mango_po, fl_inter_tam_po, file = "plots/fl_inter_jmt_po.html")
```


```{r fruiting}
# jack
fr_inter_jack_po <- clmm2(
  fr_intensity_fac ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs,
  random = User_Tree_id,
  data = jacktrain, Hess = T
)

summary(fr_inter_jack_po)
fr_inter_jack_po

pred_jack <- predict(fr_inter_jack_po, newdata = jacktest)
plot(jacktest$fr_intensity, pred_jack)

saveRDS(fr_inter_jack_po, file = "models/fr_inter_jack_po.rds")


png("plots/jack_SM_aspect_effect_fr.png", width = 8, height = 8, res = 300, units = "in")
plot(Effect(focal.predictors = c("FNMean_MeanSM_cs", "aspect_cs"), fr_inter_jack_po), rug = T)
dev.off()

png("plots/jack_DryDays_light_effect_fr.png", width = 8, height = 8, res = 300, units = "in")
plot(Effect(focal.predictors = c("FNSum_DryDays_cs", "light_cs"), fr_inter_jack_po), rug = T)
dev.off()

png("plots/jack_SumPCP_elev_effect_fr.png", width = 8, height = 8, res = 300, units = "in")
plot(Effect(focal.predictors = c("FNSum_PCP_cs", "elev_cs"), fr_inter_jack_po), rug = T)
dev.off()

# mango
fr_inter_mango_po <- clmm2(
  fr_intensity_fac ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs,
  random = User_Tree_id,
  data = mangotrain, Hess = T
)

summary(fr_inter_mango_po)

pred_mango <- predict(fr_inter_mango_po, newdata = mangotest)
plot(mangotest$fr_intensity, pred_mango)

saveRDS(fr_inter_mango_po, file = "models/fr_inter_mango_po.rds")


png("plots/mango_SM_aspect_effect_fr.png", width = 8, height = 8, res = 300, units = "in")
plot(Effect(focal.predictors = c("FNMean_MeanSM_cs", "aspect_cs"), fr_inter_mango_po), rug = T)
dev.off()

png("plots/mango_DryDays_light_effect_fr.png", width = 8, height = 8, res = 300, units = "in")
plot(Effect(focal.predictors = c("FNSum_DryDays_cs", "light_cs"), fr_inter_mango_po), rug = T)
dev.off()

png("plots/mango_SumPCP_elev_effect_fr.png", width = 8, height = 8, res = 300, units = "in")
plot(Effect(focal.predictors = c("FNSum_PCP_cs", "elev_cs"), fr_inter_mango_po), rug = T)
dev.off()


# tamarind
fr_inter_tam_po <- clmm2(
  fr_intensity_fac ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs,
  random = User_Tree_id,
  data = tamtrain, Hess = T
)

summary(fr_inter_tam_po)

pred_tam <- predict(fr_inter_tam_po, newdata = tamtest)
plot(tamtest$fr_intensity, pred_tam)

saveRDS(fr_inter_tam_po, file = "models/fr_inter_tam_po.rds")


png("plots/tam_SM_aspect_effect_fr.png", width = 8, height = 8, res = 300, units = "in")
plot(Effect(focal.predictors = c("FNMean_MeanSM_cs", "aspect_cs"), fr_inter_tam_po), rug = T)
dev.off()

png("plots/tam_DryDays_light_effect_fr.png", width = 8, height = 8, res = 300, units = "in")
plot(Effect(focal.predictors = c("FNSum_DryDays_cs", "light_cs"), fr_inter_tam_po), rug = T)
dev.off()

png("plots/tam_SumPCP_elev_effect_fr.png", width = 8, height = 8, res = 300, units = "in")
plot(Effect(focal.predictors = c("FNSum_PCP_cs", "elev_cs"), fr_inter_tam_po), rug = T)
dev.off()

tab_model(fr_inter_jack_po, fr_inter_mango_po, fr_inter_tam_po, file = "plots/fr_inter_jmt_po.html")
```

```{r ordinal model predictions - fruiting and flowering}
# figure out how to do this---------
# the code is giving classes now, but it is giving 1/3rd as many as it should

library(seqinr)

jacktest.nona <- jacktest[complete.cases(jacktest), ]
jacktest.nona <- jacktest.nona[order(jacktest.nona$Observation_ID), ]
jacktest3 <- rbind(jacktest.nona, jacktest.nona, jacktest.nona)
jacktest3 <- jacktest3[order(jacktest3$Observation_ID), ]
jacktest3$fl_intensity_fac <- factor(rep(c(0:2), nrow(jacktest.nona)))
pmat.clm <- matrix(predict(fl_inter_jack_po, newdata = jacktest3, allow.new.levels = T), ncol = 3, byrow = TRUE)

pred_jack <- as.numeric(unlist(apply(as.data.frame(pmat.clm), 1, which.max))) - 1

rjack <- round(cor(jacktest.nona$fl_intensity, pred_jack, use = "pairwise.complete.obs"), 2)


mangotest.nona <- mangotest[complete.cases(mangotest), ]
mangotest.nona <- mangotest.nona[order(mangotest.nona$Observation_ID), ]
mangotest3 <- rbind(mangotest.nona, mangotest.nona, mangotest.nona)
mangotest3 <- mangotest3[order(mangotest3$Observation_ID), ]
mangotest3$fl_intensity_fac <- factor(rep(c(0:2), nrow(mangotest.nona)))
pmat.clm <- matrix(predict(fl_inter_mango_po, newdata = mangotest3, allow.new.levels = T), ncol = 3, byrow = TRUE)

pred_mango <- as.numeric(unlist(apply(as.data.frame(pmat.clm), 1, which.max))) - 1

rmango <- round(cor(mangotest.nona$fl_intensity, pred_mango, use = "pairwise.complete.obs"), 2)

tamtest.nona <- tamtest[complete.cases(tamtest), ]
tamtest.nona <- tamtest.nona[order(tamtest.nona$Observation_ID), ]
tamtest3 <- rbind(tamtest.nona, tamtest.nona, tamtest.nona)
tamtest3 <- tamtest3[order(tamtest3$Observation_ID), ]
tamtest3$fl_intensity_fac <- factor(rep(c(0:2), nrow(tamtest.nona)))
pmat.clm <- matrix(predict(fl_inter_tam_po, newdata = tamtest3, allow.new.levels = T), ncol = 3, byrow = TRUE)

pred_tam <- as.numeric(unlist(apply(as.data.frame(pmat.clm), 1, which.max))) - 1

rtam <- round(cor(tamtest.nona$fl_intensity, pred_tam, use = "pairwise.complete.obs"), 2)

png("plots/jmt_pred_fl_po.png", width = 12, height = 4, res = 300, units = "in")
par(mfrow = c(1, 3))
plot(jitter(jacktest.nona$fl_intensity), jitter(pred_jack),
  col = col2alpha(rgb(0, 0, 0), alpha = 0.1), pch = 16,
  main = paste("jack\nr=", rjack, sep = ""), xlab = "test data", ylab = "predicted"
)
plot(jitter(mangotest.nona$fl_intensity), jitter(pred_mango),
  col = col2alpha(rgb(0, 0, 0), alpha = 0.1), pch = 16,
  main = paste("mango\nr=", rmango, sep = ""), xlab = "test data", ylab = "predicted"
)
plot(jitter(tamtest.nona$fl_intensity), jitter(pred_tam),
  col = col2alpha(rgb(0, 0, 0), alpha = 0.1), pch = 16,
  main = paste("tamarind\nr=", rtam, sep = ""), xlab = "test data", ylab = "predicted"
)
dev.off()

# fruiting

# jack
jacktest3$fr_intensity_fac <- factor(rep(c(0:2), nrow(jacktest.nona)))
pmat.clm <- matrix(predict(fr_inter_jack_po, newdata = jacktest3, allow.new.levels = T), ncol = 3, byrow = TRUE)
pred_jack <- as.numeric(unlist(apply(as.data.frame(pmat.clm), 1, which.max))) - 1
rjack <- round(cor(jacktest.nona$fr_intensity, pred_jack, use = "pairwise.complete.obs"), 2)

# mango
mangotest3$fr_intensity_fac <- factor(rep(c(0:2), nrow(mangotest.nona)))
pmat.clm <- matrix(predict(fr_inter_mango_po, newdata = mangotest3, allow.new.levels = T), ncol = 3, byrow = TRUE)
pred_mango <- as.numeric(unlist(apply(as.data.frame(pmat.clm), 1, which.max))) - 1
rmango <- round(cor(mangotest.nona$fr_intensity, pred_mango, use = "pairwise.complete.obs"), 2)

# tamarind
tamtest3$fr_intensity_fac <- factor(rep(c(0:2), nrow(tamtest.nona)))
pmat.clm <- matrix(predict(fr_inter_tam_po, newdata = tamtest3, allow.new.levels = T), ncol = 3, byrow = TRUE)
pred_tam <- as.numeric(unlist(apply(as.data.frame(pmat.clm), 1, which.max))) - 1
rtam <- round(cor(tamtest.nona$fr_intensity, pred_tam, use = "pairwise.complete.obs"), 2)

png("plots/jmt_pred_fr_po.png", width = 12, height = 4, res = 300, units = "in")
par(mfrow = c(1, 3))
plot(jitter(jacktest.nona$fr_intensity), jitter(pred_jack),
  col = col2alpha(rgb(0, 0, 0), alpha = 0.1), pch = 16,
  main = paste("jack\nr=", rjack, sep = ""), xlab = "test data", ylab = "predicted"
)
plot(jitter(mangotest.nona$fr_intensity), jitter(pred_mango),
  col = col2alpha(rgb(0, 0, 0), alpha = 0.1), pch = 16,
  main = paste("mango\nr=", rmango, sep = ""), xlab = "test data", ylab = "predicted"
)
plot(jitter(tamtest.nona$fr_intensity), jitter(pred_tam),
  col = col2alpha(rgb(0, 0, 0), alpha = 0.1), pch = 16,
  main = paste("tamarind\nr=", rtam, sep = ""), xlab = "test data", ylab = "predicted"
)
dev.off()
```

```{r jack ordinal LOOCV}
yr <- unique(jack$year)
jack.ord.ests_fl <- data.frame(matrix(0, nrow = 12, ncol = 9))
jack.ord.std.er_fl <- data.frame(matrix(0, nrow = 12, ncol = 9))
jack.ord.ests_fr <- data.frame(matrix(0, nrow = 12, ncol = 9))
jack.ord.std.er_fr <- data.frame(matrix(0, nrow = 12, ncol = 9))
jack.ord.mse.fl <- data.frame(yr, mse = rep(0, 9), cor = rep(0, 9), p_val = rep(0, 9))
jack.ord.mse.fr <- data.frame(yr, mse = rep(0, 9), cor = rep(0, 9), p_val = rep(0, 9))

library(ordinal)
i <- 1
for (i in 1:length(yr)) {
  dat <- jack[which(jack$year != yr[i]), ]
  test <- jack[which(jack$year == yr[i]), ]
  dat$User_Tree_id <- as.factor(dat$User_Tree_id)
  fl_inter_jack_po_loocv <- clmm2(
    fl_intensity_fac ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
      FNMean_MeanSM_cs + FNSum_DryDays_cs +
      elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
      light_cs * FNSum_DryDays_cs,
    random = User_Tree_id,
    data = dat, Hess = T, control = clmm2.control("nlminb")
  )

  saveRDS(fl_inter_jack_po_loocv, paste0("models/loocv/jack.ord.ests_fl_", yr[i], ".rds"))


  jack.ord.ests_fl[, i] <- coef(summary(fl_inter_jack_po_loocv))[, 1]
  jack.ord.std.er_fl[, i] <- coef(summary(fl_inter_jack_po_loocv))[, 2]

  test.nona <- test[complete.cases(test), ]
  test.nona <- test.nona[order(test.nona$Observation_ID), ]
  test3 <- rbind(test.nona, test.nona, test.nona)
  test3 <- test3[order(test3$Observation_ID), ]
  test3$fl_intensity_fac <- factor(rep(c(0:2), nrow(test.nona)))
  pmat.clm <- matrix(predict(fl_inter_jack_po_loocv, newdata = test3, allow.new.levels = T), ncol = 3, byrow = TRUE)

  pred_fl <- as.numeric(unlist(apply(as.data.frame(pmat.clm), 1, which.max))) - 1

  rjack <- round(cor(test.nona$fl_intensity, pred_fl, use = "pairwise.complete.obs"), 2)


  jack.ord.mse.fl[i, 2] <- mean((test.nona$fl_intensity - pred_fl)^2, na.rm = T)
  jack.ord.mse.fl[i, 3] <- cor.test(test.nona$fl_intensity, pred_fl, use = "complete.obs")[[4]] # correlation value
  jack.ord.mse.fl[i, 4] <- cor.test(test.nona$fl_intensity, pred_fl, use = "complete.obs")[[3]] # p-value


  write.csv(jack.ord.ests_fl, "models/loocv/jack.ord.ests_fl.csv")
  write.csv(jack.ord.std.er_fl, "models/loocv/jack.ord.std.er_fl.csv")
  write.csv(jack.ord.mse.fl, "models/loocv/jack.ord.mse.fl.csv")

  fr_inter_jack_po_loocv <- clmm2(
    fr_intensity_fac ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
      FNMean_MeanSM_cs + FNSum_DryDays_cs +
      elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
      light_cs * FNSum_DryDays_cs,
    random = User_Tree_id,
    data = dat, Hess = T, control = clmm2.control("nlminb")
  )
  saveRDS(fr_inter_jack_po_loocv, paste0("models/loocv/jack.ord.ests_fr_", yr[i], ".rds"))

  jack.ord.ests_fr[, i] <- coef(summary(fr_inter_jack_po_loocv))[, 1]
  jack.ord.std.er_fr[, i] <- coef(summary(fr_inter_jack_po_loocv))[, 2]

  test3$fr_intensity_fac <- factor(rep(c(0:2), nrow(test.nona)))
  pmat.clm <- matrix(predict(fr_inter_jack_po_loocv, newdata = test3, allow.new.levels = T), ncol = 3, byrow = TRUE)

  pred_fr <- as.numeric(unlist(apply(as.data.frame(pmat.clm), 1, which.max))) - 1

  jack.ord.mse.fr[i, 2] <- mean((test.nona$fr_intensity - pred_fr)^2, na.rm = T)
  jack.ord.mse.fr[i, 3] <- cor(test.nona$fr_intensity, pred_fr, use = "complete.obs")[[4]] # correlation
  jack.ord.mse.fr[i, 4] <- cor.test(test.nona$fr_intensity, pred_fr, use = "complete.obs")[[3]] # p-value

  write.csv(jack.ord.ests_fr, "models/loocv/jack.ord.ests_fr.csv")
  write.csv(jack.ord.std.er_fr, "models/loocv/jack.ord.std.er_fr.csv")
  write.csv(jack.ord.mse.fr, "models/loocv/jack.ord.mse.fr.csv")
}

colnames(jack.ord.ests_fl) <- colnames(jack.ord.ests_fr) <- colnames(jack.ord.std.er_fl) <- colnames(jack.ord.std.er_fr) <- yr
row.names(jack.ord.ests_fl) <- row.names(jack.ord.ests_fr) <- row.names(jack.ord.std.er_fl) <- row.names(jack.ord.std.er_fr) <- row.names(coef(summary(fr_inter_jack_po_loocv)))

write.csv(jack.ord.ests_fl, "models/jack.ord.ests_fl.csv")
write.csv(jack.ord.std.er_fl, "models/jack.ord.std.er_fl.csv")
write.csv(jack.ord.mse.fl, "models/jack.ord.mse.fl.csv")

write.csv(jack.ord.ests_fr, "models/jack.ord.ests_fr.csv")
write.csv(jack.ord.std.er_fr, "models/jack.ord.std.er_fr.csv")
write.csv(jack.ord.mse.fr, "models/jack.ord.mse.fr.csv")
```

```{r mango ordinal LOOCV}
yr <- unique(mango$year)
mango.ord.ests_fl <- data.frame(matrix(0, nrow = 12, ncol = 9))
mango.ord.std.er_fl <- data.frame(matrix(0, nrow = 12, ncol = 9))
mango.ord.ests_fr <- data.frame(matrix(0, nrow = 12, ncol = 9))
mango.ord.std.er_fr <- data.frame(matrix(0, nrow = 12, ncol = 9))
mango.ord.mse.fl <- data.frame(yr, mse = rep(0, 9), cor = rep(0, 9))
mango.ord.mse.fr <- data.frame(yr, mse = rep(0, 9), cor = rep(0, 9))

library(ordinal)
i <- 9
for (i in 1:length(yr)) {
  dat <- mango[which(mango$year != yr[i]), ]
  test <- mango[which(mango$year == yr[i]), ]
  dat$User_Tree_id <- as.factor(dat$User_Tree_id)
  fl_inter_mango_po_loocv <- clmm2(
    fl_intensity_fac ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
      FNMean_MeanSM_cs + FNSum_DryDays_cs +
      elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
      light_cs * FNSum_DryDays_cs,
    random = User_Tree_id,
    data = dat, Hess = T, control = clmm2.control("nlminb")
  )



  mango.ord.ests_fl[, i] <- coef(summary(fl_inter_mango_po_loocv))[, 1]
  mango.ord.std.er_fl[, i] <- coef(summary(fl_inter_mango_po_loocv))[, 2]

  test.nona <- test[complete.cases(test), ]
  test.nona <- test.nona[order(test.nona$Observation_ID), ]
  test3 <- rbind(test.nona, test.nona, test.nona)
  test3 <- test3[order(test3$Observation_ID), ]
  test3$fl_intensity_fac <- factor(rep(c(0:2), nrow(test.nona)))
  pmat.clm <- matrix(predict(fl_inter_mango_po_loocv, newdata = test3, allow.new.levels = T), ncol = 3, byrow = TRUE)

  pred_fl <- as.numeric(unlist(apply(as.data.frame(pmat.clm), 1, which.max))) - 1

  rmango <- round(cor(test.nona$fl_intensity, pred_fl, use = "pairwise.complete.obs"), 2)


  mango.ord.mse.fl[i, 2] <- mean((test.nona$fl_intensity - pred_fl)^2, na.rm = T)
  mango.ord.mse.fl[i, 3] <- cor(test.nona$fl_intensity, pred_fl, use = "complete.obs")

  write.csv(mango.ord.ests_fl, "models/mango.ord.ests_fl.csv")
  write.csv(mango.ord.std.er_fl, "models/mango.ord.std.er_fl.csv")
  write.csv(mango.ord.mse.fl, "models/mango.ord.mse.fl.csv")

  fr_inter_mango_po_loocv <- clmm2(
    fr_intensity_fac ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
      FNMean_MeanSM_cs + FNSum_DryDays_cs +
      elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
      light_cs * FNSum_DryDays_cs,
    random = User_Tree_id,
    data = dat, Hess = T, control = clmm2.control("nlminb")
  )

  mango.ord.ests_fr[, i] <- coef(summary(fr_inter_mango_po_loocv))[, 1]
  mango.ord.std.er_fr[, i] <- coef(summary(fr_inter_mango_po_loocv))[, 2]

  test3$fr_intensity_fac <- factor(rep(c(0:2), nrow(test.nona)))
  pmat.clm <- matrix(predict(fr_inter_mango_po_loocv, newdata = test3, allow.new.levels = T), ncol = 3, byrow = TRUE)

  pred_fr <- as.numeric(unlist(apply(as.data.frame(pmat.clm), 1, which.max))) - 1

  mango.ord.mse.fr[i, 2] <- mean((test.nona$fr_intensity - pred_fr)^2, na.rm = T)
  mango.ord.mse.fr[i, 3] <- cor(test.nona$fr_intensity, pred_fr, use = "complete.obs")

  write.csv(mango.ord.ests_fr, "models/mango.ord.ests_fr.csv")
  write.csv(mango.ord.std.er_fr, "models/mango.ord.std.er_fr.csv")
  write.csv(mango.ord.mse.fr, "models/mango.ord.mse.fr.csv")
}

colnames(mango.ord.ests_fl) <- colnames(mango.ord.ests_fr) <- colnames(mango.ord.std.er_fl) <- colnames(mango.ord.std.er_fr) <- yr
row.names(mango.ord.ests_fl) <- row.names(mango.ord.ests_fr) <- row.names(mango.ord.std.er_fl) <- row.names(mango.ord.std.er_fr) <- row.names(coef(summary(fr_inter_mango_po_loocv)))

write.csv(mango.ord.ests_fl, "models/mango.ord.ests_fl.csv")
write.csv(mango.ord.std.er_fl, "models/mango.ord.std.er_fl.csv")
write.csv(mango.ord.mse.fl, "models/mango.ord.mse.fl.csv")

write.csv(mango.ord.ests_fr, "models/mango.ord.ests_fr.csv")
write.csv(mango.ord.std.er_fr, "models/mango.ord.std.er_fr.csv")
write.csv(mango.ord.mse.fr, "models/mango.ord.mse.fr.csv")
```

```{r tamarind ordinal LOOCV}
yr <- unique(tam$year)
tam.ord.ests_fl <- data.frame(matrix(0, nrow = 12, ncol = 9))
tam.ord.std.er_fl <- data.frame(matrix(0, nrow = 12, ncol = 9))
tam.ord.ests_fr <- data.frame(matrix(0, nrow = 12, ncol = 9))
tam.ord.std.er_fr <- data.frame(matrix(0, nrow = 12, ncol = 9))
tam.ord.mse.fl <- data.frame(yr, mse = rep(0, 9), cor = rep(0, 9))
tam.ord.mse.fr <- data.frame(yr, mse = rep(0, 9), cor = rep(0, 9))

library(ordinal)
i <- 1
for (i in 1:length(yr)) {
  dat <- tam[which(tam$year != yr[i]), ]
  test <- tam[which(tam$year == yr[i]), ]
  dat$User_Tree_id <- as.factor(dat$User_Tree_id)
  fl_inter_tam_po_loocv <- clmm2(
    fl_intensity_fac ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
      FNMean_MeanSM_cs + FNSum_DryDays_cs +
      elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
      light_cs * FNSum_DryDays_cs,
    random = User_Tree_id,
    data = dat, Hess = T, control = clmm2.control("nlminb")
  )



  tam.ord.ests_fl[, i] <- coef(summary(fl_inter_tam_po_loocv))[, 1]
  tam.ord.std.er_fl[, i] <- coef(summary(fl_inter_tam_po_loocv))[, 2]

  test.nona <- test[complete.cases(test), ]
  test.nona <- test.nona[order(test.nona$Observation_ID), ]
  test3 <- rbind(test.nona, test.nona, test.nona)
  test3 <- test3[order(test3$Observation_ID), ]
  test3$fl_intensity_fac <- factor(rep(c(0:2), nrow(test.nona)))
  pmat.clm <- matrix(predict(fl_inter_tam_po_loocv, newdata = test3, allow.new.levels = T), ncol = 3, byrow = TRUE)

  pred_fl <- as.numeric(unlist(apply(as.data.frame(pmat.clm), 1, which.max))) - 1

  rtam <- round(cor(test.nona$fl_intensity, pred_fl, use = "pairwise.complete.obs"), 2)



  tam.ord.mse.fl[i, 2] <- mean((test.nona$fl_intensity - pred_fl)^2, na.rm = T)
  tam.mse.fl[i, 3] <- cor(test.nona$fl_intensity, pred_fl, use = "complete.obs")

  fr_inter_tam_po_loocv <- clmm2(
    fr_intensity_fac ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
      FNMean_MeanSM_cs + FNSum_DryDays_cs +
      elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
      light_cs * FNSum_DryDays_cs,
    random = User_Tree_id,
    data = dat, Hess = T, control = clmm2.control("nlminb")
  )

  tam.bin.ests_fr[, i] <- coef(summary(fr_inter_tam_bin_loocv))[, 1]
  tam.bin.std.er_fr[, i] <- coef(summary(fr_inter_tam_bin_loocv))[, 2]

  test3$fr_intensity_fac <- factor(rep(c(0:2), nrow(test.nona)))
  pmat.clm <- matrix(predict(fr_inter_tam_po_loocv, newdata = test3, allow.new.levels = T), ncol = 3, byrow = TRUE)

  pred_fr <- as.numeric(unlist(apply(as.data.frame(pmat.clm), 1, which.max))) - 1

  tam.ord.mse.fr[i, 2] <- mean((test.nona$fr_intensity - pred_fr)^2, na.rm = T)
  tam.mse.fr[i, 3] <- cor(test.nona$fr_intensity, pred_fr, use = "complete.obs")
}

colnames(tam.ord.ests_fl) <- colnames(tam.ord.ests_fr) <- colnames(tam.ord.std.er_fl) <- colnames(tam.ord.std.er_fr) <- yr
row.names(tam.ord.ests_fl) <- row.names(tam.ord.ests_fr) <- row.names(tam.ord.std.er_fl) <- row.names(tam.ord.std.er_fr) <- row.names(coef(summary(fr_inter_tam_po_loocv)))

write.csv(tam.ord.ests_fl, "models/tam.ord.ests_fl.csv")
write.csv(tam.ord.ests_fr, "models/tam.ord.ests_fr.csv")
write.csv(tam.ord.std.er_fl, "models/tam.ord.std.er_fl.csv")
write.csv(tam.ord.std.er_fr, "models/tam.ord.std.er_fr.csv")
write.csv(tam.ord.mse.fl, "models/tam.ord.mse.fl.csv")
write.csv(tam.ord.mse.fr, "models/tam.ord.mse.fr.csv")
```

```{r only tam fruiting}
i <- 2
for (i in 1:length(yr)) {
  dat <- tam[which(tam$year != yr[i]), ]
  test <- tam[which(tam$year == yr[i]), ]
  dat$User_Tree_id <- as.factor(dat$User_Tree_id)
  test.nona <- test[complete.cases(test), ]
  test.nona <- test.nona[order(test.nona$Observation_ID), ]
  test3 <- rbind(test.nona, test.nona, test.nona)
  test3 <- test3[order(test3$Observation_ID), ]

  fr_inter_tam_po_loocv <- clmm2(
    fr_intensity_fac ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
      FNMean_MeanSM_cs + FNSum_DryDays_cs +
      elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
      light_cs * FNSum_DryDays_cs,
    random = User_Tree_id,
    data = dat, Hess = T, control = clmm2.control("nlminb")
  )

  tam.ord.ests_fr[, i] <- coef(summary(fr_inter_tam_po_loocv))[, 1]
  tam.ord.std.er_fr[, i] <- coef(summary(fr_inter_tam_po_loocv))[, 2]

  test3$fr_intensity_fac <- factor(rep(c(0:2), nrow(test.nona)))
  pmat.clm <- matrix(predict(fr_inter_tam_po_loocv, newdata = test3, allow.new.levels = T), ncol = 3, byrow = TRUE)

  pred_fr <- as.numeric(unlist(apply(as.data.frame(pmat.clm), 1, which.max))) - 1

  tam.ord.mse.fr[i, 2] <- mean((test.nona$fr_intensity - pred_fr)^2, na.rm = T)
  tam.ord.mse.fr[i, 3] <- cor(test.nona$fr_intensity, pred_fr, use = "complete.obs")
}

colnames(tam.ord.ests_fr) <- colnames(tam.ord.std.er_fr) <- yr
row.names(tam.ord.ests_fr) <- row.names(tam.ord.std.er_fr) <- row.names(coef(summary(fr_inter_tam_po_loocv)))

write.csv(tam.ord.ests_fr, "models/tam.ord.ests_fr.csv")
write.csv(tam.ord.std.er_fr, "models/tam.ord.std.er_fr.csv")
write.csv(tam.ord.mse.fr, "models/tam.ord.mse.fr.csv")
```

```{r [SUPP] fruiting with flowering as predictor}
library(dplyr)
jack$User_Tree_id_ch <- as.character(jack$User_Tree_id)
jack_fl_tmin1 <- mutate(jack, group_by(User_Tree_id_ch),
  fl_tmin1 = lag(fl_intensity, n = 1, default = NA)
)

jack$week <- week(jack$Date)
library(lubridate)
jack_fl_tmin1 <-
  jack %>%
  group_by(User_Tree_id_ch) %>%
  mutate(
    fl_tmin1 = dplyr::lag(fl_intensity, n = 1, default = NA),
    diff = week - lag(week, n = 1, default = NA)
  )

jack$fort <- ceiling(week(jack$Date) / 2)
jack_fl_tmin1_2 <- aggregate(fl_intensity ~ year + fort + User_Tree_id, jack, mean)
```

```{r binary models flowering}
library(lme4)
# jack
fl_inter_jack_bin <- glmer(
  fl_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs + (1 | User_Tree_id),
  family = binomial, data = jacktrain
)

summary(fl_inter_jack_bin)
fl_inter_jack_bin

colnames(jacktest)
jacktest2 <- jacktest[, c(5, 8, 12, 42:49)]



saveRDS(fl_inter_jack_bin, file = "models/fl_inter_jack_bin.rds")

library(sjPlot)
library(patchwork)

p1 <- plot_model(fl_inter_jack_bin, type = "est", show.values = T, show.p = T)
# plot_model(fl_inter_jack_bin, type="eff", transform, show.data = T)
p2 <- plot_model(fl_inter_jack_bin, type = "eff", terms = c("FNSum_PCP_cs", "elev_cs"), transform, show.data = T)
p3 <- plot_model(fl_inter_jack_bin, type = "eff", terms = c("FNMean_MeanSM_cs", "aspect_cs"), transform, show.data = T)
p4 <- plot_model(fl_inter_jack_bin, type = "eff", terms = c("FNSum_DryDays_cs", "light_cs"), transform, show.data = T)

png("plots/jack_bin_est_eff.png", width = 12, height = 8, res = 300, units = "in")
p1 + p2 + p3 + p4
dev.off()

# mango
fl_inter_mango_bin <- glmer(
  fl_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs + (1 | User_Tree_id),
  family = binomial, data = mangotrain
)

summary(fl_inter_mango_bin)

pred_mango <- predict(fl_inter_mango_bin, newdata = mangotest)
plot(mangotest$fl_binary, pred_mango)

saveRDS(fl_inter_mango_bin, file = "models/fl_inter_mango_bin.rds")


p1 <- plot_model(fl_inter_mango_bin, type = "est", show.values = T, show.p = T)
# plot_model(fl_inter_jack_bin, type="eff", transform, show.data = T)
p2 <- plot_model(fl_inter_mango_bin, type = "eff", terms = c("FNSum_PCP_cs", "elev_cs"), transform, show.data = T)
p3 <- plot_model(fl_inter_mango_bin, type = "eff", terms = c("FNMean_MeanSM_cs", "aspect_cs"), transform, show.data = T)
p4 <- plot_model(fl_inter_mango_bin, type = "eff", terms = c("FNSum_DryDays_cs", "light_cs"), transform, show.data = T)

png("plots/mango_bin_est_eff.png", width = 12, height = 8, res = 300, units = "in")
p1 + p2 + p3 + p4
dev.off()

# tamarind
fl_inter_tam_bin <- glmer(
  fl_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs + (1 | User_Tree_id),
  family = binomial, data = tamtrain,
  glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

summary(fl_inter_tam_bin)

pred_tam <- predict(fl_inter_tam_bin, newdata = tamtest)
plot(tamtest$fl_binary, pred_tam)

saveRDS(fl_inter_tam_bin, file = "models/fl_inter_tam_bin.rds")


p1 <- plot_model(fl_inter_tam_bin, type = "est", show.values = T, show.p = T)
# plot_model(fl_inter_jack_bin, type="eff", transform, show.data = T)
p2 <- plot_model(fl_inter_tam_bin, type = "eff", terms = c("FNSum_PCP_cs", "elev_cs"), transform, show.data = T)
p3 <- plot_model(fl_inter_tam_bin, type = "eff", terms = c("FNMean_MeanSM_cs", "aspect_cs"), transform, show.data = T)
p4 <- plot_model(fl_inter_tam_bin, type = "eff", terms = c("FNSum_DryDays_cs", "light_cs"), transform, show.data = T)

png("plots/tam_bin_est_eff.png", width = 12, height = 8, res = 300, units = "in")
p1 + p2 + p3 + p4
dev.off()

tab_model(fl_inter_jack_bin, fl_inter_mango_bin, fl_inter_tam_bin, file = "plots/fl_inter_jmt_bin.html")
```

```{r LOOCV for jackfruit}
library(lme4)
yr <- unique(jack$year)
jack.bin.ests_fl <- data.frame(matrix(0, nrow = 11, ncol = 9))
jack.bin.std.er_fl <- data.frame(matrix(0, nrow = 11, ncol = 9))
jack.bin.ests_fr <- data.frame(matrix(0, nrow = 11, ncol = 9))
jack.bin.std.er_fr <- data.frame(matrix(0, nrow = 11, ncol = 9))
jack.mse.fl <- data.frame(yr, mse = rep(0, 9), cor = rep(0, 9))
jack.mse.fr <- data.frame(yr, mse = rep(0, 9), cor = rep(0, 9))


for (i in 1:length(yr)) {
  dat <- jack[which(jack$year != yr[i]), ]
  test <- jack[which(jack$year == yr[i]), ]
  fl_inter_jack_bin_loocv <- glmer(
    fl_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
      FNMean_MeanSM_cs + FNSum_DryDays_cs +
      elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
      light_cs * FNSum_DryDays_cs + (1 | User_Tree_id),
    family = binomial, data = dat,
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
  )

  saveRDS(fl_inter_jack_bin_loocv, file = paste0("models/loocv/fl_inter_jack_bin_loocv_", i, ".rds"))

  # jack.bin.ests_fl[,i]<-coef(summary(fl_inter_jack_bin_loocv))[,1]
  # jack.bin.std.er_fl[,i]<-coef(summary(fl_inter_jack_bin_loocv))[,2]
  # pred.fl<-plogis(predict(fl_inter_jack_bin_loocv, newdata=test, allow.new.levels=T))
  # jack.mse.fl[i,2]<-mean((test$fl_binary-pred.fl)^2, na.rm=T)
  # jack.mse.fl[i,3]<-cor(test$fl_binary, pred.fl, use="complete.obs")

  fr_inter_jack_bin_loocv <- glmer(
    fr_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
      FNMean_MeanSM_cs + FNSum_DryDays_cs +
      elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
      light_cs * FNSum_DryDays_cs + (1 | User_Tree_id),
    family = binomial, data = dat,
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
  )

  saveRDS(fr_inter_jack_bin_loocv, file = paste0("models/loocv/fr_inter_jack_bin_loocv_", i, ".rds"))


  # jack.bin.ests_fr[,i]<-coef(summary(fr_inter_jack_bin_loocv))[,1]
  # jack.bin.std.er_fr[,i]<-coef(summary(fr_inter_jack_bin_loocv))[,2]
  # pred.fr<-plogis(predict(fr_inter_jack_bin_loocv, newdata=test, allow.new.levels=T))
  # jack.mse.fr[i,2]<-mean((test$fr_binary-pred.fr)^2, na.rm=T)
  # jack.mse.fr[i,3]<-cor(test$fr_binary, pred.fr, use="complete.obs")
}

colnames(jack.bin.ests_fl) <- colnames(jack.bin.ests_fr) <- colnames(jack.bin.std.er_fl) <- colnames(jack.bin.std.er_fr) <- yr
row.names(jack.bin.ests_fl) <- row.names(jack.bin.ests_fr) <- row.names(jack.bin.std.er_fl) <- row.names(jack.bin.std.er_fr) <- row.names(coef(summary(fr_inter_jack_bin_loocv)))

write.csv(jack.bin.ests_fl, "models/jack.bin.ests_fl.csv")
write.csv(jack.bin.ests_fr, "models/jack.bin.ests_fr.csv")
write.csv(jack.bin.std.er_fl, "models/jack.bin.std.er_fl.csv")
write.csv(jack.bin.std.er_fr, "models/jack.bin.std.er_fr.csv")
write.csv(jack.mse.fl, "models/jack.mse.fl.csv")
write.csv(jack.mse.fr, "models/jack.mse.fr.csv")
```


```{r LOOCV for mango}
yr <- unique(mango$year)
mango.bin.ests_fl <- data.frame(matrix(0, nrow = 11, ncol = 9))
mango.bin.std.er_fl <- data.frame(matrix(0, nrow = 11, ncol = 9))
mango.bin.ests_fr <- data.frame(matrix(0, nrow = 11, ncol = 9))
mango.bin.std.er_fr <- data.frame(matrix(0, nrow = 11, ncol = 9))
mango.mse.fl <- data.frame(yr, mse = rep(0, 9), cor = rep(0, 9))
mango.mse.fr <- data.frame(yr, mse = rep(0, 9), cor = rep(0, 9))

i <- 1
for (i in 1:length(yr)) {
  dat <- mango[which(mango$year != yr[i]), ]
  test <- mango[which(mango$year == yr[i]), ]
  fl_inter_mango_bin_loocv <- glmer(
    fl_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
      FNMean_MeanSM_cs + FNSum_DryDays_cs +
      elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
      light_cs * FNSum_DryDays_cs + (1 | User_Tree_id),
    family = binomial, data = dat, glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
  )
  saveRDS(fl_inter_mango_bin_loocv, file = paste0("models/loocv/fl_inter_mango_bin_loocv_", i, ".rds"))

  # mango.bin.ests_fl[,i]<-coef(summary(fl_inter_mango_bin_loocv))[,1]
  # mango.bin.std.er_fl[,i]<-coef(summary(fl_inter_mango_bin_loocv))[,2]
  # pred.fl<-plogis(predict(fl_inter_mango_bin_loocv, newdata=test, allow.new.levels=T))
  # mango.mse.fl[i,2]<-mean((test$fl_binary-pred.fl)^2, na.rm=T)
  # mango.mse.fl[i,3]<-cor(test$fl_binary, pred.fl, use="complete.obs")


  fr_inter_mango_bin_loocv <- glmer(
    fr_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
      FNMean_MeanSM_cs + FNSum_DryDays_cs +
      elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
      light_cs * FNSum_DryDays_cs + (1 | User_Tree_id),
    family = binomial, data = dat,
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
  )
  saveRDS(fr_inter_mango_bin_loocv, file = paste0("models/loocv/fr_inter_mango_bin_loocv_", i, ".rds"))

  # mango.bin.ests_fr[,i]<-coef(summary(fr_inter_mango_bin_loocv))[,1]
  # mango.bin.std.er_fr[,i]<-coef(summary(fr_inter_mango_bin_loocv))[,2]
  # pred.fr<-plogis(predict(fr_inter_mango_bin_loocv, newdata=test, allow.new.levels=T))
  # mango.mse.fr[i,2]<-mean((test$fr_binary-pred.fr)^2, na.rm=T)
  # mango.mse.fr[i,3]<-cor(test$fr_binary, pred.fr, use="complete.obs")
}

colnames(mango.bin.ests_fl) <- colnames(mango.bin.ests_fr) <- colnames(mango.bin.std.er_fl) <- colnames(mango.bin.std.er_fr) <- yr
row.names(mango.bin.ests_fl) <- row.names(mango.bin.ests_fr) <- row.names(mango.bin.std.er_fl) <- row.names(mango.bin.std.er_fr) <- row.names(coef(summary(fr_inter_mango_bin_loocv)))

write.csv(mango.bin.ests_fl, "models/mango.bin.ests_fl.csv")
write.csv(mango.bin.ests_fr, "models/mango.bin.ests_fr.csv")
write.csv(mango.bin.std.er_fl, "models/mango.bin.std.er_fl.csv")
write.csv(mango.bin.std.er_fr, "models/mango.bin.std.er_fr.csv")
write.csv(mango.mse.fl, "models/mango.mse.fl.csv")
write.csv(mango.mse.fr, "models/mango.mse.fr.csv")
```

```{r LOOCV for tam}
yr <- unique(tam$year)
tam.bin.ests_fl <- data.frame(matrix(0, nrow = 11, ncol = 9))
tam.bin.std.er_fl <- data.frame(matrix(0, nrow = 11, ncol = 9))
tam.bin.ests_fr <- data.frame(matrix(0, nrow = 11, ncol = 9))
tam.bin.std.er_fr <- data.frame(matrix(0, nrow = 11, ncol = 9))
tam.mse.fl <- data.frame(yr, mse = rep(0, 9), cor = rep(0, 9))
tam.mse.fr <- data.frame(yr, mse = rep(0, 9), cor = rep(0, 9))

i <- 1
for (i in 1:length(yr)) {
  dat <- tam[which(tam$year != yr[i]), ]
  test <- tam[which(tam$year == yr[i]), ]
  fl_inter_tam_bin_loocv <- glmer(
    fl_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
      FNMean_MeanSM_cs + FNSum_DryDays_cs +
      elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
      light_cs * FNSum_DryDays_cs + (1 | User_Tree_id),
    family = binomial, data = dat,
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
  )
  saveRDS(fl_inter_tam_bin_loocv, file = paste0("models/loocv/fl_inter_tam_bin_loocv_", i, ".rds"))

  # tam.bin.ests_fl[,i]<-coef(summary(fl_inter_tam_bin_loocv))[,1]
  # tam.bin.std.er_fl[,i]<-coef(summary(fl_inter_tam_bin_loocv))[,2]

  # pred.fl<-plogis(predict(fl_inter_tam_bin_loocv, newdata=test, allow.new.levels=T))
  # tam.mse.fl[i,2]<-mean((test$fl_binary-pred.fl)^2, na.rm=T)
  # tam.mse.fl[i,3]<-cor(test$fl_binary, pred.fl, use="complete.obs")

  fr_inter_tam_bin_loocv <- glmer(
    fr_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
      FNMean_MeanSM_cs + FNSum_DryDays_cs +
      elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
      light_cs * FNSum_DryDays_cs + (1 | User_Tree_id),
    family = binomial, data = dat,
    glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
  )
  saveRDS(fr_inter_tam_bin_loocv, file = paste0("models/loocv/fr_inter_tam_bin_loocv_", i, ".rds"))

  # tam.bin.ests_fr[,i]<-coef(summary(fr_inter_tam_bin_loocv))[,1]
  # tam.bin.std.er_fr[,i]<-coef(summary(fr_inter_tam_bin_loocv))[,2]

  # pred.fr<-plogis(predict(fr_inter_tam_bin_loocv, newdata=test, allow.new.levels=T))
  # tam.mse.fr[i,2]<-mean((test$fr_binary-pred.fr)^2, na.rm=T)
  # tam.mse.fr[i,3]<-cor(test$fr_binary, pred.fr, use="complete.obs")
}

colnames(tam.bin.ests_fl) <- colnames(tam.bin.ests_fr) <- colnames(tam.bin.std.er_fl) <- colnames(tam.bin.std.er_fr) <- yr
row.names(tam.bin.ests_fl) <- row.names(tam.bin.ests_fr) <- row.names(tam.bin.std.er_fl) <- row.names(tam.bin.std.er_fr) <- row.names(coef(summary(fr_inter_tam_bin_loocv)))

write.csv(tam.bin.ests_fl, "models/tam.bin.ests_fl.csv")
write.csv(tam.bin.ests_fr, "models/tam.bin.ests_fr.csv")
write.csv(tam.bin.std.er_fl, "models/tam.bin.std.er_fl.csv")
write.csv(tam.bin.std.er_fr, "models/tam.bin.std.er_fr.csv")
write.csv(tam.mse.fl, "models/tam.mse.fl.csv")
write.csv(tam.mse.fr, "models/tam.mse.fr.csv")
```

```{r cross validation from the models}
cv.fn <- function(glmfit, dat, ests) {
  preds <- sum(ests * dat)
}
```

```{r calculating f1 scores}
all3_data <- readRDS("models/all3_data.rds")

# calculate predicted values
library(lme4)

jack <- all3_data[[1]]
mango <- all3_data[[2]]
tam <- all3_data[[3]]

# jack flowering

fl_jack_name <- list.files("models/loocv/", "fl_inter_jack_bin_loocv_", full.names = T)
fl_jack <- mapply(function(x) readRDS(fl_jack_name[x]), x = 1:9)

pred.jack.fl <- unlist(mapply(function(x) plogis(predict(fl_jack[[x]], newdata = jack[which(jack$year == (x + 2013)), ], allow.new.levels = T)), x = 1:9))
pred.jack.fl2 <- ifelse(pred.jack.fl > 0.5, 1, 0)
obs.jack.fl <- unlist(mapply(function(x) jack$fl_binary[which(jack$year == x + 2013)], x = 1:9))

fl.acc <- rep(NA, length(obs.jack.fl))
fl.acc[which(pred.jack.fl2 == 1 & obs.jack.fl == 1)] <- "TP"
fl.acc[which(pred.jack.fl2 == 0 & obs.jack.fl == 0)] <- "TN"
fl.acc[which(pred.jack.fl2 == 0 & obs.jack.fl == 1)] <- "FN"
fl.acc[which(pred.jack.fl2 == 1 & obs.jack.fl == 0)] <- "FP"

# F1 score is 2*(precision*recall)/(precision+recall)
# precision = TP/(TP+FP)
# recall = TP/ (TP+FN)

fl.precision <- table(fl.acc)[4] / (table(fl.acc)[4] + table(fl.acc)[2])
fl.recall <- table(fl.acc)[4] / (table(fl.acc)[4] + table(fl.acc)[1])

fl.jack.f1 <- 2 * (fl.precision * fl.recall) / (fl.precision + fl.recall)

# jack fruiting
fr_jack_name <- list.files("models/loocv/", "fr_inter_jack_bin_loocv_", full.names = T)
fr_jack <- mapply(function(x) readRDS(fr_jack_name[x]), x = 1:9)

pred.jack.fr <- unlist(mapply(function(x) plogis(predict(fr_jack[[x]], newdata = jack[which(jack$year == (x + 2013)), ], allow.new.levels = T)), x = 1:9))
pred.jack.fr2 <- ifelse(pred.jack.fr > 0.5, 1, 0)
obs.jack.fr <- unlist(mapply(function(x) jack$fr_binary[which(jack$year == x + 2013)], x = 1:9))

fr.acc <- rep(NA, length(obs.jack.fr))
fr.acc[which(pred.jack.fr2 == 1 & obs.jack.fr == 1)] <- "TP"
fr.acc[which(pred.jack.fr2 == 0 & obs.jack.fr == 0)] <- "TN"
fr.acc[which(pred.jack.fr2 == 0 & obs.jack.fr == 1)] <- "FN"
fr.acc[which(pred.jack.fr2 == 1 & obs.jack.fr == 0)] <- "FP"

# F1 score is 2*(precision*recall)/(precision+recall)
# precision = TP/(TP+FP)
# recall = TP/ (TP+FN)

fr.precision <- table(fr.acc)[4] / (table(fr.acc)[4] + table(fr.acc)[2])
fr.recall <- table(fr.acc)[4] / (table(fr.acc)[4] + table(fr.acc)[1])

fr.jack.f1 <- 2 * (fr.precision * fr.recall) / (fr.precision + fr.recall)


# mango flowering

fl_mango_name <- list.files("models/loocv/", "fl_inter_mango_bin_loocv_", full.names = T)
fl_mango <- mapply(function(x) readRDS(fl_mango_name[x]), x = 1:9)

pred.mango.fl <- unlist(mapply(function(x) plogis(predict(fl_mango[[x]], newdata = mango[which(mango$year == (x + 2013)), ], allow.new.levels = T)), x = 1:9))
pred.mango.fl2 <- ifelse(pred.mango.fl > 0.5, 1, 0)
obs.mango.fl <- unlist(mapply(function(x) mango$fl_binary[which(mango$year == x + 2013)], x = 1:9))

fl.acc <- rep(NA, length(obs.mango.fl))
fl.acc[which(pred.mango.fl2 == 1 & obs.mango.fl == 1)] <- "TP"
fl.acc[which(pred.mango.fl2 == 0 & obs.mango.fl == 0)] <- "TN"
fl.acc[which(pred.mango.fl2 == 0 & obs.mango.fl == 1)] <- "FN"
fl.acc[which(pred.mango.fl2 == 1 & obs.mango.fl == 0)] <- "FP"

# F1 score is 2*(precision*recall)/(precision+recall)
# precision = TP/(TP+FP)
# recall = TP/ (TP+FN)

fl.precision <- table(fl.acc)[4] / (table(fl.acc)[4] + table(fl.acc)[2])
fl.recall <- table(fl.acc)[4] / (table(fl.acc)[4] + table(fl.acc)[1])

fl.mango.f1 <- 2 * (fl.precision * fl.recall) / (fl.precision + fl.recall)

# mango fruiting
fr_mango_name <- list.files("models/loocv/", "fr_inter_mango_bin_loocv_", full.names = T)
fr_mango <- mapply(function(x) readRDS(fr_mango_name[x]), x = 1:9)

pred.mango.fr <- unlist(mapply(function(x) plogis(predict(fr_mango[[x]], newdata = mango[which(mango$year == (x + 2013)), ], allow.new.levels = T)), x = 1:9))
pred.mango.fr2 <- ifelse(pred.mango.fr > 0.5, 1, 0)
obs.mango.fr <- unlist(mapply(function(x) mango$fr_binary[which(mango$year == x + 2013)], x = 1:9))

fr.acc <- rep(NA, length(obs.mango.fr))
fr.acc[which(pred.mango.fr2 == 1 & obs.mango.fr == 1)] <- "TP"
fr.acc[which(pred.mango.fr2 == 0 & obs.mango.fr == 0)] <- "TN"
fr.acc[which(pred.mango.fr2 == 0 & obs.mango.fr == 1)] <- "FN"
fr.acc[which(pred.mango.fr2 == 1 & obs.mango.fr == 0)] <- "FP"

# F1 score is 2*(precision*recall)/(precision+recall)
# precision = TP/(TP+FP)
# recall = TP/ (TP+FN)

fr.precision <- table(fr.acc)[4] / (table(fr.acc)[4] + table(fr.acc)[2])
fr.recall <- table(fr.acc)[4] / (table(fr.acc)[4] + table(fr.acc)[1])

fr.mango.f1 <- 2 * (fr.precision * fr.recall) / (fr.precision + fr.recall)

# tam flowering

fl_tam_name <- list.files("models/loocv/", "fl_inter_tam_bin_loocv_", full.names = T)
fl_tam <- mapply(function(x) readRDS(fl_tam_name[x]), x = 1:9)

pred.tam.fl <- unlist(mapply(function(x) plogis(predict(fl_tam[[x]], newdata = tam[which(tam$year == (x + 2013)), ], allow.new.levels = T)), x = 1:9))
pred.tam.fl2 <- ifelse(pred.tam.fl > 0.5, 1, 0)
obs.tam.fl <- unlist(mapply(function(x) tam$fl_binary[which(tam$year == x + 2013)], x = 1:9))

fl.acc <- rep(NA, length(obs.tam.fl))
fl.acc[which(pred.tam.fl2 == 1 & obs.tam.fl == 1)] <- "TP"
fl.acc[which(pred.tam.fl2 == 0 & obs.tam.fl == 0)] <- "TN"
fl.acc[which(pred.tam.fl2 == 0 & obs.tam.fl == 1)] <- "FN"
fl.acc[which(pred.tam.fl2 == 1 & obs.tam.fl == 0)] <- "FP"

# F1 score is 2*(precision*recall)/(precision+recall)
# precision = TP/(TP+FP)
# recall = TP/ (TP+FN)

fl.precision <- table(fl.acc)[4] / (table(fl.acc)[4] + table(fl.acc)[2])
fl.recall <- table(fl.acc)[4] / (table(fl.acc)[4] + table(fl.acc)[1])

fl.tam.f1 <- 2 * (fl.precision * fl.recall) / (fl.precision + fl.recall)

# tam fruiting
fr_tam_name <- list.files("models/loocv/", "fr_inter_tam_bin_loocv_", full.names = T)
fr_tam <- mapply(function(x) readRDS(fr_tam_name[x]), x = 1:9)

pred.tam.fr <- unlist(mapply(function(x) plogis(predict(fr_tam[[x]], newdata = tam[which(tam$year == (x + 2013)), ], allow.new.levels = T)), x = 1:9))
pred.tam.fr2 <- ifelse(pred.tam.fr > 0.5, 1, 0)
obs.tam.fr <- unlist(mapply(function(x) tam$fr_binary[which(tam$year == x + 2013)], x = 1:9))

fr.acc <- rep(NA, length(obs.tam.fr))
fr.acc[which(pred.tam.fr2 == 1 & obs.tam.fr == 1)] <- "TP"
fr.acc[which(pred.tam.fr2 == 0 & obs.tam.fr == 0)] <- "TN"
fr.acc[which(pred.tam.fr2 == 0 & obs.tam.fr == 1)] <- "FN"
fr.acc[which(pred.tam.fr2 == 1 & obs.tam.fr == 0)] <- "FP"

# F1 score is 2*(precision*recall)/(precision+recall)
# precision = TP/(TP+FP)
# recall = TP/ (TP+FN)

fr.precision <- table(fr.acc)[4] / (table(fr.acc)[4] + table(fr.acc)[2])
fr.recall <- table(fr.acc)[4] / (table(fr.acc)[4] + table(fr.acc)[1])

fr.tam.f1 <- 2 * (fr.precision * fr.recall) / (fr.precision + fr.recall)
```


```{r binary models fruiting}
# jack
fr_inter_jack_bin <- glmer(
  fr_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs + (1 | User_Tree_id),
  family = binomial, data = jacktrain
)

summary(fr_inter_jack_bin)
fr_inter_jack_bin

saveRDS(fr_inter_jack_bin, file = "models/fr_inter_jack_bin.rds")

library(sjPlot)
library(patchwork)

p1 <- plot_model(fr_inter_jack_bin, type = "est", show.values = T, show.p = T)
# plot_model(fr_inter_jack_bin, type="eff", transform, show.data = T)
p2 <- plot_model(fr_inter_jack_bin, type = "eff", terms = c("FNSum_PCP_cs", "elev_cs"), transform, show.data = T)
p3 <- plot_model(fr_inter_jack_bin, type = "eff", terms = c("FNMean_MeanSM_cs", "aspect_cs"), transform, show.data = T)
p4 <- plot_model(fr_inter_jack_bin, type = "eff", terms = c("FNSum_DryDays_cs", "light_cs"), transform, show.data = T)

png("plots/jack_bin_est_eff_fr.png", width = 12, height = 8, res = 300, units = "in")
p1 + p2 + p3 + p4
dev.off()

# mango
fr_inter_mango_bin <- glmer(
  fr_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs + (1 | User_Tree_id),
  family = binomial, data = mangotrain
)

summary(fr_inter_mango_bin)

saveRDS(fr_inter_mango_bin, file = "models/fr_inter_mango_bin.rds")


p1 <- plot_model(fr_inter_mango_bin, type = "est", show.values = T, show.p = T)
# plot_model(fr_inter_jack_bin, type="eff", transform, show.data = T)
p2 <- plot_model(fr_inter_mango_bin, type = "eff", terms = c("FNSum_PCP_cs", "elev_cs"), transform, show.data = T)
p3 <- plot_model(fr_inter_mango_bin, type = "eff", terms = c("FNMean_MeanSM_cs", "aspect_cs"), transform, show.data = T)
p4 <- plot_model(fr_inter_mango_bin, type = "eff", terms = c("FNSum_DryDays_cs", "light_cs"), transform, show.data = T)

png("plots/mango_bin_est_eff_fr.png", width = 12, height = 8, res = 300, units = "in")
p1 + p2 + p3 + p4
dev.off()

# tamarind
fr_inter_tam_bin <- glmer(
  fr_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs + (1 | User_Tree_id),
  family = binomial, data = tamtrain
)

summary(fr_inter_tam_bin)

saveRDS(fr_inter_tam_bin, file = "models/fr_inter_tam_bin.rds")


p1 <- plot_model(fr_inter_tam_bin, type = "est", show.values = T, show.p = T)
# plot_model(fr_inter_jack_bin, type="eff", transform, show.data = T)
p2 <- plot_model(fr_inter_tam_bin, type = "eff", terms = c("FNSum_PCP_cs", "elev_cs"), transform, show.data = T)
p3 <- plot_model(fr_inter_tam_bin, type = "eff", terms = c("FNMean_MeanSM_cs", "aspect_cs"), transform, show.data = T)
p4 <- plot_model(fr_inter_tam_bin, type = "eff", terms = c("FNSum_DryDays_cs", "light_cs"), transform, show.data = T)

png("plots/tam_bin_est_eff_fr.png", width = 12, height = 8, res = 300, units = "in")
p1 + p2 + p3 + p4
dev.off()

tab_model(fr_inter_jack_bin, fr_inter_mango_bin, fr_inter_tam_bin, file = "plots/fr_inter_jmt_bin.html")
```


```{r binary model predictions - fruiting and flowering}
library(seqinr)
pred_jack <- plogis(predict(fl_inter_jack_bin, newdata = jacktest2, allow.new.levels = T))

pred_jack2 <- predict(fl_inter_jack_bin, newdata = jacktest2, allow.new.levels = T, type = "response")
rjack <- round(cor(jacktest2$fl_binary, pred_jack, use = "pairwise.complete.obs"), 2)

pred_mango <- plogis(predict(fl_inter_mango_bin, newdata = mangotest, allow.new.levels = T))
rmango <- round(cor(mangotest$fl_binary, pred_mango, use = "pairwise.complete.obs"), 2)

pred_tam <- plogis(predict(fl_inter_tam_bin, newdata = tamtest, allow.new.levels = T))
rtam <- round(cor(tamtest$fl_binary, pred_tam, use = "pairwise.complete.obs"), 2)

png("plots/jmt_pred_fl_bin.png", width = 12, height = 4, res = 300, units = "in")
par(mfrow = c(1, 3))
plot(jitter(jacktest2$fl_binary), pred_jack,
  col = col2alpha(rgb(0, 0, 0), alpha = 0.1), pch = 16,
  main = paste("jack\nr=", rjack, sep = ""), xlab = "test data", ylab = "predicted"
)
plot(jitter(mangotest$fl_binary), pred_mango,
  col = col2alpha(rgb(0, 0, 0), alpha = 0.1), pch = 16,
  main = paste("mango\nr=", rmango, sep = ""), xlab = "test data", ylab = "predicted"
)
plot(jitter(tamtest$fl_binary), pred_tam,
  col = col2alpha(rgb(0, 0, 0), alpha = 0.1), pch = 16,
  main = paste("tamarind\nr=", rtam, sep = ""), xlab = "test data", ylab = "predicted"
)
dev.off()

# fruiting
pred_jack <- plogis(predict(fr_inter_jack_bin, newdata = jacktest, allow.new.levels = T))
rjack <- round(cor(jacktest$fr_binary, pred_jack, use = "pairwise.complete.obs"), 2)

pred_mango <- plogis(predict(fr_inter_mango_bin, newdata = mangotest, allow.new.levels = T))
rmango <- round(cor(mangotest$fr_binary, pred_mango, use = "pairwise.complete.obs"), 2)

pred_tam <- plogis(predict(fr_inter_tam_bin, newdata = tamtest, allow.new.levels = T))
rtam <- round(cor(tamtest$fr_binary, pred_tam, use = "pairwise.complete.obs"), 2)

png("plots/jmt_pred_fr_bin.png", width = 12, height = 4, res = 300, units = "in")
par(mfrow = c(1, 3))
plot(jitter(jacktest$fr_binary), pred_jack,
  col = col2alpha(rgb(0, 0, 0), alpha = 0.1), pch = 16,
  main = paste("jack\nr=", rjack, sep = ""), xlab = "test data", ylab = "predicted"
)
plot(jitter(mangotest$fr_binary), pred_mango,
  col = col2alpha(rgb(0, 0, 0), alpha = 0.1), pch = 16,
  main = paste("mango\nr=", rmango, sep = ""), xlab = "test data", ylab = "predicted"
)
plot(jitter(tamtest$fr_binary), pred_tam,
  col = col2alpha(rgb(0, 0, 0), alpha = 0.1), pch = 16,
  main = paste("tamarind\nr=", rtam, sep = ""), xlab = "test data", ylab = "predicted"
)
dev.off()
```

```{r models using user id as a random effect}
library(lme4)

# jack
fl_jack_bin_usr2 <- glmer(
  fl_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs +
    (1 | User_id / User_Tree_id),
  family = binomial, data = jacktrain
)

saveRDS(fl_jack_bin_usr2, file = "models/fl_jack_bin_usr.rds")

# mango
fl_mango_bin_usr2 <- glmer(
  fl_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs +
    (1 | User_id / User_Tree_id),
  family = binomial, data = mangotrain
)


saveRDS(fl_mango_bin_usr2, file = "models/fl_mango_bin_usr.rds")

# tam
fl_tam_bin_usr2 <- glmer(
  fl_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs +
    (1 | User_id / User_Tree_id),
  family = binomial, data = tamtrain
)

saveRDS(fl_tam_bin_usr2, file = "models/fl_tam_bin_usr.rds")

library(sjPlot)

tab_model(fl_jack_bin_usr2, fl_mango_bin_usr2, fl_tam_bin_usr2, file = "plots/fl_jmt_bin_usr.html")

summary(fl_jack_bin_usr2)
summary(fl_mango_bin_usr2)
summary(fl_tam_bin_usr2)

# fruiting

# jack
fr_jack_bin_usr <- glmer(
  fr_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs +
    (1 | User_id / User_Tree_id),
  family = binomial, data = jacktrain
)

saveRDS(fr_jack_bin_usr, file = "models/fr_jack_bin_usr.rds")

# mango
fr_mango_bin_usr <- glmer(
  fr_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs +
    (1 | User_id / User_Tree_id),
  family = binomial, data = mangotrain
)


saveRDS(fr_mango_bin_usr, file = "models/fr_mango_bin_usr.rds")

# tam
fr_tam_bin_usr <- glmer(
  fr_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs +
    (1 | User_id / User_Tree_id),
  family = binomial, data = tamtrain
)

saveRDS(fr_tam_bin_usr, file = "models/fr_tam_bin_usr.rds")

library(sjPlot)

tab_model(fr_jack_bin_usr, fr_mango_bin_usr, fr_tam_bin_usr, file = "models/fr_jmt_bin_usr.html")

summary(fr_jack_bin_usr)
summary(fr_mango_bin_usr)
summary(fr_tam_bin_usr)
```
```{r models using monthly env variables}
library(lme4)

# using the full data for this

# jack
fl_jack_bin_mn <- glmer(
  fl_binary ~ MNSum_PCP_cs + MNMean_MaxTemp_cs +
    MNMean_MeanSM_cs + MNSum_DryDays_cs +
    elev_cs * MNSum_PCP_cs + aspect_cs * MNMean_MeanSM_cs +
    light_cs * MNSum_DryDays_cs +
    (1 | User_Tree_id),
  family = binomial, data = sw_spat_rescale[sw_spat_rescale$Species_name == "Artocarpus heterophyllus", ]
)

saveRDS(fl_jack_bin_mn, file = "models/fl_jack_bin_mn.rds")

# mango
fl_mango_bin_mn <- glmer(
  fl_binary ~ MNSum_PCP_cs + MNMean_MaxTemp_cs +
    MNMean_MeanSM_cs + MNSum_DryDays_cs +
    elev_cs * MNSum_PCP_cs + aspect_cs * MNMean_MeanSM_cs +
    light_cs * MNSum_DryDays_cs +
    (1 | User_Tree_id),
  family = binomial, data = sw_spat_rescale[sw_spat_rescale$Species_name == "Mangifera indica", ]
)

saveRDS(fl_mango_bin_mn, file = "models/fl_mango_bin_mn.rds")

# tam
fl_tam_bin_mn <- glmer(
  fl_binary ~ MNSum_PCP_cs + MNMean_MaxTemp_cs +
    MNMean_MeanSM_cs + MNSum_DryDays_cs +
    elev_cs * MNSum_PCP_cs + aspect_cs * MNMean_MeanSM_cs +
    light_cs * MNSum_DryDays_cs +
    (1 | User_Tree_id),
  family = binomial, data = sw_spat_rescale[sw_spat_rescale$Species_name == "Tamarindus indica", ]
)

saveRDS(fl_tam_bin_mn, file = "models/fl_tam_bin_mn.rds")

library(sjPlot)

tab_model(fl_jack_bin_mn, fl_mango_bin_mn, fl_tam_bin_mn, file = "plots/fl_jmt_bin_mn.html")

summary(fl_jack_bin_mn)
summary(fl_mango_bin_mn)
summary(fl_tam_bin_mn)


# fruiting


# jack
fr_jack_bin_mn <- glmer(
  fr_binary ~ MNSum_PCP_cs + MNMean_MaxTemp_cs +
    MNMean_MeanSM_cs + MNSum_DryDays_cs +
    elev_cs * MNSum_PCP_cs + aspect_cs * MNMean_MeanSM_cs +
    light_cs * MNSum_DryDays_cs +
    (1 | User_Tree_id),
  family = binomial, data = sw_spat_rescale[sw_spat_rescale$Species_name == "Artocarpus heterophyllus", ]
)

saveRDS(fr_jack_bin_mn, file = "models/fr_jack_bin_mn.rds")

# mango
fr_mango_bin_mn <- glmer(
  fr_binary ~ MNSum_PCP_cs + MNMean_MaxTemp_cs +
    MNMean_MeanSM_cs + MNSum_DryDays_cs +
    elev_cs * MNSum_PCP_cs + aspect_cs * MNMean_MeanSM_cs +
    light_cs * MNSum_DryDays_cs +
    (1 | User_Tree_id),
  family = binomial, data = sw_spat_rescale[sw_spat_rescale$Species_name == "Mangifera indica", ]
)

saveRDS(fr_mango_bin_mn, file = "models/fr_mango_bin_mn.rds")

# tam
fr_tam_bin_mn <- glmer(
  fr_binary ~ MNSum_PCP_cs + MNMean_MaxTemp_cs +
    MNMean_MeanSM_cs + MNSum_DryDays_cs +
    elev_cs * MNSum_PCP_cs + aspect_cs * MNMean_MeanSM_cs +
    light_cs * MNSum_DryDays_cs +
    (1 | User_Tree_id),
  family = binomial, data = sw_spat_rescale[sw_spat_rescale$Species_name == "Tamarindus indica", ]
)

saveRDS(fr_tam_bin_mn, file = "models/fr_tam_bin_mn.rds")

library(sjPlot)

tab_model(fr_jack_bin_mn, fr_mango_bin_mn, fr_tam_bin_mn, file = "plots/fr_jmt_bin_mn.html")

summary(fr_jack_bin_mn)
summary(fr_mango_bin_mn)
summary(fr_tam_bin_mn)
```

```{r location and week as weights}
library(lme4)

# jack
fl_inter_jack_bin_loc_time <- glmer(
  fl_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs + lat_sc + long_sc + month + (1 | User_Tree_id),
  family = binomial, data = jack,
  glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

saveRDS(fl_inter_jack_bin_loc_time, file = "models/fl_inter_jack_bin_loc_time.rds")

fr_inter_jack_bin_loc_time <- glmer(
  fr_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs + lat_sc + long_sc + month + (1 | User_Tree_id),
  family = binomial, data = jack,
  glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

saveRDS(fr_inter_jack_bin_loc_time, file = "models/fr_inter_jack_bin_loc_time.rds")

# mango
fl_inter_mango_bin_loc_time <- glmer(
  fl_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs + lat_sc + long_sc + month + (1 | User_Tree_id),
  family = binomial, data = mango,
  glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

saveRDS(fl_inter_mango_bin_loc_time, file = "models/fl_inter_mango_bin_loc_time.rds")

fr_inter_mango_bin_loc_time <- glmer(
  fr_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs + lat_sc + long_sc + month + (1 | User_Tree_id),
  family = binomial, data = mango,
  glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1000000))
)

saveRDS(fr_inter_mango_bin_loc_time, file = "models/fr_inter_mango_bin_loc_time.rds")

# tam
fl_inter_tam_bin_loc_time <- glmer(
  fl_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs + lat_sc + long_sc + month + (1 | User_Tree_id),
  family = binomial, data = tam,
  glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

saveRDS(fl_inter_tam_bin_loc_time, file = "models/fl_inter_tam_bin_loc_time.rds")

fr_inter_tam_bin_loc_time <- glmer(
  fr_binary ~ FNSum_PCP_cs + FNMean_MaxTemp_cs +
    FNMean_MeanSM_cs + FNSum_DryDays_cs +
    elev_cs * FNSum_PCP_cs + aspect_cs * FNMean_MeanSM_cs +
    light_cs * FNSum_DryDays_cs + lat_sc + long_sc + month + (1 | User_Tree_id),
  family = binomial, data = tam,
  glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000))
)

saveRDS(fr_inter_tam_bin_loc_time, file = "models/fr_inter_tam_bin_loc_time.rds")
```
