---
title: "SW phenology figures"
author: "Krishna Anujan"
date: "17/11/2022"
output: html_document
---

```{r Figure 1}
# Figure 1: Maps - 3 species where are they? Hotspot analysis for observations from Carina (over time most places have added observations)

# jack - #1e824c
# mango - #f7ca18
# tam - #a6915c

# map of Kerala
# library(rgdal)
library(sf)
library(ggplot2)

# KL<-readOGR("C://Users/krish/Documents/SeasonWatch/data/kerala_shapefiles/kerala-master/shapefiles", "state")
KL <- st_read("C://Users/krish/Documents/SeasonWatch/data/kerala_shapefiles/kerala-master/shapefiles/state.shp")

# world_shp<-readOGR("C://Users/krish/Documents/Columbia/Naeem_Lab/Phd_thesis/Chap3/maps/world_borders", "TM_WORLD_BORDERS_SIMPL-0.3")
world_shp <- st_read("C://Users/krish/Documents/Columbia/Naeem_Lab/Phd_thesis/Chap3/maps/world_borders/TM_WORLD_BORDERS_SIMPL-0.3.shp")

# world_shp2<-unionSpatialPolygons(world_shp, IDs=world_shp@data$NAME)



plot(world_shp[world_shp$NAME == "India", ])

# ind2<-readOGR("C://Users/krish/Documents/Andamans/Fejervarya/India-shape/india_shape2", "India_Country_Boundary")
ind2 <- st_read("C://Users/krish/Documents/Andamans/Fejervarya/India-shape/india_shape2/India_Country_Boundary.shp")


# ind<-spTransform(ind2, crs(KL))
ind <- st_transform(ind2, st_crs(KL))
# mybb <- SpatialPolygons(list(Polygons(list(Polygon(extent(and_shp))),"1")), proj4string=crs(and_shp))
# ind<-project(ind2,)


# other images
library(png)
SW_diag <- readPNG("C://Users/krish/Documents/SeasonWatch/MSG/images/1.png")
jack_illus <- readPNG("C://Users/krish/Documents/SeasonWatch/MSG/images/jackfruit_illus.png")
mango_illus <- readPNG("C://Users/krish/Documents/SeasonWatch/MSG/images/mango_illus.png")
tam_illus <- readPNG("C://Users/krish/Documents/SeasonWatch/MSG/images/tamarind_illus.png")


# png("plots/figs/Fig1.png", width=8, height =4, units="in", res=300)
png("plots/figs/Fig1_greyfill.png", width = 8, height = 4, units = "in", res = 300)

par(mfrow = c(1, 3))

# jack
plot(KL)
plot(ind, add = T, border = "grey", col = "grey80")
plot(KL, add = T, col = "white")
points(lat ~ long, sw_spat_forcor[sw_spat_forcor$Species_name == "Artocarpus heterophyllus", ], pch = 16, col = "#1e824c", cex = 0.5)
text(77, 12.5, "jackfruit\nn.tree = 7930\nn.obs=72100", cex = 0.7)
axis(2)
grid()
# points(lat~long, #sw_spat_forcor[sw_spat_forcor$Species_name=="Artocarpus heterophyllus",], cex=0.5)

# mango
plot(KL)
plot(ind, add = T, border = "grey", col = "grey80")
plot(KL, add = T, col = "white")
points(lat ~ long, sw_spat_forcor[sw_spat_forcor$Species_name == "Mangifera indica", ], pch = 16, col = "#f7ca18", cex = 0.5)
text(77, 12.5, "mango\nn.tree = 9924\nn.obs=68701", cex = 0.7)
axis(1)
grid()
# points(lat~long, #sw_spat_forcor[sw_spat_forcor$Species_name=="Mangifera indica",], cex=0.5)

# tamarind
plot(KL)
plot(ind, add = T, border = "grey", col = "grey80")
plot(KL, add = T, col = "white")
points(lat ~ long, sw_spat_forcor[sw_spat_forcor$Species_name == "Tamarindus indica", ], pch = 16, col = "#a6915c", cex = 0.5)
text(77, 12.5, "tamarind\nn.tree = 1751\nn.obs=24214", cex = 0.7)
# points(lat~long, #sw_spat_forcor[sw_spat_forcor$Species_name=="Tamarindus indica",], cex=0.5)
grid()
dev.off()

# images and maps-------------------

png("plots/figs/indmap2.png", width = 4, height = 4, units = "in", res = 300)

# plot(ind, col="white", border="grey40")
# plot(KL, add=T, col="grey")
# axis(1)
# axis(2)
# grid()

# grid()

# plot world, then India, then Kerala using ggplot and sf
ggplot() +
  geom_sf(data = st_as_sf(world_shp), fill = "white", colour = "grey40") +
  geom_sf(data = st_as_sf(ind), fill = "white", colour = "grey40", size = 2, linewidth = 0.4) +
  geom_sf(data = st_as_sf(KL), fill = "grey80", colour = "grey20", size = 2, linewidth = 0.4) +
  # bound to india's borders
  coord_sf(xlim = c(68, 98), ylim = c(6, 38)) +
  theme_bw() +
  xlab("") +
  ylab("") +
  theme(
    axis.line.y = element_line(colour = "black"),
    axis.line.x = element_line(colour = "black"),
    # panel.border = element_blank(),
    # axis.text.x = element_text(colour="white"),
    # axis.ticks.x = element_blank(),
    panel.background = element_rect(fill = "aliceblue", colour = "aliceblue"),
  )

dev.off()
```

```{r maps in ggplot, echo=F}
library(sf)
lat <- as.numeric(sw_spat_forcor$lat)
long <- as.numeric(sw_spat_forcor$long)

sw_coords <- as.data.frame(cbind(long, lat))
sw_points <- SpatialPointsDataFrame(coords = sw_coords, data = sw_spat_forcor, proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))

jack_pts <- ggplot() +
  geom_sf(data = st_as_sf(ind), fill = "grey80", colour = "grey40") +
  geom_sf(data = st_as_sf(KL), fill = "white", colour = "grey20", size = 0.4) +
  geom_sf(data = st_as_sf(sw_points[sw_points$Species_name == "Artocarpus heterophyllus", ]), col = "#1e824c", size = 0.5) +
  coord_sf(xlim = c(75, 77.5), ylim = c(8, 13)) +
  annotate(geom = "text", x = 77, y = 12.5, label = "jackfruit\nn.tree = 7930\nn.obs=72100", size = 3) +
  theme_bw() +
  xlab("") +
  ylab("") +
  theme(
    axis.line.y = element_line(colour = "black"),
    panel.border = element_blank(),
    axis.text.x = element_text(colour = "white"),
    axis.ticks.x = element_blank(),
    # panel.background= element_rect(fill = "aliceblue", colour = "aliceblue"),
  )
jack_pts

mango_pts <- ggplot() +
  geom_sf(data = st_as_sf(ind), fill = "grey80", colour = "grey40") +
  geom_sf(data = st_as_sf(KL), fill = "white", colour = "grey20", size = 0.4) +
  geom_sf(data = st_as_sf(sw_points[sw_points$Species_name == "Mangifera indica", ]), col = "#f7ca18", size = 0.5) +
  coord_sf(xlim = c(75, 77.5), ylim = c(8, 13)) +
  annotate(geom = "text", x = 77, y = 12.5, label = "mango\nn.tree = 9924\nn.obs=68701", size = 3) +
  scale_x_continuous(breaks = seq(75, 78)) +
  theme_bw() +
  xlab("") +
  ylab("") +
  theme(
    axis.line.x = element_line(colour = "black"),
    panel.border = element_blank(),
    axis.text.y = element_text(colour = "white"),
    axis.ticks.y = element_blank(),
    # panel.background= element_rect(fill = "aliceblue", colour = "aliceblue"),
  )
mango_pts

tam_pts <- ggplot() +
  geom_sf(data = st_as_sf(ind), fill = "grey80", colour = "grey40") +
  geom_sf(data = st_as_sf(KL), fill = "white", colour = "grey20", size = 0.4) +
  geom_sf(data = st_as_sf(sw_points[sw_points$Species_name == "Tamarindus indica", ]), col = "#a6915c", size = 0.5) +
  coord_sf(xlim = c(75, 77.5), ylim = c(8, 13)) +
  annotate(geom = "text", x = 77, y = 12.5, label = "tamarind\nn.tree = 1751\nn.obs=24214", size = 3) +
  theme_bw() +
  xlab("") +
  ylab("") +
  theme( # axis.line =  element_line(colour = "black"),
    panel.border = element_blank(),
    axis.text = element_text(colour = "white"),
    axis.ticks = element_blank(),
    # panel.background= element_rect(fill = "aliceblue", colour = "aliceblue"),
  )
tam_pts

library(gridExtra)
png("plots/figs/all_pts2.png", width = 8, height = 4, units = "in", res = 300)
grid.arrange(jack_pts, mango_pts, tam_pts, nrow = 1)
dev.off()
```

```{r read all GLMMs}
# ordinal flowering
fl_inter_jack_po <- readRDS("models/fl_inter_jack_po.rds")
fl_inter_mango_po <- readRDS("models/fl_inter_mango_po.rds")
fl_inter_tam_po <- readRDS("models/fl_inter_tam_po.rds")

# ordinal fruiting
fr_inter_jack_po <- readRDS("models/fr_inter_jack_po.rds")
fr_inter_mango_po <- readRDS("models/fr_inter_mango_po.rds")
fr_inter_tam_po <- readRDS("models/fr_inter_tam_po.rds")

# binary flowering
fl_inter_jack_bin <- readRDS("models/fl_inter_jack_bin.rds")
fl_inter_mango_bin <- readRDS("models/fl_inter_mango_bin.rds")
fl_inter_tam_bin <- readRDS("models/fl_inter_tam_bin.rds")

# binary fruiting
fr_inter_jack_bin <- readRDS("models/fr_inter_jack_bin.rds")
fr_inter_mango_bin <- readRDS("models/fr_inter_mango_bin.rds")
fr_inter_tam_bin <- readRDS("models/fr_inter_tam_bin.rds")
```

```{r Figure 2}
# Figure 2: estimates for all ordinal and binary models - 6 panels (one panel each for one species, each panel has estimates for both fruiting and flowering models - stacked. Grey out the non-significant predictors) (see if ML also fits here)

# read all the binary and ordinal fruiting and flowering models first


library(ordinal)

# estimates calculated from here - https://github.com/runehaubo/ordinal/issues/25

mod <- fl_inter_jack_bin

clmm.est <- function(mod) {
  est <- coef(mod)
  sdhat <- sqrt(diag(vcov(mod)))
  CI <- est + tcrossprod(1.96 * sdhat, c(-1, 1))
  rownames(CI) <- names(est)
  CI <- CI[seq_len(length(est) - 1), ]
  ests <- as.data.frame(cbind(CI, est[seq_len(length(est) - 1)]))
  ests <- cbind(ests, rownames(CI))
  colnames(ests) <- c("lower", "upper", "est", "vars")
  ests$vars <- factor(ests$vars, levels = rownames(CI))
  return(ests)
}

bin.est <- function(mod) {
  est <- coef(summary(mod))[, 1]
  sdhat <- sqrt(diag(vcov(mod)))
  CI <- est + tcrossprod(1.96 * sdhat, c(-1, 1))
  rownames(CI) <- names(est)
  CI <- CI[seq_len(length(est)), ]
  ests <- as.data.frame(cbind(CI, est[seq_len(length(est))]))
  ests <- cbind(ests, rownames(CI))
  colnames(ests) <- c("lower", "upper", "est", "vars")
  ests$vars <- factor(ests$vars, levels = rownames(CI))
  return(ests)
}

# CHECK WHETHER THIS IS ODDS RATIO OR TRANSFORMED
# ANS: It is transformed and not odds ratio

# jack panel
jack_fl_po_est <- clmm.est(fl_inter_jack_po)
jack_fl_po_est$phenophase <- rep("flowering", nrow(jack_fl_po_est))
jack_fr_po_est <- clmm.est(fr_inter_jack_po)
jack_fr_po_est$phenophase <- rep("fruiting", nrow(jack_fr_po_est))
jack_ord_est <- rbind(jack_fl_po_est, jack_fr_po_est)
jack_ord_est$phenophase <- factor(jack_ord_est$phenophase)
jack_ord_est$sig <- mapply(function(p) ifelse(jack_ord_est$upper[p] >= 0 & jack_ord_est$lower[p] <= 0, "no", "yes"), p = seq(1:nrow(jack_ord_est)))
ymin <- min(jack_ord_est$lower)
ymax <- max(jack_ord_est$upper)

# mango panel
mango_fl_po_est <- clmm.est(fl_inter_mango_po)
mango_fl_po_est$phenophase <- rep("flowering", nrow(mango_fl_po_est))
mango_fr_po_est <- clmm.est(fr_inter_mango_po)
mango_fr_po_est$phenophase <- rep("fruiting", nrow(mango_fr_po_est))
mango_ord_est <- rbind(mango_fl_po_est, mango_fr_po_est)
mango_ord_est$phenophase <- factor(mango_ord_est$phenophase)
mango_ord_est$sig <- mapply(function(p) ifelse(mango_ord_est$upper[p] >= 0 & mango_ord_est$lower[p] <= 0, "no", "yes"), p = seq(1:nrow(mango_ord_est)))

# tam panel
tam_fl_po_est <- clmm.est(fl_inter_tam_po)
tam_fl_po_est$phenophase <- rep("flowering", nrow(tam_fl_po_est))
tam_fr_po_est <- clmm.est(fr_inter_tam_po)
tam_fr_po_est$phenophase <- rep("fruiting", nrow(tam_fr_po_est))
tam_ord_est <- rbind(tam_fl_po_est, tam_fr_po_est)
tam_ord_est$phenophase <- factor(tam_ord_est$phenophase)
tam_ord_est$sig <- mapply(function(p) ifelse(tam_ord_est$upper[p] >= 0 & tam_ord_est$lower[p] <= 0, "no", "yes"), p = seq(1:nrow(tam_ord_est)))


library(ggplot2)

# jack - #1e824c
# mango - #f7ca18
# tam - #a6915c

jack_ord_est2 <- jack_ord_est[c(1:9, 13:21), ]
mango_ord_est2 <- mango_ord_est[c(1:9, 13:21), ]
tam_ord_est2 <- tam_ord_est[c(1:9, 13:21), ]

jack_ord <- ggplot(jack_ord_est, aes(y = est, x = vars, group = pheno, shape = pheno, color = sig)) +
  geom_hline(yintercept = 0, col = "grey", size = 1.5) +
  geom_point(aes(color = sig), position = position_dodge(width = 0.5)) +
  scale_color_manual(values = c("grey", "#1e824c")) +
  geom_errorbar(aes(x = vars, ymin = lower, ymax = upper, color = sig), position = position_dodge(width = 0.9)) +
  ggtitle("jackfruit") +
  # scale_y_discrete(breaks=seq(-0.5, 0.5, 0.2), labels=as.character(seq(-0.5, 0.5, 0.2)))+
  # ylim(ymin, ymax)+
  coord_flip() +
  theme_minimal()

jack_ord2 <- ggplot(jack_ord_est2, aes(y = est, x = vars, group = phenophase, shape = phenophase, color = sig)) +
  geom_hline(yintercept = 0, col = "grey", size = 1.5) +
  geom_point(aes(color = sig), position = position_dodge(width = 0.5)) +
  scale_color_manual(values = c("grey", "#1e824c")) +
  geom_linerange(aes(x = vars, ymin = lower, ymax = upper, color = sig), position = position_dodge(width = 0.5)) +
  ggtitle("jackfruit") +
  xlab("ordinal model predictors") +
  ylab("model coefficient") +
  # scale_y_discrete(breaks=seq(-0.5, 0.5, 0.2), labels=as.character(seq(-0.5, 0.5, 0.2)))+
  # ylim(ymin, ymax)+
  coord_flip() +
  theme_minimal() +
  guides(shape = FALSE, color = FALSE)
jack_ord2

mango_ord <- ggplot(mango_ord_est, aes(y = est, x = vars, group = pheno, shape = pheno, color = sig)) +
  geom_hline(yintercept = 0, col = "grey", size = 1.5) +
  geom_point(aes(color = sig), position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(x = vars, ymin = lower, ymax = upper, color = sig), position = position_dodge(width = 0.9)) +
  scale_color_manual(values = c("grey", "#f7ca18")) +
  ggtitle("mango") +
  ylab("model coefficient") +
  scale_x_discrete(labels = NULL, name = NULL) +
  # ylim(NA, NA)+
  coord_flip() +
  theme_minimal()

mango_ord2 <- ggplot(mango_ord_est2, aes(y = est, x = vars, group = phenophase, shape = phenophase, color = sig)) +
  geom_hline(yintercept = 0, col = "grey", size = 1.5) +
  geom_point(aes(color = sig), position = position_dodge(width = 0.5)) +
  geom_linerange(aes(x = vars, ymin = lower, ymax = upper, color = sig), position = position_dodge(width = 0.5)) +
  scale_color_manual(values = c("grey", "#f7ca18")) +
  ggtitle("mango") +
  ylab("model coefficient") +
  scale_x_discrete(labels = NULL, name = NULL) +
  xlab("") +
  guides(shape = FALSE, color = FALSE) +
  # ylim(NA, NA)+
  coord_flip() +
  theme_minimal()
mango_ord2

tam_ord <- ggplot(tam_ord_est, aes(y = est, x = vars, group = pheno, shape = pheno, color = sig)) +
  geom_hline(yintercept = 0, col = "grey", size = 1.5) +
  geom_point(aes(color = sig), position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(x = vars, ymin = lower, ymax = upper, color = sig), position = position_dodge(width = 0.9)) +
  scale_color_manual(values = c("grey", "#a6915c")) +
  ggtitle("tamarind") +
  ylab("model coefficient") +
  scale_x_discrete(labels = NULL, name = NULL) +
  # ylim(NA, NA)+
  coord_flip() +
  theme_minimal()

tam_ord2 <- ggplot(tam_ord_est2, aes(y = est, x = vars, group = phenophase, shape = phenophase, color = sig)) +
  geom_hline(yintercept = 0, col = "grey", size = 1.5) +
  geom_point(aes(color = sig), position = position_dodge(width = 0.5)) +
  geom_linerange(aes(x = vars, ymin = lower, ymax = upper, color = sig), position = position_dodge(width = 0.5)) +
  scale_color_manual(values = c("grey", "#a6915c")) +
  ggtitle("tamarind") +
  xlab("") +
  ylab("model coefficient") +
  scale_x_discrete(labels = NULL, name = NULL) +
  guides(color = FALSE) +
  # ylim(NA, NA)+
  coord_flip() +
  theme_minimal()
tam_ord2

library(gridExtra)

library(patchwork)

# binomial models
library(lme4)
jack_fl_bin_est <- bin.est(fl_inter_jack_bin)
jack_fl_bin_est$phenophase <- rep("flowering", nrow(jack_fl_bin_est))
jack_fr_bin_est <- bin.est(summary(fr_inter_jack_bin))
jack_fr_bin_est$phenophase <- rep("fruiting", nrow(jack_fr_bin_est))
jack_bin_est <- rbind(jack_fl_bin_est, jack_fr_bin_est)
jack_bin_est$phenophase <- factor(jack_bin_est$phenophase)
jack_bin_est$sig <- mapply(function(p) ifelse(jack_bin_est$upper[p] >= 0 & jack_bin_est$lower[p] <= 0, "no", "yes"), p = seq(1:nrow(jack_bin_est)))

# mango panel
mango_fl_bin_est <- bin.est(fl_inter_mango_bin)
mango_fl_bin_est$phenophase <- rep("flowering", nrow(mango_fl_bin_est))
mango_fr_bin_est <- bin.est(fr_inter_mango_bin)
mango_fr_bin_est$phenophase <- rep("fruiting", nrow(mango_fr_bin_est))
mango_bin_est <- rbind(mango_fl_bin_est, mango_fr_bin_est)
mango_bin_est$phenophase <- factor(mango_bin_est$phenophase)
mango_bin_est$sig <- mapply(function(p) ifelse(mango_bin_est$upper[p] >= 0 & mango_bin_est$lower[p] <= 0, "no", "yes"), p = seq(1:nrow(mango_bin_est)))

# tam panel
tam_fl_bin_est <- bin.est(fl_inter_tam_bin)
tam_fl_bin_est$phenophase <- rep("flowering", nrow(tam_fl_bin_est))
tam_fr_bin_est <- bin.est(fr_inter_tam_bin)
tam_fr_bin_est$phenophase <- rep("fruiting", nrow(tam_fr_bin_est))
tam_bin_est <- rbind(tam_fl_bin_est, tam_fr_bin_est)
tam_bin_est$phenophase <- factor(tam_bin_est$phenophase)
tam_bin_est$sig <- mapply(function(p) ifelse(tam_bin_est$upper[p] >= 0 & tam_bin_est$lower[p] <= 0, "no", "yes"), p = seq(1:nrow(tam_bin_est)))

jack_bin_est2 <- jack_bin_est[c(1:8, 12:19), ]
mango_bin_est2 <- mango_bin_est[c(1:8, 12:19), ]
tam_bin_est2 <- tam_bin_est[c(1:8, 12:19), ]

jack_bin <- ggplot(jack_bin_est, aes(y = est, x = vars, group = pheno, shape = pheno, color = sig)) +
  geom_hline(yintercept = 0, col = "grey", size = 1.5) +
  geom_point(aes(color = sig), position = position_dodge(width = 0.9)) +
  scale_color_manual(values = c("grey", "#1e824c")) +
  geom_errorbar(aes(x = vars, ymin = lower, ymax = upper, color = sig), position = position_dodge(width = 0.9)) +
  ggtitle("jackfruit") +
  # scale_y_discrete(breaks=seq(-0.5, 0.5, 0.2), labels=as.character(seq(-0.5, 0.5, 0.2)))+
  # ylim(ymin, ymax)+
  coord_flip() +
  theme_minimal()

jack_bin2 <- ggplot(jack_bin_est2, aes(y = est, x = vars, group = phenophase, shape = phenophase, color = sig)) +
  geom_hline(yintercept = 0, col = "grey", size = 1.5) +
  geom_point(aes(color = sig), position = position_dodge(width = 0.5)) +
  scale_color_manual(values = c("grey", "#1e824c")) +
  geom_linerange(aes(x = vars, ymin = lower, ymax = upper, color = sig), position = position_dodge(width = 0.5)) +
  ggtitle("jackfruit") +
  xlab("binary model predictors") +
  ylab("model coefficient") +
  guides(shape = "none", color = "none") +
  # scale_y_discrete(breaks=seq(-0.5, 0.5, 0.2), labels=as.character(seq(-0.5, 0.5, 0.2)))+
  # ylim(ymin, ymax)+
  coord_flip() +
  theme_minimal()
jack_bin2

mango_bin <- ggplot(mango_bin_est, aes(y = est, x = vars, group = pheno, shape = pheno, color = sig)) +
  geom_hline(yintercept = 0, col = "grey", size = 1.5) +
  geom_point(aes(color = sig), position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(x = vars, ymin = lower, ymax = upper, color = sig), position = position_dodge(width = 0.9)) +
  scale_color_manual(values = c("grey", "#f7ca18")) +
  ggtitle("mango") +
  scale_x_discrete(labels = NULL, name = NULL) +
  # ylim(NA, NA)+
  coord_flip() +
  theme_minimal()

mango_bin2 <- ggplot(mango_bin_est2, aes(y = est, x = vars, group = phenophase, shape = phenophase, color = sig)) +
  geom_hline(yintercept = 0, col = "grey", size = 1.5) +
  geom_point(aes(color = sig), position = position_dodge(width = 0.5)) +
  geom_linerange(aes(x = vars, ymin = lower, ymax = upper, color = sig), position = position_dodge(width = 0.5)) +
  scale_color_manual(values = c("grey", "#f7ca18")) +
  ggtitle("mango") +
  guides(shape = "none", color = "none") +
  scale_x_discrete(labels = NULL, name = NULL) +
  xlab("") +
  ylab("model coefficient") +
  # ylim(NA, NA)+
  coord_flip() +
  theme_minimal()
mango_bin2

tam_bin <- ggplot(tam_bin_est, aes(y = est, x = vars, group = phenophase, shape = phenophase, color = sig)) +
  geom_hline(yintercept = 0, col = "grey", size = 1.5) +
  geom_point(aes(color = sig), position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(x = vars, ymin = lower, ymax = upper, color = sig), position = position_dodge(width = 0.9)) +
  scale_color_manual(values = c("grey", "#a6915c")) +
  ggtitle("tamarind") +
  scale_x_discrete(labels = NULL, name = NULL) +

  # ylim(NA, NA)+
  coord_flip() +
  theme_minimal()


tam_bin2 <- ggplot(tam_bin_est2, aes(y = est, x = vars, group = phenophase, shape = phenophase, color = sig)) +
  geom_hline(yintercept = 0, col = "grey", size = 1.5) +
  geom_point(aes(color = sig), position = position_dodge(width = 0.5)) +
  geom_linerange(aes(x = vars, ymin = lower, ymax = upper, color = sig), position = position_dodge(width = 0.5)) +
  scale_color_manual(values = c("grey", "#a6915c")) +
  ggtitle("tamarind") +
  scale_x_discrete(labels = NULL, name = NULL) +
  guides(color = "none") +
  xlab("") +
  ylab("model coefficient") +
  # ylim(NA, NA)+
  coord_flip() +
  theme_minimal()
tam_bin2

png("plots/figs/Fig2.png", width = 12, height = 8, units = "in", res = 300)
(jack_ord + mango_ord + tam_ord) / (jack_bin + mango_bin + tam_bin)
dev.off()

png("plots/figs/Fig2_noint_Jan24.png", width = 12, height = 8, units = "in", res = 300)
(jack_ord2 + mango_ord2 + tam_ord2) / (jack_bin2 + mango_bin2 + tam_bin2)
dev.off()
```

```{r Figure 2 LOOCV}
# read all the csvs

# ordinal
jack.bin.ests_fl <- read.csv("models/jack.bin.ests_fl.csv")
jack.bin.ests_fr <- read.csv("models/jack.bin.ests_fr.csv")
jack.bin.std.er_fl <- read.csv("models/jack.bin.std.er_fl.csv")
jack.bin.std.er_fr <- read.csv("models/jack.bin.std.er_fr.csv")

mango.bin.ests_fl <- read.csv("models/mango.bin.ests_fl.csv")
mango.bin.ests_fr <- read.csv("models/mango.bin.ests_fr.csv")
mango.bin.std.er_fl <- read.csv("models/mango.bin.std.er_fl.csv")
mango.bin.std.er_fr <- read.csv("models/mango.bin.std.er_fr.csv")

tam.bin.ests_fl <- read.csv("models/tam.bin.ests_fl.csv")
tam.bin.ests_fr <- read.csv("models/tam.bin.ests_fr.csv")
tam.bin.std.er_fl <- read.csv("models/tam.bin.std.er_fl.csv")
tam.bin.std.er_fr <- read.csv("models/tam.bin.std.er_fr.csv")

# ordinal
jack.ord.ests_fl <- read.csv("models/jack.ord.ests_fl.csv")
jack.ord.ests_fr <- read.csv("models/jack.ord.ests_fr.csv")
jack.ord.std.er_fl <- read.csv("models/jack.ord.std.er_fl.csv")
jack.ord.std.er_fr <- read.csv("models/jack.ord.std.er_fr.csv")

mango.ord.ests_fl <- read.csv("models/mango.ord.ests_fl.csv")
mango.ord.ests_fr <- read.csv("models/mango.ord.ests_fr.csv")
mango.ord.std.er_fl <- read.csv("models/mango.ord.std.er_fl.csv")
mango.ord.std.er_fr <- read.csv("models/mango.ord.std.er_fr.csv")

tam.ord.ests_fl <- read.csv("models/tam.ord.ests_fl.csv")
tam.ord.ests_fr <- read.csv("models/tam.ord.ests_fr.csv")
tam.ord.std.er_fl <- read.csv("models/tam.ord.std.er_fl.csv")
tam.ord.std.er_fr <- read.csv("models/tam.ord.std.er_fr.csv")

# binding them

# binary
all_glmm_bin <- rbind(jack.bin.ests_fl, jack.bin.ests_fr, mango.bin.ests_fl, mango.bin.ests_fr, tam.bin.ests_fl, tam.bin.ests_fr)
all_glmm_bin$sp <- c(rep("jackfruit", 22), rep("mango", 22), rep("tamarind", 22))
all_glmm_bin$phenology <- rep(c(rep("flowering", 11), rep("fruiting", 11)), 3)
all_glmm_bin$mean <- apply(all_glmm_bin[, 2:10], 1, mean)
all_glmm_bin$sd <- apply(all_glmm_bin[, 2:10], 1, sd)

# binary with std.er
all_glmm_bin_er <- rbind(jack.bin.std.er_fl, jack.bin.std.er_fr, mango.bin.std.er_fl, mango.bin.std.er_fr, tam.bin.std.er_fl, tam.bin.std.er_fr)

all_glmm_bin_er$sp <- c(rep("jackfruit", 22), rep("mango", 22), rep("tamarind", 22))
all_glmm_bin_er$phenology <- rep(c(rep("flowering", 11), rep("fruiting", 11)), 3)
all_glmm_bin_er$mean <- apply(all_glmm_bin_er[, 2:10], 1, mean, na.rm = T)

all_glmm_bin$std.er <- all_glmm_bin_er$mean

all_glmm_bin$sig <- mapply(function(p) ifelse((all_glmm_bin$mean[p] - (1.96 * all_glmm_bin$std.er[p])) >= 0 & (all_glmm_bin$mean[p] - (1.96 * all_glmm_bin$std.er[p])) <= 0, "no", "yes"), p = seq(1:nrow(all_glmm_bin)))


# stop-gap fixes for plot - change later
mango.ord.ests_fl[, 1] <- NULL
mango.ord.ests_fr[, 1] <- NULL
jack.ord.ests_fl[, 1] <- NULL
jack.ord.ests_fr[, 1] <- NULL
# jack.ord.ests_fl[,2]<-rep(NA, nrow(jack.ord.ests_fl))
jack.ord.ests_fr[, 2] <- rep(NA, nrow(jack.ord.ests_fr))

# ordinal
all_glmm_ord <- rbind(jack.ord.ests_fl, jack.ord.ests_fr, mango.ord.ests_fl, mango.ord.ests_fr, tam.ord.ests_fl, tam.ord.ests_fr)

all_glmm_ord$sp <- c(rep("jackfruit", 24), rep("mango", 24), rep("tamarind", 24))
all_glmm_ord$phenology <- rep(c(rep("flowering", 12), rep("fruiting", 12)), 3)
all_glmm_ord$mean <- apply(all_glmm_ord[, 2:10], 1, mean, na.rm = T)
all_glmm_ord$sd <- apply(all_glmm_ord[, 2:10], 1, sd, na.rm = T)

# std error - stop gap fixes
jack.ord.std.er_fl[, 3] <- rep(NA, nrow(jack.ord.std.er_fl))
jack.ord.std.er_fr[, 3] <- jack.ord.std.er_fr[, 12]
jack.ord.std.er_fr[, 12] <- NULL
jack.ord.ests_fr[, 2] <- rep(NA, nrow(jack.ord.ests_fr))
mango.ord.std.er_fl[, 4] <- rep(NA, nrow(mango.ord.std.er_fl))

jack.ord.std.er_fl[, 1] <- NULL
jack.ord.std.er_fr[, 1] <- NULL
mango.ord.std.er_fl[, 1] <- NULL
mango.ord.std.er_fr[, 1] <- NULL

all_glmm_ord_er <- rbind(jack.ord.std.er_fl, jack.ord.std.er_fr, mango.ord.std.er_fl, mango.ord.std.er_fr, tam.ord.std.er_fl, tam.ord.std.er_fr)

all_glmm_ord_er$sp <- c(rep("jackfruit", 24), rep("mango", 24), rep("tamarind", 24))
all_glmm_ord_er$phenology <- rep(c(rep("flowering", 12), rep("fruiting", 12)), 3)
all_glmm_ord_er$mean <- apply(all_glmm_ord_er[, 2:10], 1, mean, na.rm = T)

all_glmm_ord$std.er <- all_glmm_ord_er$mean

all_glmm_ord$sig <- mapply(function(p) ifelse((all_glmm_ord$mean[p] - (1.96 * all_glmm_ord$std.er[p])) >= 0 & (all_glmm_ord$mean[p] - (1.96 * all_glmm_ord$std.er[p])) <= 0, "no", "yes"), p = seq(1:nrow(all_glmm_ord)))


# jack - #1e824c
# mango - #f7ca18
# tam - #a6915c

# the plots

newlabs <- c(
  "intercept", "precipitation", "maximum temperature", "mean soil moisture", "consecutive dry days", "elevation", "aspect", "urbanisation",
  "precipitation:elevation", "soil moisture:aspect", "dry days:urbanisation"
)

all_glmm_bin$newlabs <- rep(newlabs, 6)
all_glmm_bin$newlabs <- factor(all_glmm_bin$newlabs, levels = newlabs)

newlabs_o <- c(
  "0|1", "1|2", "precipitation", "maximum temperature", "mean soil moisture", "consecutive dry days", "elevation", "aspect", "urbanisation",
  "precipitation:elevation", "soil moisture:aspect", "dry days:urbanisation"
)
all_glmm_ord$newlabs <- rep(newlabs_o, 6)
all_glmm_ord$newlabs <- factor(all_glmm_ord$newlabs, levels = newlabs_o)


library(boot)
library(ggplot2)
jack_glmm_bin_loocv <- ggplot(all_glmm_bin[all_glmm_bin$sp == "jackfruit", ], aes(x = newlabs, y = mean, shape = phenology, color = sig)) +
  geom_point(position = position_dodge(width = 0.9), col = "#1e824c", size = 2) +
  geom_errorbar(aes(ymin = mean - (1.96 * std.er), ymax = mean + (1.96 * std.er)), position = position_dodge(width = 0.9), col = "#1e824c", width = 0) +
  coord_flip() +
  geom_hline(yintercept = 0, col = "grey", size = 1.5) +
  ggtitle("Jackfruit") +
  ylab("Model coefficient") +
  xlab("Binary predictors") +
  guides(color = "none") +
  labs(tag = "d)") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 12))

mango_glmm_bin_loocv <- ggplot(all_glmm_bin[all_glmm_bin$sp == "mango", ], aes(x = newlabs, y = mean, shape = phenology)) +
  geom_point(position = position_dodge(width = 0.9), col = "#f7ca18", size = 2) +
  geom_errorbar(aes(ymin = mean - (1.96 * std.er), ymax = mean + (1.96 * std.er)), position = position_dodge(width = 0.9), col = "#f7ca18", width = 0) +
  labs(tag = "e)") +
  coord_flip() +
  geom_hline(yintercept = 0, col = "grey", size = 1.5) +
  ggtitle("Mango") +
  ylab("Model coefficient") +
  guides(color = "none") +
  xlab("Binary predictors") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 12))

tam_glmm_bin_loocv <- ggplot(all_glmm_bin[all_glmm_bin$sp == "tamarind", ], aes(x = newlabs, y = mean, shape = phenology)) +
  geom_point(position = position_dodge(width = 0.9), col = "#a6915c", size = 2) +
  geom_errorbar(aes(ymin = mean - (1.96 * std.er), ymax = mean + (1.96 * std.er)), position = position_dodge(width = 0.9), col = "#a6915c", width = 0) +
  coord_flip() +
  labs(tag = "f)") +
  geom_hline(yintercept = 0, col = "grey", size = 1.5) +
  ggtitle("Tamarind") +
  ylab("Model coefficient") +
  xlab("Binary predictors") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 12))

library(gridExtra)

# png("plots/figs/Fig2_loocv_binonly.png", width=16, height=8, units="in", res=300)
# grid.arrange(jack_glmm_bin_loocv, mango_glmm_bin_loocv, tam_glmm_bin_loocv, ncol=3)
# dev.off()


# ordinal
jack_glmm_ord_loocv <- ggplot(all_glmm_ord[all_glmm_ord$sp == "jackfruit", ], aes(x = newlabs, y = mean, shape = phenology)) +
  geom_point(position = position_dodge(width = 0.9), col = "#1e824c", size = 2) +
  #  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd),
  geom_errorbar(aes(ymin = mean - (1.96 * std.er), ymax = mean + (1.96 * std.er)), position = position_dodge(width = 0.9), col = "#1e824c", width = 0) +
  coord_flip() +
  labs(tag = "a)") +
  geom_hline(yintercept = 0, col = "grey", size = 1.5) +
  ggtitle("Jackfruit") +
  ylab("Model coefficient") +
  xlab("Ordinal predictors") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 12))

mango_glmm_ord_loocv <- ggplot(all_glmm_ord[all_glmm_ord$sp == "mango", ], aes(x = newlabs, y = mean, shape = phenology)) +
  geom_point(position = position_dodge(width = 0.9), col = "#f7ca18", size = 2) +
  #  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd),
  geom_errorbar(aes(ymin = mean - (1.96 * std.er), ymax = mean + (1.96 * std.er)), position = position_dodge(width = 0.9), col = "#f7ca18", width = 0) +
  coord_flip() +
  labs(tag = "b)") +
  geom_hline(yintercept = 0, col = "grey", size = 1.5) +
  ggtitle("Mango") +
  ylab("Model coefficient") +
  xlab("Ordinal predictors") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 12))

tam_glmm_ord_loocv <- ggplot(all_glmm_ord[all_glmm_ord$sp == "tamarind", ], aes(x = newlabs, y = mean, shape = phenology)) +
  geom_point(position = position_dodge(width = 0.9), col = "#a6915c", size = 2) +
  #  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd),
  geom_errorbar(aes(ymin = mean - (1.96 * std.er), ymax = mean + (1.96 * std.er)), position = position_dodge(width = 0.9), col = "#a6915c", width = 0) +
  coord_flip() +
  labs(tag = "c)") +
  geom_hline(yintercept = 0, col = "grey", size = 1.5) +
  ggtitle("Tamarind") +
  ylab("Model coefficient") +
  xlab("Ordinal predictors") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 12))

png("plots/figs/Fig2_loocv_stderJan24.png", width = 16, height = 8, units = "in", res = 300)
grid.arrange(jack_glmm_ord_loocv, mango_glmm_ord_loocv, tam_glmm_ord_loocv,
  jack_glmm_bin_loocv, mango_glmm_bin_loocv, tam_glmm_bin_loocv,
  ncol = 3
)
dev.off()
```

```{r Figure 3}
# Figure 3: predicted probabilities fixed effects for three species - from ML models

# reading the files

# fl binary
ml_jack_fl_bin <- read.csv("C://Users/krish/Documents/SeasonWatch/MSG/ML_results/Jackfruit_ObsLevel_FNVarsOnly/Jackfruit_ObsLevel_FNVarsOnly/RFImportanceMDA_Jackfruit_ObsLevel_FNVarsOnly_FLBinary.csv")
ml_mango_fl_bin <- read.csv("C://Users/krish/Documents/SeasonWatch/MSG/ML_results/Mango_ObsLevel_FNVarsOnly/Mango_ObsLevel_FNVarsOnly/RFImportanceMDA_Mango_ObsLevel_FNVarsOnly_FLBinary.csv")
ml_tam_fl_bin <- read.csv("C://Users/krish/Documents/SeasonWatch/MSG/ML_results/Tamarind_ObsLevel_FNVarsOnly/Tamarind_ObsLevel_FNVarsOnly/RFImportanceMDA_Tamarind_ObsLevel_FNVarsOnly_FLBinary.csv")

# fr binary
ml_jack_fr_bin <- read.csv("C://Users/krish/Documents/SeasonWatch/MSG/ML_results/Jackfruit_ObsLevel_FNVarsOnly/Jackfruit_ObsLevel_FNVarsOnly/RFImportanceMDA_Jackfruit_ObsLevel_FNVarsOnly_FRBinary.csv")
ml_mango_fr_bin <- read.csv("C://Users/krish/Documents/SeasonWatch/MSG/ML_results/Mango_ObsLevel_FNVarsOnly/Mango_ObsLevel_FNVarsOnly/RFImportanceMDA_Mango_ObsLevel_FNVarsOnly_FRBinary.csv")
ml_tam_fr_bin <- read.csv("C://Users/krish/Documents/SeasonWatch/MSG/ML_results/Tamarind_ObsLevel_FNVarsOnly/Tamarind_ObsLevel_FNVarsOnly/RFImportanceMDA_Tamarind_ObsLevel_FNVarsOnly_FRBinary.csv")

# fl ordinal
ml_jack_fl_ord <- read.csv("C://Users/krish/Documents/SeasonWatch/MSG/ML_results/Jackfruit_ObsLevel_FNVarsOnly/Jackfruit_ObsLevel_FNVarsOnly/CatBoostImportanceMDA_Jackfruit_ObsLevel_FNVarsOnly_FLIntensity.csv")
ml_mango_fl_ord <- read.csv("C://Users/krish/Documents/SeasonWatch/MSG/ML_results/Mango_ObsLevel_FNVarsOnly/Mango_ObsLevel_FNVarsOnly/CatBoostImportanceMDA_Mango_ObsLevel_FNVarsOnly_FLIntensity.csv")
ml_tam_fl_ord <- read.csv("C://Users/krish/Documents/SeasonWatch/MSG/ML_results/Tamarind_ObsLevel_FNVarsOnly/Tamarind_ObsLevel_FNVarsOnly/CatBoostImportanceMDA_Tamarind_ObsLevel_FNVarsOnly_FLIntensity.csv")

# fr ordinal
ml_jack_fr_ord <- read.csv("C://Users/krish/Documents/SeasonWatch/MSG/ML_results/Jackfruit_ObsLevel_FNVarsOnly/Jackfruit_ObsLevel_FNVarsOnly/CatBoostImportanceMDA_Jackfruit_ObsLevel_FNVarsOnly_FRIntensity.csv")
ml_mango_fr_ord <- read.csv("C://Users/krish/Documents/SeasonWatch/MSG/ML_results/Mango_ObsLevel_FNVarsOnly/Mango_ObsLevel_FNVarsOnly/CatBoostImportanceMDA_Mango_ObsLevel_FNVarsOnly_FRIntensity.csv")
ml_tam_fr_ord <- read.csv("C://Users/krish/Documents/SeasonWatch/MSG/ML_results/Tamarind_ObsLevel_FNVarsOnly/Tamarind_ObsLevel_FNVarsOnly/CatBoostImportanceMDA_Tamarind_ObsLevel_FNVarsOnly_FRIntensity.csv")

# joining
all_ml_ord <- rbind(ml_jack_fl_ord, ml_jack_fr_ord, ml_mango_fl_ord, ml_mango_fr_ord, ml_tam_fl_ord, ml_tam_fr_ord)
all_ml_ord$sp <- c(rep("jackfruit", 22), rep("mango", 22), rep("tamarind", 22))
all_ml_ord$phenology <- rep(c(rep("flowering", 11), rep("fruiting", 11)), 3)
all_ml_ord$sd <- apply(all_ml_ord[, 3:11], 1, sd)
all_ml_ord$X <- factor(all_ml_ord$X, levels = c(
  "FNSum_PCP", "FNMean_MaxTemp", "FNMean_MeanTemp", "FNMean_MinTemp",
  "FNMean_MeanSM", "FNSum_SolarRad", "FNSum_ConsDryDays", "elev", "slope", "aspect", "light_vals"
))


all_ml_bin <- rbind(ml_jack_fl_bin, ml_jack_fr_bin, ml_mango_fl_bin, ml_mango_fr_bin, ml_tam_fl_bin, ml_tam_fr_bin)
all_ml_bin$sp <- c(rep("jackfruit", 22), rep("mango", 22), rep("tamarind", 22))
all_ml_bin$phenology <- rep(c(rep("flowering", 11), rep("fruiting", 11)), 3)
all_ml_bin$sd <- apply(all_ml_bin[, 3:11], 1, sd)
all_ml_bin$X <- factor(all_ml_bin$X, levels = c(
  "FNSum_PCP", "FNMean_MaxTemp", "FNMean_MeanTemp", "FNMean_MinTemp",
  "FNMean_MeanSM", "FNSum_SolarRad", "FNSum_ConsDryDays", "elev", "slope", "aspect", "light_vals"
))

# remotes::install_github("coolbutuseless/ggpattern")
library(ggplot2)
library(ggpattern)

# jack - #1e824c
# mango - #f7ca18
# tam - #a6915c

# jack_ml_ord<-ggplot(all_ml_ord[all_ml_ord$sp=="jackfruit",], aes(x=X, y=Overall..average., fill=phenology))+
jack_ml_ord <- ggplot(all_ml_ord[all_ml_ord$sp == "jackfruit", ], aes(x = X, y = Overall..average., shape = phenology)) +
  # geom_bar(stat="identity", position = position_dodge())+
  geom_point(position = position_dodge(width = 0.9), col = "#1e824c") +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd), position = position_dodge()) +
  coord_flip() +
  ggtitle("jackfruit") +
  ylab("variable importance") +
  xlab("ordinal predictors") #+
# scale_fill_manual(values=c("#1e824c", "#1e824c"))+
# geom_col_pattern(aes(pattern_fill=phenology), pattern="stripe", fill="white", colour="#1e824c", position=position_dodge())
# geom_bar_pattern(stat="identity", pattern_color="white",
# pattern_fill="black", position=position_dodge(), aes(pattern=phenology))

mango_ml_ord <- ggplot(all_ml_ord[all_ml_ord$sp == "mango", ], aes(x = X, y = Overall..average., fill = phenology)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd), position = position_dodge()) +
  coord_flip() +
  ggtitle("mango") +
  ylab("variable importance") +
  xlab("ordinal predictors") +
  # scale_color_manual(values=c("#1e824c", "#1e824c"))+
  scale_fill_manual(values = c("#f7ca18", "#f7ca18")) +
  # geom_col_pattern(aes(pattern_fill=phenology), pattern="stripe", fill="white", colour="#1e824c", position=position_dodge())
  geom_bar_pattern(
    stat = "identity", pattern_color = "white",
    pattern_fill = "black", position = position_dodge(), aes(pattern = phenology)
  )

tam_ml_ord <- ggplot(all_ml_ord[all_ml_ord$sp == "tamarind", ], aes(x = X, y = Overall..average., fill = phenology)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd), position = position_dodge()) +
  coord_flip() +
  ggtitle("tamarind") +
  ylab("variable importance") +
  xlab("ordinal predictors") +
  # scale_color_manual(values=c("#1e824c", "#1e824c"))+
  scale_fill_manual(values = c("#a6915c", "#a6915c")) +
  # geom_col_pattern(aes(pattern_fill=phenology), pattern="stripe", fill="white", colour="#1e824c", position=position_dodge())
  geom_bar_pattern(
    stat = "identity", pattern_color = "white",
    pattern_fill = "black", position = position_dodge(), aes(pattern = phenology)
  )


jack_ml_bin <- ggplot(all_ml_bin[all_ml_bin$sp == "jackfruit", ], aes(x = X, y = Overall..average., fill = phenology)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd), position = position_dodge()) +
  coord_flip() +
  ggtitle("jackfruit") +
  ylab("variable importance") +
  xlab("binary predictors") +
  # scale_color_manual(values=c("#1e824c", "#1e824c"))+
  scale_fill_manual(values = c("#1e824c", "#1e824c")) +
  # geom_col_pattern(aes(pattern_fill=phenology), pattern="stripe", fill="white", colour="#1e824c", position=position_dodge())
  geom_bar_pattern(
    stat = "identity", pattern_color = "white",
    pattern_fill = "black", position = position_dodge(), aes(pattern = phenology)
  )

mango_ml_bin <- ggplot(all_ml_bin[all_ml_bin$sp == "mango", ], aes(x = X, y = Overall..average., fill = phenology)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd), position = position_dodge()) +
  coord_flip() +
  ggtitle("mango") +
  ylab("variable importance") +
  xlab("binary predictors") +
  # scale_color_manual(values=c("#1e824c", "#1e824c"))+
  scale_fill_manual(values = c("#f7ca18", "#f7ca18")) +
  # geom_col_pattern(aes(pattern_fill=phenology), pattern="stripe", fill="white", colour="#1e824c", position=position_dodge())
  geom_bar_pattern(
    stat = "identity", pattern_color = "white",
    pattern_fill = "black", position = position_dodge(), aes(pattern = phenology)
  )

tam_ml_bin <- ggplot(all_ml_bin[all_ml_bin$sp == "tamarind", ], aes(x = X, y = Overall..average., fill = phenology)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd), position = position_dodge()) +
  coord_flip() +
  ggtitle("tamarind") +
  ylab("variable importance") +
  xlab("binary predictors") +
  # scale_color_manual(values=c("#1e824c", "#1e824c"))+
  scale_fill_manual(values = c("#a6915c", "#a6915c")) +
  # geom_col_pattern(aes(pattern_fill=phenology), pattern="stripe", fill="white", colour="#1e824c", position=position_dodge())
  geom_bar_pattern(
    stat = "identity", pattern_color = "white",
    pattern_fill = "black", position = position_dodge(), aes(pattern = phenology)
  )

library(gridExtra)

png("plots/figs/Fig3_new.png", width = 16, height = 8, units = "in", res = 300)
grid.arrange(jack_ml_ord, mango_ml_ord, tam_ml_ord,
  jack_ml_bin, mango_ml_bin, tam_ml_bin,
  ncol = 3
)
dev.off()
```

```{r Figure 3 12 panels and bars}
newlabs_ml <- c("precipitation", "minimum temperature", "mean temperature", "maximum temperature", "mean soil moisture", "solar radiation", "consecutive dry days", "elevation", "slope", "aspect", "urbanisation")

all_ml_ord$newlabs <- rep(newlabs_ml, 6)
all_ml_ord$newlabs <- factor(all_ml_ord$newlabs, levels = newlabs_ml)
all_ml_bin$newlabs <- rep(newlabs_ml, 6)
all_ml_bin$newlabs <- factor(all_ml_bin$newlabs, levels = newlabs_ml)

library(ggplot2)
jack_ml_ord_fl <- ggplot(all_ml_ord[all_ml_ord$sp == "jackfruit" & all_ml_ord$phenology == "flowering", ], aes(x = reorder(newlabs, Overall..average.), y = Overall..average.)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), fill = "#1e824c") +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd),
    position = position_dodge()
  ) +
  coord_flip() +
  ggtitle("Jackfruit flowering") +
  ylab("Variable importance") +
  xlab("Ordinal predictors") +
  labs(tag = "a)") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 12))

jack_ml_ord_fr <- ggplot(all_ml_ord[all_ml_ord$sp == "jackfruit" & all_ml_ord$phenology == "fruiting", ], aes(x = reorder(newlabs, Overall..average.), y = Overall..average.)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), fill = "#1e824c") +
  labs(tag = "d)") +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd),
    position = position_dodge()
  ) +
  coord_flip() +
  ggtitle("Jackfruit fruiting") +
  ylab("Variable importance") +
  xlab("Ordinal predictors") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 12))

mango_ml_ord_fl <- ggplot(all_ml_ord[all_ml_ord$sp == "mango" & all_ml_ord$phenology == "flowering", ], aes(x = reorder(newlabs, Overall..average.), y = Overall..average., shape = phenology)) +
  labs(tag = "b)") +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), fill = "#f7ca18") +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd),
    position = position_dodge()
  ) +
  coord_flip() +
  ggtitle("Mango flowering") +
  ylab("Variable importance") +
  xlab("Ordinal predictors") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 12))

mango_ml_ord_fr <- ggplot(all_ml_ord[all_ml_ord$sp == "mango" & all_ml_ord$phenology == "fruiting", ], aes(x = reorder(newlabs, Overall..average.), y = Overall..average., shape = phenology)) +
  labs(tag = "e)") +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), fill = "#f7ca18") +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd),
    position = position_dodge()
  ) +
  coord_flip() +
  ggtitle("Mango fruiting") +
  ylab("Variable importance") +
  xlab("Ordinal predictors") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 12))

tam_ml_ord_fl <- ggplot(all_ml_ord[all_ml_ord$sp == "tamarind" & all_ml_ord$phenology == "flowering", ], aes(x = reorder(newlabs, Overall..average.), y = Overall..average., shape = phenology)) +
  labs(tag = "c)") +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), fill = "#a6915c") +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd),
    position = position_dodge()
  ) +
  coord_flip() +
  ggtitle("Tamarind flowering") +
  ylab("Variable importance") +
  xlab("Ordinal predictors") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 12))

tam_ml_ord_fr <- ggplot(all_ml_ord[all_ml_ord$sp == "tamarind" & all_ml_ord$phenology == "fruiting", ], aes(x = reorder(newlabs, Overall..average.), y = Overall..average., shape = phenology)) +
  labs(tag = "f)") +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), fill = "#a6915c") +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd),
    position = position_dodge()
  ) +
  coord_flip() +
  ggtitle("Tamarind fruiting") +
  ylab("Variable importance") +
  xlab("Ordinal predictors") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 12))

# binary

jack_ml_bin_fl <- ggplot(all_ml_bin[all_ml_bin$sp == "jackfruit" & all_ml_bin$phenology == "flowering", ], aes(x = reorder(newlabs, Overall..average.), y = Overall..average.)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), fill = "#1e824c") +
  labs(tag = "g)") +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd),
    position = position_dodge()
  ) +
  coord_flip() +
  ggtitle("Jackfruit flowering") +
  ylab("Variable importance") +
  xlab("Binary predictors") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 12))

jack_ml_bin_fr <- ggplot(all_ml_bin[all_ml_bin$sp == "jackfruit" & all_ml_bin$phenology == "fruiting", ], aes(x = reorder(newlabs, Overall..average.), y = Overall..average.)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), fill = "#1e824c") +
  labs(tag = "j)") +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd),
    position = position_dodge()
  ) +
  coord_flip() +
  ggtitle("Jackfruit fruiting") +
  ylab("Variable importance") +
  xlab("Binary predictors") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 12))

mango_ml_bin_fl <- ggplot(all_ml_bin[all_ml_bin$sp == "mango" & all_ml_bin$phenology == "flowering", ], aes(x = reorder(newlabs, Overall..average.), y = Overall..average., shape = phenology)) +
  labs(tag = "h)") +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), fill = "#f7ca18") +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd),
    position = position_dodge()
  ) +
  coord_flip() +
  ggtitle("Mango flowering") +
  ylab("Variable importance") +
  xlab("Binary predictors") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 12))

mango_ml_bin_fr <- ggplot(all_ml_bin[all_ml_bin$sp == "mango" & all_ml_bin$phenology == "fruiting", ], aes(x = reorder(newlabs, Overall..average.), y = Overall..average., shape = phenology)) +
  labs(tag = "k)") +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), fill = "#f7ca18") +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd),
    position = position_dodge()
  ) +
  coord_flip() +
  ggtitle("Mango fruiting") +
  ylab("Variable importance") +
  xlab("Binary predictors") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 12))

tam_ml_bin_fl <- ggplot(all_ml_bin[all_ml_bin$sp == "tamarind" & all_ml_bin$phenology == "flowering", ], aes(x = reorder(newlabs, Overall..average.), y = Overall..average., shape = phenology)) +
  labs(tag = "i)") +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), fill = "#a6915c") +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd),
    position = position_dodge()
  ) +
  coord_flip() +
  ggtitle("Tamarind flowering") +
  ylab("Variable importance") +
  xlab("Binary predictors") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 12))

tam_ml_bin_fr <- ggplot(all_ml_bin[all_ml_bin$sp == "tamarind" & all_ml_bin$phenology == "fruiting", ], aes(x = reorder(newlabs, Overall..average.), y = Overall..average., shape = phenology)) +
  labs(tag = "l)") +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9), fill = "#a6915c") +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd),
    position = position_dodge()
  ) +
  coord_flip() +
  ggtitle("Tamarind fruiting") +
  ylab("Variable importance") +
  xlab("Binary predictors") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 12))

library(gridExtra)
png("plots/figs/Fig3_sep_bars_tag.png", width = 12, height = 18, units = "in", res = 300)
grid.arrange(jack_ml_ord_fl, mango_ml_ord_fl, tam_ml_ord_fl,
  jack_ml_ord_fr, mango_ml_ord_fr, tam_ml_ord_fr,
  jack_ml_bin_fl, mango_ml_bin_fl, tam_ml_bin_fl,
  jack_ml_bin_fr, mango_ml_bin_fr, tam_ml_bin_fr,
  ncol = 3
)
dev.off()
```

```{r Figure 3 like a coef plot}
# jack - #1e824c
# mango - #f7ca18
# tam - #a6915c

jack_ml_ord_cf <- ggplot(all_ml_ord[all_ml_ord$sp == "jackfruit", ], aes(x = X, y = Overall..average., shape = phenology)) +
  geom_point(position = position_dodge(width = 0.9), col = "#1e824c") +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd),
    position = position_dodge(), col = "#1e824c"
  ) +
  coord_flip() +
  ggtitle("jackfruit") +
  ylab("variable importance") +
  xlab("ordinal predictors") +
  theme_minimal()

mango_ml_ord_cf <- ggplot(all_ml_ord[all_ml_ord$sp == "mango", ], aes(x = X, y = Overall..average., shape = phenology)) +
  geom_point(position = position_dodge(width = 0.9), col = "#f7ca18") +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd),
    position = position_dodge(), col = "#f7ca18"
  ) +
  coord_flip() +
  ggtitle("mango") +
  ylab("variable importance") +
  xlab("ordinal predictors") +
  theme_minimal()

tam_ml_ord_cf <- ggplot(all_ml_ord[all_ml_ord$sp == "tamarind", ], aes(x = X, y = Overall..average., shape = phenology)) +
  geom_point(position = position_dodge(width = 0.9), col = "#a6915c") +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd),
    position = position_dodge(), col = "#a6915c"
  ) +
  coord_flip() +
  ggtitle("tamarind") +
  ylab("variable importance") +
  xlab("ordinal predictors") +
  theme_minimal()

jack_ml_bin_cf <- ggplot(all_ml_bin[all_ml_bin$sp == "jackfruit", ], aes(x = X, y = Overall..average., shape = phenology)) +
  geom_point(position = position_dodge(width = 0.9), col = "#1e824c") +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd),
    position = position_dodge(), col = "#1e824c"
  ) +
  coord_flip() +
  ggtitle("jackfruit") +
  ylab("variable importance") +
  xlab("binary predictors") +
  theme_minimal()

mango_ml_bin_cf <- ggplot(all_ml_bin[all_ml_bin$sp == "mango", ], aes(x = X, y = Overall..average., shape = phenology)) +
  geom_point(position = position_dodge(width = 0.9), col = "#f7ca18") +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd),
    position = position_dodge(), col = "#f7ca18"
  ) +
  coord_flip() +
  ggtitle("mango") +
  ylab("variable importance") +
  xlab("binary predictors") +
  theme_minimal()

tam_ml_bin_cf <- ggplot(all_ml_bin[all_ml_bin$sp == "tamarind", ], aes(x = X, y = Overall..average., shape = phenology)) +
  geom_point(position = position_dodge(width = 0.9), col = "#a6915c") +
  geom_errorbar(aes(ymin = Overall..average. - sd, ymax = Overall..average. + sd),
    position = position_dodge(), col = "#a6915c"
  ) +
  coord_flip() +
  ggtitle("tamarind") +
  ylab("variable importance") +
  xlab("binary predictors") +
  theme_minimal()

library(gridExtra)

png("plots/figs/Fig3_coefs.png", width = 16, height = 8, units = "in", res = 300)
grid.arrange(jack_ml_ord_cf, mango_ml_ord_cf, tam_ml_ord_cf,
  jack_ml_bin_cf, mango_ml_bin_cf, tam_ml_bin_cf,
  ncol = 3
)
dev.off()
```

```{r Figure 4}
# Figure 4: Interactions between effects from GLMMs (only significant) (3x3 table - for each species x interaction. Each row - one interaction that is significant - fruiting and flowering. Blank for not significant)

# find the mean, +/-SD of the static variables across all three species
# we need this for elevation, aspect and light

cont <- rep(seq(-3, 3, 0.05), 3)
stat <- c(rep(-1, 121), rep(0, 121), rep(1, 121))
cons <- rep(0, 363)

library(lme4)

# functionalising this plot

all_pars <- c(
  "FNSum_PCP_cs", "elev_cs", "FNMean_MaxTemp_cs", "FNMean_MeanSM_cs",
  "FNSum_DryDays_cs", "aspect_cs", "light_cs"
)

`%nin%` <- Negate(`%in%`)

labs_list <- list(
  "FNSum_PCP_cs" = "precipitation", "elev_cs" = "elevation",
  "FNMean_MaxTemp_cs" = "max temperature", "FNMean_MeanSM_cs" = "soil moisture",
  "FNSum_DryDays_cs" = "dry days", "aspect_cs" = "aspect",
  "light_cs" = "urbanisation", "User_Tree_id" = "Tree ID"
)

# parameters are model, temporal variable, static variable

temp_var <- c(rep("FNSum_PCP_cs", 6), rep("FNMean_MeanSM_cs", 6), rep("FNSum_DryDays_cs", 6))
static_var <- c(rep("elev_cs", 6), rep("aspect_cs", 6), rep("light_cs", 6))
static_name <- c(rep("elevation", 6), rep("aspect", 6), rep("urbanisation", 6))
model_list <- list(
  fl_inter_jack_bin, fl_inter_mango_bin, fl_inter_tam_bin,
  fr_inter_jack_bin, fr_inter_mango_bin, fr_inter_tam_bin
)
model_num <- rep(1:6, 3)
pheno_var <- rep(c(rep("flowering", 3), rep("fruiting", 3)), 3)
alph <- c(
  NA, NA, NA, "a", "b", "c", "d", "e", "f", "g", "h",
  "i", "j", "k", "l", "m", "n", "o"
)

library(lme4)
library(ggplot2)
int_plot <- function(model, temp, static, static_new, pheno, alph) {
  cons_pars <- all_pars[which(all_pars %nin% c(temp, static))]
  newdat <- data.frame(cont, stat, cons, cons, cons, cons, cons, NA)
  colnames(newdat) <- c(temp, static, cons_pars, "User_Tree_id")
  # newdat$pred<-predict(model, allow.new.levels=T, newdata=newdat,  level=0)
  newdat$pred <- predict(model, allow.new.levels = T, newdata = newdat, type = "response", re.form = NA)

  # need to add confidence intervals to plot

  newdat <- newdat[newdat$pred >= 0 & newdat$pred <= 1, ] # making sure that there are no impossible probs
  newdat[, 2] <- factor(newdat[, 2])
  temp1 <- sym(temp)
  static1 <- sym(static)
  p <- ggplot(newdat, aes(x = !!temp1, y = pred, color = !!static1)) +
    geom_line(size = 2) +
    scale_color_manual(values = c("firebrick4", "grey60", "navy")) +
    ylab(paste(pheno, " probability", sep = "")) +
    xlab("") +
    labs(color = static_new, tag = alph) +
    theme_bw() +
    theme(
      axis.title.x = element_text(size = 16),
      axis.title.y = element_text(size = 16)
    )
  return(p)
}


plotlist <- mapply(function(x) {
  int_plot(
    model = model_list[[model_num[x]]], temp = temp_var[x],
    static = static_var[x], pheno = pheno_var[x]
  )
}, x = 1:18)

x <- 1
p1 <- ggplot() +
  geom_blank() +
  theme_bw() +
  ggtitle("jackfruit") +
  guides(colour = "none") +
  ylab("flowering probability") +
  xlab("")
x <- 2
p2 <- ggplot() +
  geom_blank() +
  ggtitle("mango") +
  xlab("precipitation") +
  ylab("") +
  theme_bw()

x <- 3
p3 <- ggplot() +
  geom_blank() +
  ggtitle("tamarind") +
  xlab("") +
  ylab("") +
  theme_bw()

x <- 4
p4 <- int_plot(
  model = model_list[[model_num[x]]], temp = temp_var[x],
  static = static_var[x], static_new = static_name[x], pheno = pheno_var[x], alph = alph[x]
) +
  guides(colour = "none")

x <- 5
p5 <- int_plot(
  model = model_list[[model_num[x]]], temp = temp_var[x],
  static = static_var[x], static_new = static_name[x], pheno = pheno_var[x], alph = alph[x]
) +
  guides(colour = "none") + xlab("precipitation") + ylab("")

x <- 6 # this needs to be blank
p6 <- int_plot(
  model = model_list[[model_num[x]]], temp = temp_var[x],
  static = static_var[x], static_new = static_name[x], pheno = pheno_var[x], alph = alph[x]
) +
  geom_rect(aes(xmin = -3, xmax = 3, ymin = 0, ymax = 2),
    fill = "grey100", color = "grey100"
  ) + ylab("") + theme_void() + xlab("")

# p6<-ggplot()+geom_blank()+guides(fill=guide_legend(override.aes = list(alpha=1)))

x <- 7
p7 <- int_plot(
  model = model_list[[model_num[x]]], temp = temp_var[x],
  static = static_var[x], static_new = static_name[x], pheno = pheno_var[x], alph = alph[x]
) +
  guides(colour = "none")
x <- 8
p8 <- ggplot() +
  geom_blank() +
  xlab("soil moisture") +
  ylab("") +
  guides(colour = "none") +
  geom_text(x = 0.5, y = 0.5, label = "n.s.") +
  theme_bw() +
  theme(
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16)
  )
x <- 9
p9 <- int_plot(
  model = model_list[[model_num[x]]], temp = temp_var[x],
  static = static_var[x], static_new = static_name[x], alph = alph[x], pheno = pheno_var[x]
) + ylab("")
x <- 10
p10 <- ggplot() +
  geom_blank() +
  ylab("fruiting probability") +
  xlab("") +
  theme_bw() +
  theme(
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16)
  )

x <- 11
p11 <- int_plot(
  model = model_list[[model_num[x]]], temp = temp_var[x],
  static = static_var[x], static_new = static_name[x], pheno = pheno_var[x], alph = alph[x]
) +
  guides(colour = "none") + xlab("soil moisture") + ylab("")
x <- 12
p12 <- int_plot(
  model = model_list[[model_num[x]]], temp = temp_var[x],
  static = static_var[x], static_new = static_name[x], alph = alph[x], pheno = pheno_var[x]
) + ylab("")
x <- 13
p13 <- int_plot(
  model = model_list[[model_num[x]]], temp = temp_var[x], alph = alph[x],
  static = static_var[x], static_new = static_name[x], pheno = pheno_var[x]
) + guides(colour = "none")
x <- 14
p14 <- ggplot() +
  geom_blank() +
  xlab("consecutive dry days") +
  ylab("") +
  theme_bw() +
  theme(
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16)
  )
x <- 15
p15 <- int_plot(
  model = model_list[[model_num[x]]], temp = temp_var[x], alph = alph[x],
  static = static_var[x], static_new = static_name[x], pheno = pheno_var[x]
) + ylab("")
x <- 16
p16 <- int_plot(
  model = model_list[[model_num[x]]], temp = temp_var[x],
  static = static_var[x], static_new = static_name[x], alph = alph[x], pheno = pheno_var[x]
) + guides(colour = "none")
x <- 17
p17 <- int_plot(
  model = model_list[[model_num[x]]], temp = temp_var[x], alph = alph[x],
  static = static_var[x], static_new = static_name[x], pheno = pheno_var[x]
) +
  guides(colour = "none") + xlab("consecutive dry days") + ylab("")
x <- 18
p18 <- int_plot(
  model = model_list[[model_num[x]]], temp = temp_var[x], alph = alph[x],
  static = static_var[x], static_new = static_name[x], pheno = pheno_var[x]
) + ylab("")


library(gridExtra)
png("plots/figs/Fig4_new.png", width = 12, height = 16, units = "in", res = 300)
grid.arrange(p1, p2, p3, p4, p5, p6, p7, p8, p9,
  p10, p11, p12, p13, p14, p15, p16, p17, p18,
  ncol = 3, widths = c(10, 10, 12.5)
)

dev.off()

png("plots/figs/Fig4_noline1.png", width = 12, height = 16, units = "in", res = 300)
grid.arrange(p4, p5, p6, p7, p8, p9,
  p10, p11, p12, p13, p14, p15, p16, p17, p18,
  ncol = 3, widths = c(10, 10, 12.5)
)

dev.off()
```

for(i in 1:2){
  x<-i
  px<-"x"
  assign(px, paste0("p", x, sep=""))
  #px<-(paste("p", x, sep=""))
  eval(as.symbol(px))<-int_plot(model=model_list[[model_num[x]]], temp=temp_var[x],
                            static=static_var[x], pheno=pheno_var[x])
}

jack_fr_elev_pred<-predict(fr_inter_jack_bin, allow.new.levels=T,
                           #terms=c("FNSum_PCP_cs", "elev_cs"), 
                           newdata=data.frame("FNSum_PCP_cs"=cont, "elev_cs"=stat,
                                              "FNMean_MaxTemp_cs"=cons, "FNMean_MeanSM_cs"= cons,
                                              "FNSum_DryDays_cs"=cons, "aspect_cs"=cons, "light_cs"=cons,
                                              "User_Tree_id"=mean(sw_spat_rescale$User_Tree_id)))

ggplot()+geom_line(aes(x=cont, y=jack_fr_elev_pred, color=factor(stat)), size=2)

library(sjPlot)
#jack_fl_elev<-plot_model(fl_inter_jack_bin, terms=c("FNSum_PCP_cs", "elev_cs"), type="pred")+ggtitle("jackfruit")
jack_fr_elev<-plot_model(fr_inter_jack_bin, terms=c("FNSum_PCP_cs", "elev_cs"), type="pred")+ggtitle("")
jack_fl_aspect<-plot_model(fl_inter_jack_bin, terms=c("FNMean_MeanSM_cs", "aspect_cs"), type="pred")+ggtitle("")
#jack_fr_aspect (gap)

jack_fl_light<-plot_model(fl_inter_jack_bin, terms=c("FNSum_DryDays_cs", "light_cs"), type="pred", mdrt.values = "minmax")+ggtitle("jackfruit")+ylab("flowering probability")

jack_fr_light<-plot_model(fr_inter_jack_bin, terms=c("FNSum_DryDays_cs", "light_cs"), type="pred")+ggtitle("")

#mango_fl_elev<-plot_model(fl_inter_mango_bin, terms=c("FNSum_PCP_cs", "elev_cs"), type="pred")+ggtitle("mango")
mango_fr_elev<-plot_model(fr_inter_mango_bin, terms=c("FNSum_PCP_cs", "elev_cs"), type="pred")+ggtitle("")
#mango_fl_aspect<-plot_model(fl_inter_mango_bin, terms=c("FNMean_MeanSM_cs", "aspect_cs"), type="pred")+ggtitle("")
mango_fr_aspect<-plot_model(fr_inter_mango_bin, terms=c("FNMean_MeanSM_cs", "aspect_cs"), type="pred")+ggtitle("")
#mango_fl_light<-plot_model(fl_inter_mango_bin, terms=c("FNSum_DryDays_cs", "light_cs"), type="pred")+ggtitle("")
mango_fr_light<-plot_model(fr_inter_mango_bin, terms=c("FNSum_DryDays_cs", "light_cs"), type="pred")+ggtitle("")


#tam_fl_elev (gap)
#tam_fr_elev (gap)
tam_fl_aspect<-plot_model(fl_inter_tam_bin, terms=c("FNMean_MeanSM_cs", "aspect_cs"), type="pred")+ggtitle("")
tam_fr_aspect<-plot_model(fr_inter_tam_bin, terms=c("FNMean_MeanSM_cs", "aspect_cs"), type="pred")+ggtitle("")
tam_fl_light<-plot_model(fl_inter_tam_bin, terms=c("FNSum_DryDays_cs", "light_cs"), type="pred")+ggtitle("tamarind")
tam_fr_light<-plot_model(fr_inter_tam_bin, terms=c("FNSum_DryDays_cs", "light_cs"), type="pred")+ggtitle("")

#png("plots/figs/Fig4_fl.png", width=12, height =12, units="in", res=300)
png("plots/figs/Fig4.png", width=12, height =16, units="in", res=300)
(jack_fl_light+plot_spacer()+tam_fl_light)/(jack_fr_light+mango_fr_light+tam_fr_light)/(plot_spacer()+plot_spacer()+plot_spacer())/(jack_fr_elev+mango_fr_elev+plot_spacer())/(jack_fl_aspect+plot_spacer()+tam_fl_aspect)/(plot_spacer()+mango_fr_aspect+tam_fr_aspect)
dev.off()

plotlist[[3]]

p7


```{r Figure 5}
# Figure 5: hotspot analysis from Carina - of environmental variables - where do we need more observations from - data gaps
```

```{r Table S2 - User ID random effect}
fl_jack_bin_usr2 <- readRDS("models/fl_jack_bin_usr.rds")
fl_mango_bin_usr2 <- readRDS("models/fl_mango_bin_usr.rds")
fl_tam_bin_usr2 <- readRDS("models/fl_tam_bin_usr.rds")

library(sjPlot)
tab_model(fl_jack_bin_usr2, fl_mango_bin_usr2, fl_tam_bin_usr2,
  transform = NULL, auto.label = FALSE, collapse.ci = TRUE,
  file = "models/fl_all_usr.xls"
)

fr_jack_bin_usr2 <- readRDS("models/fr_jack_bin_usr.rds")
fr_mango_bin_usr2 <- readRDS("models/fr_mango_bin_usr.rds")
fr_tam_bin_usr2 <- readRDS("models/fr_tam_bin_usr.rds")

tab_model(fr_jack_bin_usr2, fr_mango_bin_usr2, fr_tam_bin_usr2,
  transform = NULL, auto.label = FALSE, collapse.ci = TRUE,
  file = "models/fr_all_usr.xls"
)
```
```{r Table S3 monthly predictors}
fl_jack_bin_mn <- readRDS("models/fl_jack_bin_mn.rds")
fl_mango_bin_mn <- readRDS("models/fl_mango_bin_mn.rds")
fl_tam_bin_mn <- readRDS("models/fl_tam_bin_mn.rds")

library(sjPlot)
tab_model(fl_jack_bin_mn, fl_mango_bin_mn, fl_tam_bin_mn,
  transform = NULL, auto.label = FALSE, collapse.ci = TRUE,
  file = "models/fl_all_mn.xls"
)

fr_jack_bin_mn <- readRDS("models/fr_jack_bin_mn.rds")
fr_mango_bin_mn <- readRDS("models/fr_mango_bin_mn.rds")
fr_tam_bin_mn <- readRDS("models/fr_tam_bin_mn.rds")

tab_model(fr_jack_bin_mn, fr_mango_bin_mn, fr_tam_bin_mn,
  transform = NULL, auto.label = FALSE, collapse.ci = TRUE,
  file = "models/fr_all_mn.xls"
)
```
```{r Table S4 loc time}
fl_inter_jack_bin_loc_time <- readRDS("models/fl_inter_jack_bin_loc_time.rds")
fl_inter_mango_bin_loc_time <- readRDS("models/fl_inter_mango_bin_loc_time.rds")
fl_inter_tam_bin_loc_time <- readRDS("models/fl_inter_tam_bin_loc_time.rds")

fr_inter_jack_bin_loc_time <- readRDS("models/fr_inter_jack_bin_loc_time.rds")
fr_inter_mango_bin_loc_time <- readRDS("models/fr_inter_mango_bin_loc_time.rds")
fr_inter_tam_bin_loc_time <- readRDS("models/fr_inter_tam_bin_loc_time.rds")

library(lme4)
library(performance)

png("plots/fl_inter_jack_bin_loc_time_diag.png", width = 8, height = 12, units = "in", res = 300)
check_model(fl_inter_jack_bin_loc_time)
dev.off()

png("plots/fl_inter_mango_bin_loc_time_diag.png", width = 8, height = 12, units = "in", res = 300)
check_model(fl_inter_mango_bin_loc_time)
dev.off()

png("plots/fl_inter_tam_bin_loc_time_diag.png", width = 8, height = 12, units = "in", res = 300)
check_model(fl_inter_tam_bin_loc_time)
dev.off()

png("plots/fr_inter_jack_bin_loc_time_diag.png", width = 8, height = 12, units = "in", res = 300)
check_model(fr_inter_jack_bin_loc_time)
dev.off()

png("plots/fr_inter_mango_bin_loc_time_diag.png", width = 8, height = 12, units = "in", res = 300)
check_model(fr_inter_mango_bin_loc_time)
dev.off()

png("plots/fr_inter_tam_bin_loc_time_diag.png", width = 8, height = 12, units = "in", res = 300)
check_model(fr_inter_tam_bin_loc_time)
dev.off()


library(sjPlot)
tab_model(fl_inter_jack_bin_loc_time, fl_inter_mango_bin_loc_time, fl_inter_tam_bin_loc_time,
  transform = NULL, auto.label = FALSE, collapse.ci = TRUE,
  file = "models/fl_all_loc_time.xls"
)

fr_jack_bin_mn <- readRDS("models/fr_jack_bin_mn.rds")
fr_mango_bin_mn <- readRDS("models/fr_mango_bin_mn.rds")
fr_tam_bin_mn <- readRDS("models/fr_tam_bin_mn.rds")

tab_model(fr_inter_jack_bin_loc_time, fr_inter_mango_bin_loc_time, fr_inter_tam_bin_loc_time,
  transform = NULL, auto.label = FALSE, collapse.ci = TRUE,
  file = "models/fr_all_loc_time.xls"
)
```
