---
title: "03_ml_analysis"
author: "Krishna Anujan"
#date: "2024-04-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r functions}
# F1 scores------------------------

f1 <- function(pred, obs) {
  acc <- rep(NA, length(obs))
  acc[which(pred == 1 & obs == 1)] <- "TP" # true positive
  acc[which(pred == 0 & obs == 0)] <- "TN" # true negative
  acc[which(pred == 0 & obs == 1)] <- "FN" # false negative
  acc[which(pred == 1 & obs == 0)] <- "FP" # false positive

  # F1 score is 2*(precision*recall)/(precision+recall)
  # precision = TP/(TP+FP)
  # recall = TP/ (TP+FN)

  precision <- table(acc)[4] / (table(acc)[4] + table(acc)[2])
  recall <- table(acc)[4] / (table(acc)[4] + table(acc)[1])

  f1 <- 2 * (precision * recall) / (precision + recall)

  return(f1)
}
```

```{r ml}
library(randomForestSRC)
library(dplyr)
library(randomForest)
# install.packages("ggRandomForests")
# library(ggRandomForests)
library(caret)

rm(list = ls())

sw_spat <- readRDS("data/modified_data/sw_spat.RDS")

# jack<-sw_spat %>%
#  filter(diffdays==7,
#         Species_name=="Artocarpus heterophyllus")

jack <- sw_spat %>%
  filter(Species_name == "Artocarpus heterophyllus")

# jack

library(tidyverse)

jack_ml <- jack %>%
  # dplyr::select(grep("FN", colnames(jack), value=T), fl_binary, year, fl_binary_prev)
  dplyr::select(grep("FN", colnames(jack), value = T), fl_binary, year)
jack_ml <- jack_ml %>%
  # dplyr::select(!grep("cs", colnames(jack_ml), value=T), fl_binary, year) %>%
  mutate(
    mean_temp = scale(FNMean_MeanTemp),
    min_temp = scale(FNMean_MinTemp),
    max_temp = scale(FNMean_MaxTemp),
    solar_rad = scale(FNSum_SolarRad),
    precip = scale(FNSum_PCP),
    dry_days = scale(FNSum_ConsDryDays),
    soil_moisture = scale(FNMean_MeanSM)
  ) %>%
  dplyr::select(!grep("FN", colnames(jack_ml), value = T))



# loop for LOOCV

preds <- data.frame(NA, NA, NA)
colnames(preds) <- c("obs", "pred", "year")
preds <- preds[F, ]

vimp <- data.frame(var = colnames(dplyr::select(jack_ml, -c("fl_binary", "year"))))

for (i in 1:length(unique(jack_ml$year))) {
  yr <- unique(jack_ml$year)[i]
  # subset
  training <- dplyr::select(jack_ml[jack_ml$year != yr, ], -"year")
  testing <- dplyr::select(jack_ml[jack_ml$year == yr, ], -"year")

  # prepping for analysis
  training <- training[complete.cases(training) == T, ]
  testing <- testing[complete.cases(testing) == T, ]
  training$fl_binary <- factor(as.character(training$fl_binary), levels = c("0", "1"))
  testing$fl_binary <- factor(as.character(testing$fl_binary), levels = c("0", "1"))

  names(training) <- make.names(names(training))
  rf_jack <- randomForest::randomForest(fl_binary ~ ., importance = T, data = training)


  predictions.rf <- predict(rf_jack, newdata = testing)
  pred.df <- data.frame(obs = testing[, "fl_binary"], pred = predictions.rf, year = rep(yr, year = length(predictions.rf)))

  preds <- rbind(preds, pred.df)


  # vimp_mean<-caret::varImp(rf_jack)
  vimp_mean <- data.frame(importance(rf_jack, type = 1))
  vimp <- cbind(vimp, vimp_mean[, 1])
  colnames(vimp)[ncol(vimp)] <- as.character(yr)
}

write.csv(vimp, "models/ML/vimp_jack_noyr.csv", row.names = F)
write.csv(preds, "models/ML/preds_jack_noyr.csv", row.names = F)
# write.csv(vimp, "models/ML/vimp_jack_withprev.csv", row.names = F)
# write.csv(preds, "models/ML/preds_jack_withprev.csv", row.names = F)

preds <- read.csv("models/ML/preds_jack_noyr.csv")

confusionMatrix(preds$pred, preds$obs)
preds$obs.num <- as.numeric(as.character(preds$obs))
preds$pred.num <- as.numeric(as.character(preds$pred))
R2_jack <- 1 - (sum((preds$obs.num - preds$pred.num)^2) / sum((preds$obs.num - mean(preds$obs.num))^2))

F1_jack <- f1(pred = preds$pred.num, obs = preds$obs.num)


library(dplyr)
vimp_jack <- vimp
# vimp_jack<-read.csv("models/ML/vimp_jack.csv")
vimp_jack <- read.csv("models/ML/vimp_jack_noyr.csv")
vimp_jack$mean <- rowMeans(vimp_jack[, 2:ncol(vimp_jack)])
vimp_jack$sd <- apply(vimp_jack[, 2:(ncol(vimp_jack) - 1)], 1, sd)
vimp_jack$sd_1.6 <- vimp_jack$sd * 1.6
```

```{r ml mango}
mango_ml <- mango %>%
  dplyr::select(grep("FN", colnames(mango), value = T), fl_binary, year)
mango_ml <- mango_ml %>%
  # dplyr::select(!grep("cs", colnames(mango_ml), value=T))%>%
  mutate(
    mean_temp = scale(FNMean_MeanTemp),
    min_temp = scale(FNMean_MinTemp),
    max_temp = scale(FNMean_MaxTemp),
    solar_rad = scale(FNSum_SolarRad),
    precip = scale(FNSum_PCP),
    dry_days = scale(FNSum_ConsDryDays),
    soil_moisture = scale(FNMean_MeanSM)
  ) %>%
  dplyr::select(!grep("FN", colnames(mango_ml), value = T))


# loop for LOOCV

preds <- data.frame(NA, NA, NA)
colnames(preds) <- c("obs", "pred", "year")
preds <- preds[F, ]

vimp <- data.frame(var = colnames(dplyr::select(mango_ml, -c("fl_binary", "year"))))

for (i in 1:length(unique(mango_ml$year))) {
  yr <- unique(mango_ml$year)[i]
  # subset
  training <- dplyr::select(mango_ml[mango_ml$year != yr, ], -"year")
  testing <- dplyr::select(mango_ml[mango_ml$year == yr, ], -"year")

  # prepping for analysis
  training <- training[complete.cases(training) == T, ]
  testing <- testing[complete.cases(testing) == T, ]
  training$fl_binary <- factor(as.character(training$fl_binary), levels = c("0", "1"))
  testing$fl_binary <- factor(as.character(testing$fl_binary), levels = c("0", "1"))

  names(training) <- make.names(names(training))
  rf_mango <- randomForest::randomForest(fl_binary ~ ., importance = T, data = training)


  predictions.rf <- predict(rf_mango, newdata = testing)
  pred.df <- data.frame(obs = testing[, "fl_binary"], pred = predictions.rf, year = rep(yr, year = length(predictions.rf)))

  preds <- rbind(preds, pred.df)


  # vimp_mean<-caret::varImp(rf_mango)
  vimp_mean <- data.frame(importance(rf_mango, type = 1))
  vimp <- cbind(vimp, vimp_mean[, 1])
  colnames(vimp)[ncol(vimp)] <- as.character(yr)
}

write.csv(vimp, "models/ML/vimp_mango_noyr.csv", row.names = F)
write.csv(preds, "models/ML/preds_mango_noyr.csv", row.names = F)

preds <- read.csv("models/ML/preds_mango_noyr.csv")


confusionMatrix(preds$pred, preds$obs)
preds$obs.num <- as.numeric(as.character(preds$obs))
preds$pred.num <- as.numeric(as.character(preds$pred))
R2_mango <- 1 - (sum((preds$obs.num - preds$pred.num)^2) / sum((preds$obs.num - mean(preds$obs.num))^2))

F1_mango <- f1(pred = preds$pred.num, obs = preds$obs.num)


library(dplyr)
vimp_mango <- read.csv("models/ML/vimp_mango_noyr.csv")
vimp_mango$mean <- rowMeans(vimp_mango[, 2:ncol(vimp_mango)])
vimp_mango$sd <- apply(vimp_mango[, 2:(ncol(vimp_mango) - 1)], 1, sd)
vimp_mango$sd_1.6 <- vimp_mango$sd * 1.6
```
```{r ml tamarind}
tam_ml <- tam %>%
  dplyr::select(grep("FN", colnames(tam), value = T), fl_binary, year)
tam_ml <- tam_ml %>%
  # dplyr::select(grep("cs", colnames(tam_ml), value=T), fl_binary, year)%>%
  mutate(
    mean_temp = scale(FNMean_MeanTemp),
    min_temp = scale(FNMean_MinTemp),
    max_temp = scale(FNMean_MaxTemp),
    solar_rad = scale(FNSum_SolarRad),
    precip = scale(FNSum_PCP),
    dry_days = scale(FNSum_ConsDryDays),
    soil_moisture = scale(FNMean_MeanSM)
  ) %>%
  dplyr::select(!grep("FN", colnames(tam_ml), value = T))


# loop for LOOCV

preds <- data.frame(NA, NA, NA)
colnames(preds) <- c("obs", "pred", "year")
preds <- preds[F, ]

vimp <- data.frame(var = colnames(dplyr::select(tam_ml, -c("fl_binary", "year"))))

for (i in 1:length(unique(tam_ml$year))) {
  yr <- unique(tam_ml$year)[i]
  # subset
  training <- dplyr::select(tam_ml[tam_ml$year != yr, ], -"year")
  testing <- dplyr::select(tam_ml[tam_ml$year == yr, ], -"year")

  # prepping for analysis
  training <- training[complete.cases(training) == T, ]
  testing <- testing[complete.cases(testing) == T, ]
  training$fl_binary <- factor(as.character(training$fl_binary), levels = c("0", "1"))
  testing$fl_binary <- factor(as.character(testing$fl_binary), levels = c("0", "1"))

  names(training) <- make.names(names(training))
  rf_tam <- randomForest::randomForest(fl_binary ~ ., importance = T, data = training)


  predictions.rf <- predict(rf_tam, newdata = testing)
  pred.df <- data.frame(obs = testing[, "fl_binary"], pred = predictions.rf, year = rep(yr, year = length(predictions.rf)))

  preds <- rbind(preds, pred.df)


  # vimp_mean<-caret::varImp(rf_tam)
  vimp_mean <- data.frame(importance(rf_tam, type = 1))
  vimp <- cbind(vimp, vimp_mean[, 1])
  colnames(vimp)[ncol(vimp)] <- as.character(yr)
}

write.csv(vimp, "models/ML/vimp_tam_noyr.csv", row.names = F)
write.csv(preds, "models/ML/preds_tam_noyr.csv", row.names = F)

preds <- read.csv("models/ML/preds_tam_noyr.csv")

confusionMatrix(preds$pred, preds$obs)
preds$obs.num <- as.numeric(as.character(preds$obs))
preds$pred.num <- as.numeric(as.character(preds$pred))
R2_tam <- 1 - (sum((preds$obs.num - preds$pred.num)^2) / sum((preds$obs.num - mean(preds$obs.num))^2))

F1_tam <- f1(pred = preds$pred.num, obs = preds$obs.num)


library(dplyr)
vimp_tam <- read.csv("models/ML/vimp_tam_noyr.csv")
vimp_tam$mean <- rowMeans(vimp_tam[, 2:ncol(vimp_tam)])
vimp_tam$sd <- apply(vimp_tam[, 2:(ncol(vimp_tam) - 1)], 1, sd)
vimp_tam$sd_1.6 <- vimp_tam$sd * 1.6
```

```{r plots from random forest}
labs_list <- list(
  "FNSum_PCP_cs" = "precipitation",
  "elev_cs" = "elevation",
  "FNMean_MaxTemp_cs" = "max temperature",
  "FNMean_MeanSM_cs" = "soil moisture",
  "FNSum_DryDays_cs" = "dry days",
  "aspect_cs" = "aspect",
  "light_cs" = "urbanisation",
  "User_Tree_id" = "Tree ID",
  "FNMean_MinTemp_cs" = "minimum temperature",
  "FNMean_MeanTemp_cs" = "mean temperature",
  "FNSum_SolarRad_cs" = "solar radiation"
)

labs_list2 <- list(
  "precip" = "precipitation",
  "max_temp" = "max temperature",
  "soil_moisture" = "soil moisture",
  "dry_days" = "dry days",
  "min_temp" = "minimum temperature",
  "mean_temp" = "mean temperature",
  "solar_rad" = "solar radiation"
)

library(ggplot2)

# fmt <- "R^2 == %.3f"
fmt2 <- "F1 == %.3f"
lab2 <- sprintf(fmt2, F1_jack)

vimp_jack_plot <- ggplot(vimp_jack, aes(x = reorder(var, mean), weight = mean)) +
  geom_bar(fill = "#1e824c") +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0, size = 1) +
  geom_errorbar(aes(ymin = mean - sd_1.6, ymax = mean + sd_1.6), width = 0, size = 0.5) +
  ylab("Variable Importance") +
  xlab("") +
  scale_x_discrete(labels = labs_list2) +
  guides(fill = F) +
  labs(tag = "a)") +
  ggtitle("Jackfruit flowering probability") +
  annotate("text",
    x = 1, y = 100, label = lab2,
    parse = TRUE, hjust = 0
  ) +
  coord_flip() +
  theme_bw()
vimp_jack_plot

lab2 <- sprintf(fmt2, F1_mango)

vimp_mango_plot <- ggplot(vimp_mango, aes(x = reorder(var, mean), weight = mean)) +
  geom_bar(fill = "#f7ca18") +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0, size = 1) +
  geom_errorbar(aes(ymin = mean - sd_1.6, ymax = mean + sd_1.6), width = 0, size = 0.5) +
  ylab("Variable Importance") +
  xlab("") +
  scale_x_discrete(labels = labs_list2) +
  guides(fill = F) +
  labs(tag = "b)") +
  ggtitle("Mango flowering probability") +
  annotate("text",
    x = 1, y = 100, label = lab2,
    parse = TRUE, hjust = 0
  ) +
  coord_flip() +
  theme_bw()
vimp_mango_plot

lab2 <- sprintf(fmt2, F1_tam)

vimp_tam_plot <- ggplot(vimp_tam, aes(x = reorder(var, mean), weight = mean)) +
  geom_bar(fill = "#a6915c") +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0, size = 1) +
  geom_errorbar(aes(ymin = mean - sd_1.6, ymax = mean + sd_1.6), width = 0, size = 0.5) +
  ylab("Variable Importance") +
  xlab("") +
  scale_x_discrete(labels = labs_list2) +
  guides(fill = F) +
  labs(tag = "c)") +
  ggtitle("Tamarind flowering probability") +
  annotate("text",
    x = 1, y = 100, label = lab2,
    parse = TRUE, hjust = 0
  ) +
  coord_flip() +
  theme_bw()
vimp_tam_plot

library(gridExtra)
library(patchwork)

# png("doc/display/Fig2.png", width=12, height=4, units="in", res=300)
png("doc/display/Fig2_noyr.png", width = 12, height = 4, units = "in", res = 300)
grid.arrange(vimp_jack_plot, vimp_mango_plot, vimp_tam_plot, nrow = 1)
# vimp_mean_plot+vimp_invcv_plot
dev.off()
```
```{r confusion matrix}
# jack

preds.jack <- read.csv("models/ML/preds_jack.csv")
cm_jack <- confusionMatrix(factor(preds.jack$pred, levels = c("0", "1")), factor(preds.jack$obs, levels = c("0", "1")), dnn = c("Prediction", "Reference"))

plt <- as.data.frame(cm_jack$table)
plt$Prediction <- factor(plt$Prediction, levels = rev(levels(plt$Prediction)))

cm_plot_jack <- ggplot(plt, aes(Prediction, Reference, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq)) +
  scale_fill_gradient(low = "white", high = "#009194") +
  labs(x = "Reference", y = "Prediction") +
  ggtitle("jackfruit flowering ML model") +
  theme_minimal()
cm_plot_jack

# mango
preds.mango <- read.csv("models/ML/preds_mango.csv")
cm_mango <- confusionMatrix(factor(preds.mango$pred, levels = c("0", "1")), factor(preds.mango$obs, levels = c("0", "1")), dnn = c("Prediction", "Reference"))

plt <- as.data.frame(cm_mango$table)
plt$Prediction <- factor(plt$Prediction, levels = rev(levels(plt$Prediction)))

cm_plot_mango <- ggplot(plt, aes(Prediction, Reference, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq)) +
  scale_fill_gradient(low = "white", high = "#009194") +
  labs(x = "Reference", y = "Prediction") +
  ggtitle("mango flowering ML model") +
  theme_minimal()
cm_plot_mango


# tam
preds.tam <- read.csv("models/ML/preds_tam.csv")
cm_tam <- confusionMatrix(factor(preds.tam$pred, levels = c("0", "1")), factor(preds.tam$obs, levels = c("0", "1")), dnn = c("Prediction", "Reference"))

plt <- as.data.frame(cm_tam$table)
plt$Prediction <- factor(plt$Prediction, levels = rev(levels(plt$Prediction)))

cm_plot_tam <- ggplot(plt, aes(Prediction, Reference, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq)) +
  scale_fill_gradient(low = "white", high = "#009194") +
  labs(x = "Reference", y = "Prediction") +
  ggtitle("tamarind flowering ML model") +
  theme_minimal()
cm_plot_tam

png("doc/display/SFig_confusion.png", width = 12, height = 4, units = "in", res = 300)
grid.arrange(cm_plot_jack, cm_plot_mango, cm_plot_tam, nrow = 1)
# vimp_mean_plot+vimp_invcv_plot
dev.off()
```

```{r repeat with onset data}
library(randomForestSRC)
library(dplyr)
library(randomForest)
# install.packages("ggRandomForests")
# library(ggRandomForests)
library(caret)

rm(list = ls())

start_nomiss <- readRDS("data/modified_data/start_nomiss.RDS")


# jack<-sw_spat %>%
#  filter(diffdays==7,
#         Species_name=="Artocarpus heterophyllus")

jack <- start_nomiss %>%
  filter(Species_name == "Artocarpus heterophyllus")

# jack

library(tidyverse)

jack_ml <- jack %>%
  # dplyr::select(grep("FN", colnames(jack), value=T), fl_binary, year, fl_binary_prev)
  dplyr::select(grep("FN", colnames(jack), value = T), fl_binary, year)
jack_ml <- jack_ml %>%
  # dplyr::select(!grep("cs", colnames(jack_ml), value=T), fl_binary, year) %>%
  mutate(
    mean_temp = scale(FNMean_MeanTemp),
    min_temp = scale(FNMean_MinTemp),
    max_temp = scale(FNMean_MaxTemp),
    solar_rad = scale(FNSum_SolarRad),
    precip = scale(FNSum_PCP),
    dry_days = scale(FNSum_ConsDryDays),
    soil_moisture = scale(FNMean_MeanSM)
  ) %>%
  dplyr::select(!grep("FN", colnames(jack_ml), value = T))



# loop for LOOCV

preds <- data.frame(NA, NA, NA)
colnames(preds) <- c("obs", "pred", "year")
preds <- preds[F, ]

vimp <- data.frame(var = colnames(dplyr::select(jack_ml, -c("fl_binary", "year"))))

for (i in 1:length(unique(jack_ml$year))) {
  yr <- unique(jack_ml$year)[i]
  # subset
  training <- dplyr::select(jack_ml[jack_ml$year != yr, ], -"year")
  testing <- dplyr::select(jack_ml[jack_ml$year == yr, ], -"year")

  # prepping for analysis
  training <- training[complete.cases(training) == T, ]
  testing <- testing[complete.cases(testing) == T, ]
  training$fl_binary <- factor(as.character(training$fl_binary), levels = c("0", "1"))
  testing$fl_binary <- factor(as.character(testing$fl_binary), levels = c("0", "1"))

  names(training) <- make.names(names(training))
  rf_jack <- randomForest::randomForest(fl_binary ~ ., importance = T, data = training)


  predictions.rf <- predict(rf_jack, newdata = testing)
  pred.df <- data.frame(obs = testing[, "fl_binary"], pred = predictions.rf, year = rep(yr, year = length(predictions.rf)))

  preds <- rbind(preds, pred.df)


  # vimp_mean<-caret::varImp(rf_jack)
  vimp_mean <- data.frame(importance(rf_jack, type = 1))
  vimp <- cbind(vimp, vimp_mean[, 1])
  colnames(vimp)[ncol(vimp)] <- as.character(yr)
}

write.csv(vimp, "models/ML/vimp_jack_onset.csv", row.names = F)
write.csv(preds, "models/ML/preds_jack_onset.csv", row.names = F)
# write.csv(vimp, "models/ML/vimp_jack_withprev.csv", row.names = F)
# write.csv(preds, "models/ML/preds_jack_withprev.csv", row.names = F)

preds <- read.csv("models/ML/preds_jack_onset.csv")
preds$obs <- factor(preds$fl_binary, levels = c("0", "1"))
preds$pred <- factor(preds$pred, levels = c("0", "1"))
confusionMatrix(preds$pred, preds$obs)
preds$obs.num <- as.numeric(as.character(preds$fl_binary))
preds$pred.num <- as.numeric(as.character(preds$pred))
R2_jack <- 1 - (sum((preds$obs.num - preds$pred.num)^2, na.rm = T) / sum((preds$obs.num - mean(preds$obs.num, na.rm = T))^2, na.rm = T))

F1_jack <- f1(pred = preds$pred.num, obs = preds$obs.num)


library(dplyr)
vimp_jack <- vimp
# vimp_jack<-read.csv("models/ML/vimp_jack.csv")
vimp_jack <- read.csv("models/ML/vimp_jack_onset.csv")
vimp_jack$mean <- rowMeans(vimp_jack[, 2:ncol(vimp_jack)])
vimp_jack$sd <- apply(vimp_jack[, 2:(ncol(vimp_jack) - 1)], 1, sd)
vimp_jack$sd_1.6 <- vimp_jack$sd * 1.6
```

```{r ml mango onset}
mango <- start_nomiss %>%
  filter(Species_name == "Mangifera indica")
mango_ml <- mango %>%
  dplyr::select(grep("FN", colnames(mango), value = T), fl_binary, year)
mango_ml <- mango_ml %>%
  # dplyr::select(!grep("cs", colnames(mango_ml), value=T))%>%
  mutate(
    mean_temp = scale(FNMean_MeanTemp),
    min_temp = scale(FNMean_MinTemp),
    max_temp = scale(FNMean_MaxTemp),
    solar_rad = scale(FNSum_SolarRad),
    precip = scale(FNSum_PCP),
    dry_days = scale(FNSum_ConsDryDays),
    soil_moisture = scale(FNMean_MeanSM)
  ) %>%
  dplyr::select(!grep("FN", colnames(mango_ml), value = T))


# loop for LOOCV

preds <- data.frame(NA, NA, NA)
colnames(preds) <- c("obs", "pred", "year")
preds <- preds[F, ]

vimp <- data.frame(var = colnames(dplyr::select(mango_ml, -c("fl_binary", "year"))))

for (i in 1:length(unique(mango_ml$year))) {
  yr <- unique(mango_ml$year)[i]
  # subset
  training <- dplyr::select(mango_ml[mango_ml$year != yr, ], -"year")
  testing <- dplyr::select(mango_ml[mango_ml$year == yr, ], -"year")

  # prepping for analysis
  training <- training[complete.cases(training) == T, ]
  testing <- testing[complete.cases(testing) == T, ]
  training$fl_binary <- factor(as.character(training$fl_binary), levels = c("0", "1"))
  testing$fl_binary <- factor(as.character(testing$fl_binary), levels = c("0", "1"))

  names(training) <- make.names(names(training))
  rf_mango <- randomForest::randomForest(fl_binary ~ ., importance = T, data = training)


  predictions.rf <- predict(rf_mango, newdata = testing)
  pred.df <- data.frame(obs = testing[, "fl_binary"], pred = predictions.rf, year = rep(yr, year = length(predictions.rf)))

  preds <- rbind(preds, pred.df)


  # vimp_mean<-caret::varImp(rf_mango)
  vimp_mean <- data.frame(importance(rf_mango, type = 1))
  vimp <- cbind(vimp, vimp_mean[, 1])
  colnames(vimp)[ncol(vimp)] <- as.character(yr)
}

write.csv(vimp, "models/ML/vimp_mango_onset.csv", row.names = F)
write.csv(preds, "models/ML/preds_mango_onset.csv", row.names = F)

preds <- read.csv("models/ML/preds_mango_onset.csv")

head(preds)
preds$obs <- factor(preds$fl_binary, levels = c("0", "1"))
preds$pred <- factor(preds$pred, levels = c("0", "1"))
confusionMatrix(preds$pred, preds$obs)
preds$obs.num <- as.numeric(as.character(preds$fl_binary))
preds$pred.num <- as.numeric(as.character(preds$pred))
R2_mango <- 1 - (sum((preds$obs.num - preds$pred.num)^2) / sum((preds$obs.num - mean(preds$obs.num))^2))

F1_mango <- f1(pred = preds$pred.num, obs = preds$obs.num)


library(dplyr)
vimp_mango <- read.csv("models/ML/vimp_mango_onset.csv")
vimp_mango$mean <- rowMeans(vimp_mango[, 2:ncol(vimp_mango)])
vimp_mango$sd <- apply(vimp_mango[, 2:(ncol(vimp_mango) - 1)], 1, sd)
vimp_mango$sd_1.6 <- vimp_mango$sd * 1.6
```
```{r ml tamarind}
tam <- start_nomiss %>%
  filter(Species_name == "Tamarindus indica")

tam_ml <- tam %>%
  dplyr::select(grep("FN", colnames(tam), value = T), fl_binary, year)
tam_ml <- tam_ml %>%
  # dplyr::select(grep("cs", colnames(tam_ml), value=T), fl_binary, year)%>%
  mutate(
    mean_temp = scale(FNMean_MeanTemp),
    min_temp = scale(FNMean_MinTemp),
    max_temp = scale(FNMean_MaxTemp),
    solar_rad = scale(FNSum_SolarRad),
    precip = scale(FNSum_PCP),
    dry_days = scale(FNSum_ConsDryDays),
    soil_moisture = scale(FNMean_MeanSM)
  ) %>%
  dplyr::select(!grep("FN", colnames(tam_ml), value = T))


# loop for LOOCV

preds <- data.frame(NA, NA, NA)
colnames(preds) <- c("obs", "pred", "year")
preds <- preds[F, ]

vimp <- data.frame(var = colnames(dplyr::select(tam_ml, -c("fl_binary", "year"))))

for (i in 1:length(unique(tam_ml$year))) {
  yr <- unique(tam_ml$year)[i]
  # subset
  training <- dplyr::select(tam_ml[tam_ml$year != yr, ], -"year")
  testing <- dplyr::select(tam_ml[tam_ml$year == yr, ], -"year")

  # prepping for analysis
  training <- training[complete.cases(training) == T, ]
  testing <- testing[complete.cases(testing) == T, ]
  training$fl_binary <- factor(as.character(training$fl_binary), levels = c("0", "1"))
  testing$fl_binary <- factor(as.character(testing$fl_binary), levels = c("0", "1"))

  names(training) <- make.names(names(training))
  rf_tam <- randomForest::randomForest(fl_binary ~ ., importance = T, data = training)


  predictions.rf <- predict(rf_tam, newdata = testing)
  pred.df <- data.frame(obs = testing[, "fl_binary"], pred = predictions.rf, year = rep(yr, year = length(predictions.rf)))

  preds <- rbind(preds, pred.df)


  # vimp_mean<-caret::varImp(rf_tam)
  vimp_mean <- data.frame(importance(rf_tam, type = 1))
  vimp <- cbind(vimp, vimp_mean[, 1])
  colnames(vimp)[ncol(vimp)] <- as.character(yr)
}

write.csv(vimp, "models/ML/vimp_tam_onset.csv", row.names = F)
write.csv(preds, "models/ML/preds_tam_onset.csv", row.names = F)

preds <- read.csv("models/ML/preds_tam_onset.csv")

head(preds)
preds$obs <- factor(preds$fl_binary, levels = c("0", "1"))
preds$pred <- factor(preds$pred, levels = c("0", "1"))
confusionMatrix(preds$pred, preds$obs)
preds$obs.num <- as.numeric(as.character(preds$fl_binary))
preds$pred.num <- as.numeric(as.character(preds$pred))
R2_tam <- 1 - (sum((preds$obs.num - preds$pred.num)^2) / sum((preds$obs.num - mean(preds$obs.num))^2))

F1_tam <- f1(pred = preds$pred.num, obs = preds$obs.num)


library(dplyr)
vimp_tam <- read.csv("models/ML/vimp_tam_onset.csv")
vimp_tam$mean <- rowMeans(vimp_tam[, 2:ncol(vimp_tam)])
vimp_tam$sd <- apply(vimp_tam[, 2:(ncol(vimp_tam) - 1)], 1, sd)
vimp_tam$sd_1.6 <- vimp_tam$sd * 1.6
```
```{r plots from random forest}
labs_list <- list(
  "FNSum_PCP_cs" = "precipitation",
  "elev_cs" = "elevation",
  "FNMean_MaxTemp_cs" = "max temperature",
  "FNMean_MeanSM_cs" = "soil moisture",
  "FNSum_DryDays_cs" = "dry days",
  "aspect_cs" = "aspect",
  "light_cs" = "urbanisation",
  "User_Tree_id" = "Tree ID",
  "FNMean_MinTemp_cs" = "minimum temperature",
  "FNMean_MeanTemp_cs" = "mean temperature",
  "FNSum_SolarRad_cs" = "solar radiation"
)
labs_list2 <- list(
  "precip" = "precipitation",
  "max_temp" = "max temperature",
  "soil_moisture" = "soil moisture",
  "dry_days" = "dry days",
  "min_temp" = "minimum temperature",
  "mean_temp" = "mean temperature",
  "solar_rad" = "solar radiation"
)
library(ggplot2)
# fmt <- "R^2 == %.3f"
fmt2 <- "F1 == %.3f"
lab2 <- sprintf(fmt2, F1_jack)
vimp_jack_plot <- ggplot(vimp_jack, aes(x = reorder(var, mean), weight = mean)) +
  geom_bar(fill = "#1e824c") +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0, size = 1) +
  geom_errorbar(aes(ymin = mean - sd_1.6, ymax = mean + sd_1.6), width = 0, size = 0.5) +
  ylab("Variable Importance") +
  xlab("") +
  scale_x_discrete(labels = labs_list2) +
  guides(fill = F) +
  labs(tag = "a)") +
  ggtitle("Jackfruit flowering probability") +
  annotate("text",
    x = 1, y = 50, label = lab2,
    parse = TRUE, hjust = 0
  ) +
  coord_flip() +
  theme_bw()
vimp_jack_plot

lab2 <- sprintf(fmt2, F1_mango)
vimp_mango_plot <- ggplot(vimp_mango, aes(x = reorder(var, mean), weight = mean)) +
  geom_bar(fill = "#f7ca18") +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0, size = 1) +
  geom_errorbar(aes(ymin = mean - sd_1.6, ymax = mean + sd_1.6), width = 0, size = 0.5) +
  ylab("Variable Importance") +
  xlab("") +
  scale_x_discrete(labels = labs_list2) +
  guides(fill = F) +
  labs(tag = "b)") +
  ggtitle("Mango flowering probability") +
  annotate("text",
    x = 1, y = 50, label = lab2,
    parse = TRUE, hjust = 0
  ) +
  coord_flip() +
  theme_bw()
vimp_mango_plot

lab2 <- sprintf(fmt2, F1_tam)
vimp_tam_plot <- ggplot(vimp_tam, aes(x = reorder(var, mean), weight = mean)) +
  geom_bar(fill = "#a6915c") +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd), width = 0, size = 1) +
  geom_errorbar(aes(ymin = mean - sd_1.6, ymax = mean + sd_1.6), width = 0, size = 0.5) +
  ylab("Variable Importance") +
  xlab("") +
  scale_x_discrete(labels = labs_list2) +
  guides(fill = F) +
  labs(tag = "c)") +
  ggtitle("Tamarind flowering probability") +
  annotate("text",
    x = 1, y = 50, label = lab2,
    parse = TRUE, hjust = 0
  ) +
  coord_flip() +
  theme_bw()
vimp_tam_plot

library(gridExtra)
png("doc/display/Fig2_onset.png", width = 12, height = 4, units = "in", res = 300)
grid.arrange(vimp_jack_plot, vimp_mango_plot, vimp_tam_plot, nrow = 1)
dev.off()